---
title: "2026 World Cup Betting Model"
subtitle: "Goals, match-result probabilities, and data diagnostics for the 2026 FIFA World Cup"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
root <- if (file.exists("data/processed/modeling")) "." else ".."
source(file.path(root, "R", "report_helpers.R"))
read_csv_if_exists <- function(path) {
  if (file.exists(path)) {
    readr::read_csv(path, show_col_types = FALSE)
  } else {
    data.frame()
  }
}
```

<div class="draft-label">DRAFT</div>

<div class="hero-actions">
  <a class="btn" href="#data-sources-and-expense-overview">Data Sources</a>
  <a class="btn" href="reports/01_goals_linear_regression.html">Goals Model</a>
  <a class="btn" href="reports/02_ordinal_result_model.html">Result Model</a>
  <a class="btn" href="reports/04_knn_similarity_model.html">KNN Model</a>
  <a class="btn" href="reports/03_distribution_review.html">Distributions</a>
  <a class="btn" href="#dictionary">Dictionary</a>
</div>

<div class="site-summary">
  <p>This site presents the 2026 World Cup betting model, the data used to support it, and the model results in one shareable place.</p>
  <p>Method details are available through the <strong>Show code</strong> controls. The public page focuses on model inputs, diagnostics, and results.</p>
</div>

```{r hero-metrics, echo=FALSE, results="asis"}
goals_metrics <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "goals_linear_model_metrics.csv"))
ordinal_metrics <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "result_ordinal_model_metrics.csv"))
knn_metrics <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "knn_similarity_model_metrics.csv"))

if (nrow(goals_metrics) > 0 && nrow(ordinal_metrics) > 0) {
  cat(metric_cards(data.frame(
    label = c("Rows Modeled", "Goals RMSE", "Result Accuracy", "Reference Accuracy"),
    value = c(
      fmt_integer(goals_metrics$rows_total[[1]]),
      fmt_number(goals_metrics$test_rmse[[1]], 3),
      fmt_percent(ordinal_metrics$test_accuracy[[1]], 1),
      fmt_percent(ordinal_metrics$baseline_majority_accuracy[[1]], 1)
    ),
    note = c("team-match rows", "typical goal miss", "exact result hit rate", "simple comparison")
  )))
}
```

## Data Sources and Expense Overview

::: {.api-choice}
**Primary live source: API-Football by API-Sports**

The model uses API-Football for World Cup fixtures, teams, standings, player/squad data, injuries, lineups, match events, player statistics, odds, and predictions where coverage is available.
:::

```{r data-sources, echo=FALSE}
data_sources <- data.frame(
  layer = c(
    "Live football API",
    "Historical match results",
    "Player and squad context",
    "Venue, weather, and news context",
    "Publishing"
  ),
  source = c(
    "API-Football / API-Sports",
    "Public international results and goalscorer datasets",
    "API-Football plus public squad references",
    "Open venue/weather/news sources",
    "GitHub Pages"
  ),
  use_in_model = c(
    "Fixtures, teams, standings, injuries, lineups, odds, and match details",
    "Team form, goals scored, results, Elo-style strength features",
    "Roster strength, player goal history, availability indicators",
    "Travel, venue, weather, and news-derived context features",
    "Shareable HTML reports and model summaries"
  )
)
display_table(data_sources, caption = "Data sources used by the modeling workflow.")
```

The public site is hosted without an additional hosting bill. The live football data subscription is the main recurring project expense, and the project is kept within the monthly budget cap.

## Model Iterations

The model section is organized around interpretable iterations. Each model card names the prediction target, explains the method, and lists the checks that determine whether the model is trustworthy.

::: {.model-blueprint}
::: {.model-card}
### 1. Simple OLS Goals Model

**Question:** how many goals should a team score in one match?

The simple goals model starts with team strength difference:

$$
\text{Goals} = \beta_0 + \beta_1\text{Strength Difference} + \epsilon
$$

In words, `beta_1` is the average change in expected goals for a one-unit increase in strength difference.

**Why start here:** OLS is easy to explain, fast to diagnose, and gives a clear benchmark before count models.

**Main issues to check:**

- Goals are counts, so OLS can predict impossible values below 0.
- Soccer scores are low and skewed, so residuals may not look normal.
- Error variance can change by team strength or opponent strength.
- Team-match rows are not fully independent because each match creates two rows.
- Missing roster, injury, rest, venue, and market data can bias coefficients.
:::

::: {.model-card}
### 2. Three-Outcome Result Model

**Question:** what is the probability of win, draw, or loss?

The first three-outcome model is an ordinal logistic regression because the team result has a natural order:

$$
\text{loss} < \text{draw} < \text{win}
$$

The model uses cumulative logits:

$$
\text{logit}\{P(\text{Result} \leq k)\}
= \theta_k - \eta
$$

with:

$$
\eta = \beta_1\text{Strength Difference}
+ \beta_2\text{Roster Strength}
+ \beta_3\text{Rest}
+ \beta_4\text{Venue}
$$

The output is three probabilities:

$$
P(\text{loss}), \quad P(\text{draw}), \quad P(\text{win})
$$

**Main issues to check:**

- The proportional-odds assumption may be too strong.
- Draws are not simply halfway between loss and win.
- A multinomial logit may fit better if win, draw, and loss behave as separate categories.
- Calibration matters more than raw accuracy for betting comparison.
- Market odds should be the benchmark once API odds are available.
:::

::: {.model-card}
### 3. KNN Similarity Model

**Question:** what happened in historical matchups that looked similar?

KNN stands for **K-Nearest Neighbors**. The model finds the closest historical team-match rows, then averages those neighbors.

**Outputs:**

- Expected goals from the average goals scored by similar teams.
- Loss, draw, and win probabilities from the neighbor outcomes.
- A flexible comparison against the regression models.

**Main issues to check:**

- Similarity depends on which features are included.
- Features must be standardized so one large-scale number does not dominate.
- The choice of `K` controls the smoothness of predictions.
- KNN can copy noise when the neighbor set is too small.
- KNN is harder to explain with coefficients because it is case-based, not equation-based.

<a class="project-button" href="reports/04_knn_similarity_model.html">Open KNN report</a>
:::
:::

## Dictionary

Use this as the quick reference for acronyms and modeling terms that appear across the site.

| Term | Plain-English Meaning | Why It Matters Here |
|---|---|---|
| Target variable | The thing the model is trying to predict. | For the goals model, the target is goals scored. For the result model, the target is win, draw, or loss. |
| Predictor variable | An input used to make the prediction. | Team strength, opponent strength, rest, venue, injuries, and odds can all be predictors. |
| Regression | A model that estimates how a target changes with predictors. | It gives a structured way to turn soccer information into expected goals or probabilities. |
| OLS | Ordinary Least Squares. A line-fitting method that chooses coefficients by minimizing squared prediction errors. | It is the first goals benchmark because it is easy to explain and diagnose. |
| RMSE | Root Mean Squared Error. A typical prediction miss, measured in the same units as the target. | If goals RMSE is 1.46, the model is often off by roughly 1 to 2 goals, with big misses punished more. Lower is better. |
| Coefficient | The estimated effect of one predictor, holding the other predictors fixed. | A positive strength coefficient means stronger teams are expected to score more, all else equal. |
| Residual | Observed value minus predicted value. | Residual plots show whether the model is missing patterns. |
| Logit | A transformation that converts a probability into log odds. | Logistic models use logits so probabilities stay between 0 and 1. |
| Odds | Probability of an event divided by probability of it not happening. | Model probabilities can be compared with betting-market odds after converting both to the same scale. |
| Ordinal logistic regression | A logistic model for ordered outcomes. | It fits `loss < draw < win` while producing probabilities for all three outcomes. |
| Multinomial logistic regression | A logistic model for three or more categories without assuming an order. | It is the backup if draws behave like their own special category rather than the middle of loss and win. |
| KNN | K-Nearest Neighbors. A model that predicts from similar historical examples. | It lets the project compare regression models against a flexible similarity-based method. |
| Accuracy | Share of predictions where the top predicted class matches the actual result. | Useful for a quick check, but not enough for betting because probability quality matters too. |
| Calibration | Whether predicted probabilities match real frequencies. | If the model says 60% win chance, well-calibrated teams should win about 60% of those cases. |
| Log loss | A probability score that penalizes confident wrong predictions. | Better than accuracy when the model outputs win/draw/loss probabilities. Lower is better. |
| Brier score | Average squared error between predicted probabilities and actual outcomes. | Another way to judge probability quality. Lower is better. |
| Test set | Data held out from fitting and used only to evaluate performance. | It helps estimate how the model performs on games it has not already seen. |
| Overfitting | When a model learns noise in past data instead of repeatable signal. | A model can look good historically and still fail on new World Cup matches. |
| Data grain | What one row means in a dataset. | The main modeling grain is one team in one match. One match creates two team-match rows. |
| API | Application Programming Interface. A structured way to pull data from a provider. | API-Football is the primary live football data source. |

## Project Map

::: {.panel-tabset}

## About

This project builds a reproducible data and modeling workflow for the 2026 FIFA World Cup.

The workflow has four layers:

- **Data collection:** match history, Elo ratings, 2026 venues, weather, player metadata, odds, and news signals.
- **Data storage:** local files and DuckDB tables for modeling.
- **Modeling:** interpretable models in R, with Python handling source refreshes and heavier data jobs.
- **Publishing:** Quarto renders a shareable GitHub Pages site from the local analysis.

## Data Exploration

The starting modeling unit is one row per team-match. A single game creates two observations, one for each team.

```{r goals-sample}
goals_sample <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "goals_linear_model_sample_1000.csv"))

if (nrow(goals_sample) > 0) {
  goals_sample |>
    dplyr::select(
      date,
      team,
      opponent,
      tournament,
      y_goals_for,
      pre_elo,
      opponent_pre_elo,
      pre_match_expected_result
    ) |>
    dplyr::mutate(
      pre_match_expected_result = fmt_percent(pre_match_expected_result, 1),
      pre_elo = round(pre_elo, 0),
      opponent_pre_elo = round(opponent_pre_elo, 0)
    ) |>
    display_table(caption = "Sample of the team-match modeling grain.", max_rows = 8)
} else {
  display_table(empty_note("Sample rows are available after the goals model refresh."))
}
```

```{r goals-plot, fig.cap="Distribution of goals in the public sample used for the goals model."}
if (nrow(goals_sample) > 0 && requireNamespace("ggplot2", quietly = TRUE)) {
  ggplot2::ggplot(goals_sample, ggplot2::aes(x = y_goals_for)) +
    ggplot2::geom_histogram(binwidth = 1, boundary = -0.5, fill = "#157878", color = "white") +
    ggplot2::labs(
      title = "Goals scored by team-match row",
      x = "Goals scored",
      y = "Team-match rows"
    ) +
    ggplot2::theme_minimal(base_size = 12)
} else {
  plot.new()
  text(0.5, 0.5, "Goals sample is available after the model refresh.")
}
```

## Modeling

The model suite has three core outputs:

- **Goals OLS:** predicts goals scored.
- **Ordinal logistic:** predicts `loss < draw < win`.
- **KNN similarity:** predicts goals and result probabilities from nearby historical matchups.

The reports show the model target, data grain, math, coefficients, predictions, and diagnostics.

```{r model-summary}
if (nrow(goals_metrics) > 0 && nrow(ordinal_metrics) > 0) {
  model_rows <- data.frame(
    model = c("Goals OLS", "Ordinal result logistic"),
    target = c("team goals", "loss < draw < win"),
    rows = fmt_integer(c(goals_metrics$rows_total[[1]], ordinal_metrics$rows_total[[1]])),
    primary_metric = c("Test RMSE", "Test accuracy"),
    value = c(
      fmt_number(goals_metrics$test_rmse[[1]], 3),
      fmt_percent(ordinal_metrics$test_accuracy[[1]], 1)
    ),
    context = c(
      paste("Mean test goals:", fmt_number(goals_metrics$mean_y_test[[1]], 3)),
      paste("Reference accuracy:", fmt_percent(ordinal_metrics$baseline_majority_accuracy[[1]], 1))
    )
  )
  if (nrow(knn_metrics) > 0) {
    model_rows <- rbind(
      model_rows,
      data.frame(
        model = "KNN similarity",
        target = "team goals and loss/draw/win",
        rows = fmt_integer(knn_metrics$rows_total[[1]]),
        primary_metric = "Accuracy / RMSE",
        value = paste0(
          fmt_percent(knn_metrics$result_accuracy[[1]], 1),
          " / ",
          fmt_number(knn_metrics$goals_rmse[[1]], 3)
        ),
        context = paste("Neighbors:", knn_metrics$k_neighbors[[1]])
      )
    )
  }
  model_rows |>
    display_table(caption = "Current model performance.")
} else {
  display_table(empty_note("Model metrics are available after the model refresh."))
}
```

## Data

The analysis uses a local modeling database and publishes the resulting metadata, diagnostics, and model outputs here.

```{r table-inventory}
inventory <- read_csv_if_exists(file.path(root, "data", "processed", "metadata", "table_inventory.csv"))

if (nrow(inventory) > 0) {
  inventory |>
    dplyr::arrange(table_name) |>
    dplyr::mutate(row_count = fmt_integer(row_count)) |>
    display_table(caption = "Modeling tables available for analysis.", max_rows = 16)
} else {
  display_table(empty_note("Table inventory is available after the metadata refresh."))
}
```

## Display Standards

<div class="principles">
  <div class="principle"><strong>One grain per table</strong>Rows should mean one thing: team-match, match, player, venue, or API pull.</div>
  <div class="principle"><strong>Readable labels</strong>Keep machine-safe names in source files; convert them to readable labels only in reports.</div>
  <div class="principle"><strong>Round for display only</strong>Do not round model inputs or stored outputs. Round the rendered table so readers can scan it.</div>
  <div class="principle"><strong>Publication boundary</strong>Present summaries and rendered results; keep operational files out of the public site.</div>
</div>

:::

## Reports

::: {.columns}
::: {.column width="33%"}
::: {.project-card}
### Goals Model

OLS model for team goals scored from pre-match Elo, opponent Elo, expected result, home listing, and neutral-site status.

<a class="project-button" href="reports/01_goals_linear_regression.html">Open goals report</a>
:::
:::

::: {.column width="33%"}
::: {.project-card}
### Result Model

Ordinal logistic model for `loss < draw < win`, producing win/draw/loss probability estimates.

<a class="project-button" href="reports/02_ordinal_result_model.html">Open result report</a>
:::
:::

::: {.column width="33%"}
::: {.project-card}
### KNN Model

Similarity model that compares each team-match row with nearby historical cases.

<a class="project-button" href="reports/04_knn_similarity_model.html">Open KNN report</a>
:::
:::
:::

::: {.project-card}
### Distribution Review

Distribution notes for deciding whether soccer features are counts, continuous variables, ordered outcomes, probabilities, or mixture processes.

<a class="project-button" href="reports/03_distribution_review.html">Open distribution review</a>
:::
