---
title: "2026 World Cup Betting Model"
subtitle: "Goals, match-result probabilities, and data diagnostics for the 2026 FIFA World Cup"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
root <- if (file.exists("data/processed/modeling")) "." else ".."
source(file.path(root, "R", "report_helpers.R"))
read_csv_if_exists <- function(path) {
  if (file.exists(path)) {
    readr::read_csv(path, show_col_types = FALSE)
  } else {
    data.frame()
  }
}
```

<div class="draft-label">DRAFT</div>

<div class="hero-actions">
  <a class="btn" href="#data-sources-and-expense-overview">Data Sources</a>
  <a class="btn" href="reports/00_data_overview.html">Data Overview</a>
  <a class="btn" href="reports/01_goals_linear_regression.html">Goals Model</a>
  <a class="btn" href="reports/05_poisson_goals_model.html">Poisson Model</a>
  <a class="btn" href="reports/02_ordinal_result_model.html">Result Model</a>
  <a class="btn" href="reports/04_knn_similarity_model.html">KNN Model</a>
  <a class="btn" href="reports/06_regression_diagnostics.html">Diagnostics</a>
  <a class="btn" href="reports/07_fixture_predictions.html">Predictions</a>
  <a class="btn" href="reports/08_matchday_predictions.html">Matchday</a>
  <a class="btn" href="reports/03_distribution_review.html">Distributions</a>
  <a class="btn" href="#dictionary">Dictionary</a>
</div>

<div class="site-summary">
  <p>This site presents the 2026 World Cup betting model, the data used to support it, and the model results in one shareable place.</p>
  <p>Method details are available through the <strong>Show code</strong> controls. The public page focuses on model inputs, diagnostics, and results.</p>
</div>

```{r hero-metrics, echo=FALSE, results="asis"}
goals_metrics <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "goals_linear_model_metrics.csv"))
poisson_metrics <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "goals_poisson_model_metrics.csv"))
ordinal_metrics <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "result_ordinal_model_metrics.csv"))
knn_metrics <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "knn_similarity_model_metrics.csv"))
diagnostics <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "regression_diagnostics_summary.csv"))
fixture_metrics <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "world_cup_2026_fixture_prediction_metrics.csv"))

metric_value <- function(metrics, metric_name) {
  if (nrow(metrics) == 0 || !"metric" %in% names(metrics)) {
    return(NA_real_)
  }
  value <- metrics$value[metrics$metric == metric_name]
  if (length(value) == 0) NA_real_ else value[[1]]
}

diagnostic_value <- function(summary, check_name) {
  if (nrow(summary) == 0 || !"check" %in% names(summary)) {
    return(NA_real_)
  }
  value <- summary$value[summary$check == check_name]
  if (length(value) == 0) NA_real_ else value[[1]]
}

if (nrow(goals_metrics) > 0 && nrow(ordinal_metrics) > 0) {
  poisson_random <- if (nrow(poisson_metrics) > 0) {
    poisson_metrics[poisson_metrics$validation_type == "random_80_20", ]
  } else {
    data.frame()
  }
  cat(metric_cards(data.frame(
    label = c("Rows Modeled", "OLS Goals RMSE", "Poisson Goals RMSE", "Result Accuracy", "Fixtures Scored", "Max VIF"),
    value = c(
      fmt_integer(goals_metrics$rows_total[[1]]),
      fmt_number(goals_metrics$test_rmse[[1]], 3),
      if (nrow(poisson_random) > 0) fmt_number(poisson_random$test_rmse[[1]], 3) else "pending",
      fmt_percent(ordinal_metrics$test_accuracy[[1]], 1),
      if (nrow(fixture_metrics) > 0) fmt_integer(metric_value(fixture_metrics, "fixtures_scored")) else "pending",
      if (nrow(diagnostics) > 0) fmt_number(diagnostic_value(diagnostics, "Collinearity"), 2) else "pending"
    ),
    note = c(
      "team-match rows",
      "typical goal miss",
      "count-model benchmark",
      "exact result hit rate",
      "2026 fixtures",
      "collinearity check"
    )
  )))
}
```

## Data Sources and Expense Overview

::: {.api-choice}
**Model data foundation**

The model uses open historical match results, goalscorer history, Elo-style team strength, public 2026 fixture/venue references, weather, news metadata, and connected football API coverage summaries. The public site presents model-ready summaries and results, not raw operational files.
:::

```{r data-sources, echo=FALSE}
data_sources <- data.frame(
  layer = c(
    "Historical match results",
    "Team strength history",
    "Player and squad context",
    "Venue, weather, and news context",
    "Live football API",
    "Publishing"
  ),
  source = c(
    "Public international results and goalscorer datasets",
    "Derived Elo-style ratings from historical results",
    "Public squad references and Wikidata enrichment",
    "Open venue/weather/news sources",
    "API-Football / API-Sports",
    "GitHub Pages"
  ),
  use_in_model = c(
    "Team form, goals scored, results, Elo-style strength features",
    "Pre-match strength, opponent strength, and expected result",
    "Roster strength, player goal history, availability indicators",
    "Travel, venue, weather, and news-derived context features",
    "Later enrichment for teams, standings, injuries, lineups, odds, and match details",
    "Shareable HTML reports and model summaries"
  )
)
display_table(data_sources, caption = "Data sources used by the modeling workflow.")
```

This model version is reproducible with free data and connected API coverage metadata. Detailed match, player, lineup, injury, and odds fields are structured enrichment layers for future scoring updates.

## Model Iterations

The model section is organized around interpretable iterations. Each model card names the prediction target, explains the method, and lists the checks that determine whether the model is trustworthy.

::: {.model-blueprint}
::: {.model-card}
### 1. Simple OLS Goals Model

**Question:** how many goals should a team score in one match?

The simple goals model starts with team strength difference:

$$
\text{Goals} = \beta_0 + \beta_1\text{Strength Difference} + \epsilon
$$

In words, `beta_1` is the average change in expected goals for a one-unit increase in strength difference.

**Why start here:** OLS is easy to explain, fast to diagnose, and gives a clear benchmark before count models.

**Main issues to check:**

- Goals are counts, so OLS can predict impossible values below 0.
- Soccer scores are low and skewed, so residuals may not look normal.
- Error variance can change by team strength or opponent strength.
- Team-match rows are not fully independent because each match creates two rows.
- Missing roster, injury, rest, venue, and market data can bias coefficients.
:::

::: {.model-card}
### 2. Poisson Goals Model

**Question:** how many goals should a team score when the target is a count?

The Poisson model predicts a positive expected-goals number:

$$
\log(\lambda)
= \beta_0
+ \beta_1\text{Team Elo}
+ \beta_2\text{Opponent Elo}
+ \beta_3\text{Expected Result}
+ \beta_4\text{Match Context}
$$

In words, `lambda` is expected goals. This is a better shape for soccer scoring than OLS because goals cannot be negative.

**Main issues to check:**

- The basic Poisson assumes the mean and variance are tied together.
- Soccer can be overdispersed, meaning the variance may be larger than the mean.
- A negative binomial model may be needed if Poisson is too tight.
- Forward-time validation matters more than random validation for future matches.

<a class="project-button" href="reports/05_poisson_goals_model.html">Open Poisson report</a>
:::

::: {.model-card}
### 3. Three-Outcome Result Model

**Question:** what is the probability of win, draw, or loss?

The first three-outcome model is an ordinal logistic regression because the team result has a natural order:

$$
\text{loss} < \text{draw} < \text{win}
$$

The model uses cumulative logits:

$$
\text{logit}\{P(\text{Result} \leq k)\}
= \theta_k - \eta
$$

with:

$$
\eta = \beta_1\text{Strength Difference}
+ \beta_2\text{Roster Strength}
+ \beta_3\text{Rest}
+ \beta_4\text{Venue}
$$

The output is three probabilities:

$$
P(\text{loss}), \quad P(\text{draw}), \quad P(\text{win})
$$

**Main issues to check:**

- The proportional-odds assumption may be too strong.
- Draws are not simply halfway between loss and win.
- A multinomial logit may fit better if win, draw, and loss behave as separate categories.
- Calibration matters more than raw accuracy for betting comparison.
- Market odds should be the benchmark once API odds are available.
:::

::: {.model-card}
### 4. KNN Similarity Model

**Question:** what happened in historical matchups that looked similar?

KNN stands for **K-Nearest Neighbors**. The model finds the closest historical team-match rows, then averages those neighbors.

**Outputs:**

- Expected goals from the average goals scored by similar teams.
- Loss, draw, and win probabilities from the neighbor outcomes.
- A flexible comparison against the regression models.

**Main issues to check:**

- Similarity depends on which features are included.
- Features must be standardized so one large-scale number does not dominate.
- The choice of `K` controls the smoothness of predictions.
- KNN can copy noise when the neighbor set is too small.
- KNN is harder to explain with coefficients because it is case-based, not equation-based.

<a class="project-button" href="reports/04_knn_similarity_model.html">Open KNN report</a>
:::

::: {.model-card}
### 5. Regression Diagnostics

**Question:** are the regression results stable enough to interpret?

Before the regression results are shown, the workflow checks collinearity, residual patterns, heteroskedasticity, and Poisson overdispersion.

**Outputs:**

- VIF values for predictor overlap.
- Durbin-Watson and lagged residual correlation.
- Breusch-Pagan and Jarque-Bera diagnostic checks.
- Poisson dispersion checks for goals.

<a class="project-button" href="reports/06_regression_diagnostics.html">Open diagnostics report</a>
:::

::: {.model-card}
### 6. Fixture Predictions

**Question:** what does the current model predict for the 2026 fixture list?

The fixture table applies the trained local models to 2026 matches and shows expected goals, win/draw/loss probabilities, completed-fixture checks, and weather context.

<a class="project-button" href="reports/07_fixture_predictions.html">Open predictions report</a>
:::

::: {.model-card}
### 7. Matchday Board

**Question:** what should be checked right before kickoff?

The matchday board combines predicted winner, win/draw/loss probabilities, expected goals, most likely score, totals, lineup status, and yellow-card coverage in one table.

<a class="project-button" href="reports/08_matchday_predictions.html">Open matchday board</a>
:::
:::

## Dictionary

Use this as the quick reference for acronyms and modeling terms that appear across the site.

| Term | Plain-English Meaning | Why It Matters Here |
|---|---|---|
| Target variable | The thing the model is trying to predict. | For the goals model, the target is goals scored. For the result model, the target is win, draw, or loss. |
| Predictor variable | An input used to make the prediction. | Team strength, opponent strength, rest, venue, injuries, and odds can all be predictors. |
| Regression | A model that estimates how a target changes with predictors. | It gives a structured way to turn soccer information into expected goals or probabilities. |
| OLS | Ordinary Least Squares. A line-fitting method that chooses coefficients by minimizing squared prediction errors. | It is the first goals benchmark because it is easy to explain and diagnose. |
| Poisson regression | A regression model for count outcomes. | Goals are counts, so Poisson is a natural next model after OLS. |
| Lambda | The expected count in a Poisson model. | In the Poisson goals model, lambda is expected team goals. |
| RMSE | Root Mean Squared Error. A typical prediction miss, measured in the same units as the target. | If goals RMSE is 1.46, the model is often off by roughly 1 to 2 goals, with big misses punished more. Lower is better. |
| MAE | Mean Absolute Error. The average size of the prediction miss, ignoring direction. | Easier to read than RMSE because every miss is weighted evenly. Lower is better. |
| Coefficient | The estimated effect of one predictor, holding the other predictors fixed. | A positive strength coefficient means stronger teams are expected to score more, all else equal. |
| Residual | Observed value minus predicted value. | Residual plots show whether the model is missing patterns. |
| VIF | Variance Inflation Factor. A check for predictors that repeat the same information. | High VIF makes individual regression coefficients harder to trust. |
| Autocorrelation | A pattern where one residual is related to the residual before it. | Strong autocorrelation means model errors are not acting independently. |
| Durbin-Watson | A residual autocorrelation check. Values near 2 are usually cleaner. | It is a quick test for sequential structure in ordered model errors. |
| Heteroskedasticity | Uneven residual spread across predictor values. | It can make OLS standard errors too optimistic. |
| Overdispersion | When a count outcome varies more than a basic Poisson model expects. | If goals are overdispersed, a negative-binomial or paired-goal model may handle uncertainty better. |
| Logit | A transformation that converts a probability into log odds. | Logistic models use logits so probabilities stay between 0 and 1. |
| Odds | Probability of an event divided by probability of it not happening. | Model probabilities can be compared with betting-market odds after converting both to the same scale. |
| Ordinal logistic regression | A logistic model for ordered outcomes. | It fits `loss < draw < win` while producing probabilities for all three outcomes. |
| Multinomial logistic regression | A logistic model for three or more categories without assuming an order. | It is the backup if draws behave like their own special category rather than the middle of loss and win. |
| KNN | K-Nearest Neighbors. A model that predicts from similar historical examples. | It lets the project compare regression models against a flexible similarity-based method. |
| Accuracy | Share of predictions where the top predicted class matches the actual result. | Useful for a quick check, but not enough for betting because probability quality matters too. |
| Calibration | Whether predicted probabilities match real frequencies. | If the model says 60% win chance, well-calibrated teams should win about 60% of those cases. |
| Log loss | A probability score that penalizes confident wrong predictions. | Better than accuracy when the model outputs win/draw/loss probabilities. Lower is better. |
| Brier score | Average squared error between predicted probabilities and actual outcomes. | Another way to judge probability quality. Lower is better. |
| Test set | Data held out from fitting and used only to evaluate performance. | It helps estimate how the model performs on games it has not already seen. |
| Forward-time validation | Training on older games and testing on newer games. | This is closer to the real task: using past matches to predict future matches. |
| Overfitting | When a model learns noise in past data instead of repeatable signal. | A model can look good historically and still fail on new World Cup matches. |
| Data grain | What one row means in a dataset. | The main modeling grain is one team in one match. One match creates two team-match rows. |
| API | Application Programming Interface. A structured way to pull data from a provider. | API-Football is the live enrichment source once the baseline model is stable. |
| Ensemble | A prediction that combines more than one model. | The matchday winner uses ordinal result probabilities plus Poisson score-derived probabilities. |
| O2.5 | Over 2.5 goals. The match has at least 3 total goals. | Common soccer totals view derived from the score grid. |
| BTTS | Both teams to score. | Another match-level probability derived from expected goals. |

## Project Map

::: {.panel-tabset}

## About

This project builds a reproducible data and modeling workflow for the 2026 FIFA World Cup.

The workflow has four layers:

- **Data collection:** free match history, Elo-style ratings, 2026 venues, weather, player metadata, and news signals.
- **Data storage:** local files and DuckDB tables for modeling.
- **Modeling:** interpretable models in R, with Python handling source refreshes and heavier data jobs.
- **Publishing:** Quarto renders a shareable GitHub Pages site from the local analysis.

## Data Exploration

The starting modeling unit is one row per team-match. A single game creates two observations, one for each team.

```{r goals-sample}
goals_sample <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "goals_linear_model_sample_1000.csv"))

if (nrow(goals_sample) > 0) {
  goals_sample |>
    dplyr::select(
      date,
      team,
      opponent,
      tournament,
      y_goals_for,
      pre_elo,
      opponent_pre_elo,
      pre_match_expected_result
    ) |>
    dplyr::mutate(
      pre_match_expected_result = fmt_percent(pre_match_expected_result, 1),
      pre_elo = round(pre_elo, 0),
      opponent_pre_elo = round(opponent_pre_elo, 0)
    ) |>
    display_table(caption = "Sample of the team-match modeling grain.", max_rows = 8)
} else {
  display_table(empty_note("Sample rows are available after the goals model refresh."))
}
```

```{r goals-plot, fig.cap="Distribution of goals in the public sample used for the goals model."}
if (nrow(goals_sample) > 0 && requireNamespace("ggplot2", quietly = TRUE)) {
  ggplot2::ggplot(goals_sample, ggplot2::aes(x = y_goals_for)) +
    ggplot2::geom_histogram(binwidth = 1, boundary = -0.5, fill = "#157878", color = "white") +
    ggplot2::labs(
      title = "Goals scored by team-match row",
      x = "Goals scored",
      y = "Team-match rows"
    ) +
    ggplot2::theme_minimal(base_size = 12)
} else {
  plot.new()
  text(0.5, 0.5, "Goals sample is available after the model refresh.")
}
```

## Modeling

The model suite has four core outputs:

- **Goals OLS:** predicts goals scored.
- **Poisson goals:** predicts expected goals as a count.
- **Ordinal logistic:** predicts `loss < draw < win`.
- **KNN similarity:** predicts goals and result probabilities from nearby historical matchups.

The reports show the model target, data grain, math, coefficients, predictions, and diagnostics.

```{r model-summary}
if (nrow(goals_metrics) > 0 && nrow(ordinal_metrics) > 0) {
  model_rows <- data.frame(
    model = c("Goals OLS", "Ordinal result logistic"),
    target = c("team goals", "loss < draw < win"),
    rows = fmt_integer(c(goals_metrics$rows_total[[1]], ordinal_metrics$rows_total[[1]])),
    primary_metric = c("Test RMSE", "Test accuracy"),
    value = c(
      fmt_number(goals_metrics$test_rmse[[1]], 3),
      fmt_percent(ordinal_metrics$test_accuracy[[1]], 1)
    ),
    context = c(
      paste("Mean test goals:", fmt_number(goals_metrics$mean_y_test[[1]], 3)),
      paste("Reference accuracy:", fmt_percent(ordinal_metrics$baseline_majority_accuracy[[1]], 1))
    )
  )
  if (nrow(poisson_metrics) > 0) {
    poisson_random <- poisson_metrics[poisson_metrics$validation_type == "random_80_20", ]
    poisson_forward <- poisson_metrics[poisson_metrics$validation_type == "train_through_2018_test_2019_plus", ]
    if (nrow(poisson_random) > 0) {
      model_rows <- rbind(
        model_rows,
        data.frame(
          model = "Poisson goals",
          target = "team goals",
          rows = fmt_integer(poisson_random$rows_total[[1]]),
          primary_metric = "RMSE",
          value = fmt_number(poisson_random$test_rmse[[1]], 3),
          context = if (nrow(poisson_forward) > 0) {
            paste("Forward-time RMSE:", fmt_number(poisson_forward$test_rmse[[1]], 3))
          } else {
            "Count-model benchmark"
          }
        )
      )
    }
  }
  if (nrow(knn_metrics) > 0) {
    model_rows <- rbind(
      model_rows,
      data.frame(
        model = "KNN similarity",
        target = "team goals and loss/draw/win",
        rows = fmt_integer(knn_metrics$rows_total[[1]]),
        primary_metric = "Accuracy / RMSE",
        value = paste0(
          fmt_percent(knn_metrics$result_accuracy[[1]], 1),
          " / ",
          fmt_number(knn_metrics$goals_rmse[[1]], 3)
        ),
        context = paste("Neighbors:", knn_metrics$k_neighbors[[1]])
      )
    )
  }
  if (nrow(diagnostics) > 0) {
    model_rows <- rbind(
      model_rows,
      data.frame(
        model = "Regression diagnostics",
        target = "model quality checks",
        rows = fmt_integer(goals_metrics$rows_total[[1]]),
        primary_metric = "Max VIF",
        value = fmt_number(diagnostic_value(diagnostics, "Collinearity"), 3),
        context = paste("Poisson dispersion:", fmt_number(diagnostic_value(diagnostics, "Poisson overdispersion"), 3))
      )
    )
  }
  if (nrow(fixture_metrics) > 0) {
    model_rows <- rbind(
      model_rows,
      data.frame(
        model = "2026 fixture scoring",
        target = "match goals and result probabilities",
        rows = fmt_integer(metric_value(fixture_metrics, "fixtures_scored")),
        primary_metric = "Completed fixture check",
        value = paste0(
          "Poisson RMSE ",
          fmt_number(metric_value(fixture_metrics, "poisson_completed_fixture_rmse"), 3)
        ),
        context = paste("Weather rows:", fmt_integer(metric_value(fixture_metrics, "fixtures_with_weather_context")))
      )
    )
  }
  model_rows |>
    display_table(caption = "Current model performance.")
} else {
  display_table(empty_note("Model metrics are available after the model refresh."))
}
```

## Data

The analysis uses a local modeling database and publishes the resulting metadata, diagnostics, and model outputs here.

```{r table-inventory}
inventory <- read_csv_if_exists(file.path(root, "data", "processed", "metadata", "table_inventory.csv"))

if (nrow(inventory) > 0) {
  inventory |>
    dplyr::arrange(table_name) |>
    dplyr::mutate(row_count = fmt_integer(row_count)) |>
    display_table(caption = "Modeling tables available for analysis.", max_rows = 16)
} else {
  display_table(empty_note("Table inventory is available after the metadata refresh."))
}
```

## Display Standards

<div class="principles">
  <div class="principle"><strong>One grain per table</strong>Rows should mean one thing: team-match, match, player, venue, or API pull.</div>
  <div class="principle"><strong>Readable labels</strong>Keep machine-safe names in source files; convert them to readable labels only in reports.</div>
  <div class="principle"><strong>Round for display only</strong>Do not round model inputs or stored outputs. Round the rendered table so readers can scan it.</div>
  <div class="principle"><strong>Publication boundary</strong>Present summaries and rendered results; keep operational files out of the public site.</div>
</div>

:::

## Reports

::: {.columns}
::: {.column width="25%"}
::: {.project-card}
### Data Overview

Available datasets, sample modeling rows, metadata, and simple diagnostics.

<a class="project-button" href="reports/00_data_overview.html">Open data report</a>
:::
:::

::: {.column width="25%"}
::: {.project-card}
### Goals Model

OLS model for team goals scored from pre-match Elo, opponent Elo, expected result, home listing, and neutral-site status.

<a class="project-button" href="reports/01_goals_linear_regression.html">Open goals report</a>
:::
:::

::: {.column width="25%"}
::: {.project-card}
### Poisson Model

Count model for expected team goals, with random and forward-time validation.

<a class="project-button" href="reports/05_poisson_goals_model.html">Open Poisson report</a>
:::
:::

::: {.column width="25%"}
::: {.project-card}
### Result Model

Ordinal logistic model for `loss < draw < win`, producing win/draw/loss probability estimates.

<a class="project-button" href="reports/02_ordinal_result_model.html">Open result report</a>
:::
:::
:::

::: {.columns}
::: {.column width="50%"}
::: {.project-card}
### KNN Model

Similarity model that compares each team-match row with nearby historical cases.

<a class="project-button" href="reports/04_knn_similarity_model.html">Open KNN report</a>
:::
:::

::: {.column width="50%"}
::: {.project-card}
### Diagnostics

Regression checks for collinearity, residual structure, heteroskedasticity, and count-model dispersion.

<a class="project-button" href="reports/06_regression_diagnostics.html">Open diagnostics report</a>
:::
:::
:::

::: {.columns}
::: {.column width="50%"}
::: {.project-card}
### Fixture Predictions

Expected goals, win/draw/loss probabilities, completed-fixture checks, and weather context for the 2026 fixture list.

<a class="project-button" href="reports/07_fixture_predictions.html">Open predictions report</a>
:::
:::

::: {.column width="50%"}
::: {.project-card}
### Matchday Board

One table for pre-match checks: winner, outcome probabilities, goals, totals, lineups, cards coverage, and venue context.

<a class="project-button" href="reports/08_matchday_predictions.html">Open matchday board</a>
:::
:::
:::

::: {.columns}
::: {.column width="50%"}
::: {.project-card}
### Distribution Review

Distribution notes for deciding whether soccer features are counts, continuous variables, ordered outcomes, probabilities, or mixture processes.

<a class="project-button" href="reports/03_distribution_review.html">Open distribution review</a>
:::
:::
:::
