---
title: "World Cup 2026 Forecasting Model"
subtitle: "Match probabilities, projected scores, and tournament paths."
description: "A prediction-first World Cup 2026 forecasting site with match probabilities, expected scores, tournament bracket projections, and model performance monitoring."
image: assets/social-preview.svg
toc: false
number-sections: false
code-tools: false
code-fold: false
execute:
  echo: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
root <- if (file.exists("data/processed/modeling")) "." else ".."
source(file.path(root, "R", "report_helpers.R"))
source(file.path(root, "R", "matchday_components.R"))
bundle <- load_matchday_bundle(root)
board <- bundle$board
summary <- bundle$summary
```

```{r hero, results="asis"}
cat(render_home_hero(summary, board))
cat(render_quick_status(summary))
```

```{r today, results="asis"}
cat(render_today_section(board, summary))
```

```{r next-match, results="asis"}
cat(render_next_match_section(board))
```

```{r upcoming, results="asis"}
cat(render_upcoming_section(board, limit = 6))
```

```{r model-performance, results="asis"}
cat(render_performance_section(bundle, compact = TRUE))
```

```{r bracket-preview, results="asis"}
cat('<section id="bracket-preview" class="page-section">')
cat('<div class="section-heading">')
cat('<span class="section-kicker">Tournament outlook</span>')
cat('<h2>Bracket Preview</h2>')
cat('<p>The bracket starts from projected group seeds and can be explored in more detail on the Predictions page.</p>')
cat('</div>')
cat(render_bracket(root, preview = TRUE))
cat('<p><a class="button-secondary" href="reports/08_matchday_predictions.html#bracket">Open full bracket</a></p>')
cat('</section>')
```

```{r methodology, results="asis"}
cat(render_methodology_short())
```

## Limitations

This forecast is probabilistic. Lineups, injuries, card projections, and provider feeds update when the connected sources make those records available. Current tournament accuracy should be interpreted carefully because completed-match sample sizes are still small.

<p class="responsible-note">Responsible use: this site presents statistical forecasts for education and research. Always verify official match information before making decisions.</p>
