---
title: "2026 World Cup Betting Model"
subtitle: "Goals, match-result probabilities, and data diagnostics for the 2026 FIFA World Cup"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
root <- if (file.exists("data/processed/modeling")) "." else ".."
source(file.path(root, "R", "report_helpers.R"))
read_csv_if_exists <- function(path) {
  if (file.exists(path)) {
    readr::read_csv(path, show_col_types = FALSE)
  } else {
    data.frame()
  }
}
```

<div class="draft-label">DRAFT</div>

<div class="hero-actions">
  <a class="btn" href="https://github.com/LindaData/world-cup-2026-betting-model">View on GitHub</a>
  <a class="btn" href="#data-sources-and-expense-overview">Data Sources</a>
  <a class="btn" href="reports/01_goals_linear_regression.html">Goals Model</a>
  <a class="btn" href="reports/02_ordinal_result_model.html">Result Model</a>
  <a class="btn" href="reports/03_distribution_review.html">Distributions</a>
  <a class="btn" href="#dictionary">Dictionary</a>
  <a class="btn" href="https://lindadata.github.io/">LindaData Home</a>
</div>

This site presents the 2026 World Cup betting model, the data used to support it, and the model results in one shareable place.

Code is hidden by default so the report reads cleanly. Use **Show code** to inspect the R steps.

No raw datasets, processed databases, API keys, `.Renviron`, credentials, or private files are published here.

```{r hero-metrics, echo=FALSE, results="asis"}
goals_metrics <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "goals_linear_model_metrics.csv"))
ordinal_metrics <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "result_ordinal_model_metrics.csv"))

if (nrow(goals_metrics) > 0 && nrow(ordinal_metrics) > 0) {
  cat(metric_cards(data.frame(
    label = c("Rows Modeled", "Goals RMSE", "Result Accuracy", "Reference Accuracy"),
    value = c(
      fmt_integer(goals_metrics$rows_total[[1]]),
      fmt_number(goals_metrics$test_rmse[[1]], 3),
      fmt_percent(ordinal_metrics$test_accuracy[[1]], 1),
      fmt_percent(ordinal_metrics$baseline_majority_accuracy[[1]], 1)
    ),
    note = c("team-match rows", "typical goal miss", "exact result hit rate", "simple comparison")
  )))
}
```

## Data Sources and Expense Overview

::: {.api-choice}
**Primary live source: API-Football by API-Sports**

The model uses API-Football for World Cup fixtures, teams, standings, player/squad data, injuries, lineups, match events, player statistics, odds, and predictions where coverage is available.
:::

```{r data-sources, echo=FALSE}
data_sources <- data.frame(
  layer = c(
    "Live football API",
    "Historical match results",
    "Player and squad context",
    "Venue, weather, and news context",
    "Publishing"
  ),
  source = c(
    "API-Football / API-Sports",
    "Public international results and goalscorer datasets",
    "API-Football plus public squad references",
    "Open venue/weather/news sources",
    "GitHub Pages"
  ),
  use_in_model = c(
    "Fixtures, teams, standings, injuries, lineups, odds, and match details",
    "Team form, goals scored, results, Elo-style strength features",
    "Roster strength, player goal history, availability indicators",
    "Travel, venue, weather, and news-derived context features",
    "Shareable HTML reports and model summaries"
  )
)
display_table(data_sources, caption = "Data sources used by the modeling workflow.")
```

The public site is hosted without an additional hosting bill. The live football data subscription is the main recurring project expense, and the project is kept within the monthly budget cap.

## Model Iterations

The model section is organized around interpretable iterations. Each iteration names the prediction target, shows the equation, explains the coefficients in plain English, and lists the checks that determine whether the model is trustworthy.

::: {.model-blueprint}
::: {.model-card}
### 1. Simple OLS Goals Model

**Question:** how many goals should a team score in one match?

The simple goals model starts with team strength difference:

$$
\text{Goals} = \beta_0 + \beta_1\text{Strength Difference} + \epsilon
$$

In words, `beta_1` is the average change in expected goals for a one-unit increase in strength difference.

**Why start here:** OLS is easy to explain, fast to diagnose, and gives a clear benchmark before count models.

**Main issues to check:**

- Goals are counts, so OLS can predict impossible values below 0.
- Soccer scores are low and skewed, so residuals may not look normal.
- Error variance can change by team strength or opponent strength.
- Team-match rows are not fully independent because each match creates two rows.
- Missing roster, injury, rest, venue, and market data can bias coefficients.
:::

::: {.model-card}
### 2. Three-Outcome Result Model

**Question:** what is the probability of win, draw, or loss?

The first three-outcome model is an ordinal logistic regression because the team result has a natural order:

$$
\text{loss} < \text{draw} < \text{win}
$$

The model uses cumulative logits:

$$
\text{logit}\{P(\text{Result} \leq k)\}
= \theta_k - \eta
$$

with:

$$
\eta = \beta_1\text{Strength Difference}
+ \beta_2\text{Roster Strength}
+ \beta_3\text{Rest}
+ \beta_4\text{Venue}
$$

The output is three probabilities:

$$
P(\text{loss}), \quad P(\text{draw}), \quad P(\text{win})
$$

**Main issues to check:**

- The proportional-odds assumption may be too strong.
- Draws are not simply halfway between loss and win.
- A multinomial logit may fit better if win, draw, and loss behave as separate categories.
- Calibration matters more than raw accuracy for betting comparison.
- Market odds should be the benchmark once API odds are available.
:::
:::

## Dictionary

Use this as the quick reference for acronyms and modeling terms that appear across the site.

| Term | Plain-English Meaning | Why It Matters Here |
|---|---|---|
| `Y` or target | The thing the model is trying to predict. | For the goals model, `Y` is goals scored. For the result model, `Y` is win, draw, or loss. |
| `X` or predictor | An input used to make the prediction. | Team strength, opponent strength, rest, venue, injuries, and odds can all be predictors. |
| Regression | A model that estimates how a target changes with predictors. | It gives a structured way to turn soccer information into expected goals or probabilities. |
| OLS | Ordinary Least Squares. A line-fitting method that chooses coefficients by minimizing squared prediction errors. | It is the first goals benchmark because it is easy to explain and diagnose. |
| RMSE | Root Mean Squared Error. A typical prediction miss, measured in the same units as the target. | If goals RMSE is 1.46, the model is often off by roughly 1 to 2 goals, with big misses punished more. Lower is better. |
| Coefficient | The estimated effect of one predictor, holding the other predictors fixed. | A positive strength coefficient means stronger teams are expected to score more, all else equal. |
| Residual | Observed value minus predicted value. | Residual plots show whether the model is missing patterns. |
| Logit | A transformation that converts a probability into log odds. | Logistic models use logits so probabilities stay between 0 and 1. |
| Odds | Probability of an event divided by probability of it not happening. | Model probabilities can be compared with betting-market odds after converting both to the same scale. |
| Ordinal logistic regression | A logistic model for ordered outcomes. | It fits `loss < draw < win` while producing probabilities for all three outcomes. |
| Multinomial logistic regression | A logistic model for three or more categories without assuming an order. | It is the backup if draws behave like their own special category rather than the middle of loss and win. |
| Accuracy | Share of predictions where the top predicted class matches the actual result. | Useful for a quick check, but not enough for betting because probability quality matters too. |
| Calibration | Whether predicted probabilities match real frequencies. | If the model says 60% win chance, well-calibrated teams should win about 60% of those cases. |
| Log loss | A probability score that penalizes confident wrong predictions. | Better than accuracy when the model outputs win/draw/loss probabilities. Lower is better. |
| Brier score | Average squared error between predicted probabilities and actual outcomes. | Another way to judge probability quality. Lower is better. |
| Test set | Data held out from fitting and used only to evaluate performance. | It helps estimate how the model performs on games it has not already seen. |
| Overfitting | When a model learns noise in past data instead of repeatable signal. | A model can look good historically and still fail on new World Cup matches. |
| Data grain | What one row means in a dataset. | The main modeling grain is one team in one match. One match creates two team-match rows. |
| API | Application Programming Interface. A structured way to pull data from a provider. | API-Football is the primary live football data source. |

## Project Map

::: {.panel-tabset}

## About

This project builds a reproducible data and modeling workflow for the 2026 FIFA World Cup.

The workflow has four layers:

- **Data collection:** match history, Elo ratings, 2026 venues, weather, player metadata, odds, and news signals.
- **Data storage:** local files and DuckDB tables for modeling.
- **Modeling:** interpretable models in R, with Python handling source refreshes and heavier data jobs.
- **Publishing:** Quarto renders a shareable GitHub Pages site from the local analysis.

## Data Exploration

The starting modeling unit is one row per team-match. A single game creates two observations, one for each team.

```{r goals-sample}
goals_sample <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "goals_linear_model_sample_1000.csv"))

if (nrow(goals_sample) > 0) {
  goals_sample |>
    dplyr::select(
      date,
      team,
      opponent,
      tournament,
      y_goals_for,
      pre_elo,
      opponent_pre_elo,
      pre_match_expected_result
    ) |>
    dplyr::mutate(
      pre_match_expected_result = fmt_percent(pre_match_expected_result, 1),
      pre_elo = round(pre_elo, 0),
      opponent_pre_elo = round(opponent_pre_elo, 0)
    ) |>
    display_table(caption = "Sample of the team-match modeling grain.", max_rows = 8)
} else {
  display_table(empty_note("Sample rows are generated locally by R/10_fit_linear_goals_model.R."))
}
```

```{r goals-plot, fig.cap="Distribution of goals in the public sample used for the goals model."}
if (nrow(goals_sample) > 0 && requireNamespace("ggplot2", quietly = TRUE)) {
  ggplot2::ggplot(goals_sample, ggplot2::aes(x = y_goals_for)) +
    ggplot2::geom_histogram(binwidth = 1, boundary = -0.5, fill = "#157878", color = "white") +
    ggplot2::labs(
      title = "Goals scored by team-match row",
      x = "Goals scored",
      y = "Team-match rows"
    ) +
    ggplot2::theme_minimal(base_size = 12)
} else {
  plot.new()
  text(0.5, 0.5, "Goals sample is generated locally before rendering.")
}
```

## Modeling

The model suite has two core outputs:

- **Goals OLS:** predicts goals scored.
- **Ordinal logistic:** predicts `loss < draw < win`.

The reports show the model target, data grain, math, coefficients, predictions, and diagnostics.

```{r model-summary}
if (nrow(goals_metrics) > 0 && nrow(ordinal_metrics) > 0) {
  data.frame(
    model = c("Goals OLS", "Ordinal result logistic"),
    target = c("team goals", "loss < draw < win"),
    rows = fmt_integer(c(goals_metrics$rows_total[[1]], ordinal_metrics$rows_total[[1]])),
    primary_metric = c("Test RMSE", "Test accuracy"),
    value = c(
      fmt_number(goals_metrics$test_rmse[[1]], 3),
      fmt_percent(ordinal_metrics$test_accuracy[[1]], 1)
    ),
    context = c(
      paste("Mean test goals:", fmt_number(goals_metrics$mean_y_test[[1]], 3)),
      paste("Reference accuracy:", fmt_percent(ordinal_metrics$baseline_majority_accuracy[[1]], 1))
    )
  ) |>
    display_table(caption = "Current model performance.")
} else {
  display_table(empty_note("Model metrics are generated locally before rendering."))
}
```

## Data

The modeling database stores raw and transformed tables. The rendered site publishes metadata and model outputs, while credentials and private source files stay local.

```{r table-inventory}
inventory <- read_csv_if_exists(file.path(root, "data", "processed", "metadata", "table_inventory.csv"))

if (nrow(inventory) > 0) {
  inventory |>
    dplyr::arrange(table_name) |>
    dplyr::mutate(row_count = fmt_integer(row_count)) |>
    display_table(caption = "Local tables available for modeling.", max_rows = 16)
} else {
  display_table(empty_note("Table inventory is generated locally by R/07_export_metadata.R."))
}
```

## Display Standards

<div class="principles">
  <div class="principle"><strong>One grain per table</strong>Rows should mean one thing: team-match, match, player, venue, or API pull.</div>
  <div class="principle"><strong>Readable labels</strong>Keep machine-safe names in source files; convert them to readable labels only in reports.</div>
  <div class="principle"><strong>Round for display only</strong>Do not round model inputs or stored outputs. Round the rendered table so readers can scan it.</div>
  <div class="principle"><strong>Separate public and private</strong>Publish metadata and rendered results, not raw datasets, databases, keys, or credentials.</div>
</div>

:::

## Reports

::: {.columns}
::: {.column width="50%"}
::: {.project-card}
### Goals Model

OLS model for team goals scored from pre-match Elo, opponent Elo, expected result, home listing, and neutral-site status.

<a class="project-button" href="reports/01_goals_linear_regression.html">Open goals report</a>
:::
:::

::: {.column width="50%"}
::: {.project-card}
### Result Model

Ordinal logistic model for `loss < draw < win`, producing win/draw/loss probability estimates.

<a class="project-button" href="reports/02_ordinal_result_model.html">Open result report</a>
:::
:::
:::

::: {.project-card}
### Distribution Review

Distribution notes for deciding whether soccer features are counts, continuous variables, ordered outcomes, probabilities, or mixture processes.

<a class="project-button" href="reports/03_distribution_review.html">Open distribution review</a>
:::
