---
title: "2026 World Cup Betting Model"
subtitle: "An RStudio-first modeling lab for public soccer data, explainable baselines, and future betting-market comparison"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
root <- if (file.exists("data/processed/modeling")) "." else ".."
read_csv_if_exists <- function(path) {
  if (file.exists(path)) {
    readr::read_csv(path, show_col_types = FALSE)
  } else {
    data.frame()
  }
}
```

<div class="hero-actions">
  <a class="btn" href="https://github.com/LindaData/world-cup-2026-betting-model">View on GitHub</a>
  <a class="btn" href="reports/01_goals_linear_regression.html">Goals Model</a>
  <a class="btn" href="reports/02_ordinal_result_model.html">Result Model</a>
  <a class="btn" href="https://lindadata.github.io/">LindaData Home</a>
</div>

This site is the public-facing project notebook. It follows the same teaching-report style as the earlier COVID analysis: explain the data pull, show the transformed data, write the model math, then render plots and results.

Code is hidden by default so the report reads cleanly. Use **Show code** when you want to inspect the R steps.

No raw datasets, processed databases, API keys, `.Renviron`, credentials, or private files are published here.

## Project Map

::: {.panel-tabset}

## About

This project builds a reproducible data and modeling workflow for the 2026 FIFA World Cup. The current public version is a transparent baseline, not a final betting system.

The project is designed around four layers:

- **Data collection:** public match history, Elo ratings, 2026 venues, weather, player metadata, and news signals.
- **Data storage:** local files and DuckDB tables for modeling.
- **Modeling:** interpretable first-pass models in R, with Python handling API pulls and heavier data jobs.
- **Publishing:** Quarto renders a shareable GitHub Pages site from the local analysis.

## Data Exploration

The starting modeling unit is one row per team-match. A single game creates two observations, one for each team.

```{r goals-sample}
goals_sample <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "goals_linear_model_sample_1000.csv"))

if (nrow(goals_sample) > 0) {
  head(goals_sample, 10)
} else {
  data.frame(note = "Sample rows are generated locally by R/10_fit_linear_goals_model.R.")
}
```

```{r goals-plot, fig.cap="Distribution of goals in the public sample used for the baseline model."}
if (nrow(goals_sample) > 0 && requireNamespace("ggplot2", quietly = TRUE)) {
  ggplot2::ggplot(goals_sample, ggplot2::aes(x = y_goals_for)) +
    ggplot2::geom_histogram(binwidth = 1, boundary = -0.5, fill = "#157878", color = "white") +
    ggplot2::labs(
      title = "Goals scored by team-match row",
      x = "Goals scored",
      y = "Team-match rows"
    ) +
    ggplot2::theme_minimal(base_size = 12)
} else {
  plot.new()
  text(0.5, 0.5, "Goals sample is generated locally before rendering.")
}
```

## Modeling

The first two models are intentionally simple:

- **Goals OLS:** predicts goals scored.
- **Ordinal logistic:** predicts `loss < draw < win`.

This mirrors the Shiny layout you used before: start with descriptive analytics, then model fitting, then prediction outputs.

```{r model-summary}
goals_metrics <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "goals_linear_model_metrics.csv"))
ordinal_metrics <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "result_ordinal_model_metrics.csv"))

if (nrow(goals_metrics) > 0 && nrow(ordinal_metrics) > 0) {
  data.frame(
    model = c("Goals OLS", "Ordinal result logistic"),
    target = c("team goals", "loss < draw < win"),
    rows_total = c(goals_metrics$rows_total[[1]], ordinal_metrics$rows_total[[1]]),
    primary_metric = c("test RMSE", "test accuracy"),
    value = c(
      round(goals_metrics$test_rmse[[1]], 4),
      round(ordinal_metrics$test_accuracy[[1]], 4)
    ),
    baseline_context = c(
      paste("mean test goals:", round(goals_metrics$mean_y_test[[1]], 4)),
      paste("majority baseline:", round(ordinal_metrics$baseline_majority_accuracy[[1]], 4))
    )
  )
} else {
  data.frame(note = "Model metrics are generated locally before rendering.")
}
```

## Data

The local database contains more tables than the public site publishes. The rendered site shows metadata only, which keeps credentials and source data off GitHub.

```{r table-inventory}
inventory <- read_csv_if_exists(file.path(root, "data", "processed", "metadata", "table_inventory.csv"))

if (nrow(inventory) > 0) {
  inventory |>
    dplyr::arrange(table_name) |>
    head(20)
} else {
  data.frame(note = "Table inventory is generated locally by R/07_export_metadata.R.")
}
```

:::

## Current Models

<div class="metric-strip">
  <div class="metric"><strong>Primary target</strong><span>Goals</span></div>
  <div class="metric"><strong>Result target</strong><span>W / D / L</span></div>
  <div class="metric"><strong>Workflow</strong><span>R + Python + SQL</span></div>
  <div class="metric"><strong>Publishing</strong><span>Quarto HTML</span></div>
</div>

## Reports

::: {.columns}
::: {.column width="50%"}
::: {.project-card}
### Goals Model

OLS baseline for team goals scored from pre-match Elo, opponent Elo, expected result, home listing, and neutral-site status.

<a class="project-button" href="reports/01_goals_linear_regression.html">Open goals report</a>
:::
:::

::: {.column width="50%"}
::: {.project-card}
### Result Model

Ordinal logistic baseline for `loss < draw < win`, producing win/draw/loss probability estimates.

<a class="project-button" href="reports/02_ordinal_result_model.html">Open result report</a>
:::
:::
:::

## Interactive App Direction

::: {.shiny-panel}
### Shiny First, Streamlit Later If Needed

Your older Shiny code used `navbarPage()`, `sidebarLayout()`, `DT` tables, model controls, and MathJax. That is the right pattern for the local interactive version because you already work in RStudio.

GitHub Pages can host this static Quarto report for free. A true Shiny app needs a running R server, usually ShinyApps.io, Posit Connect, or your own cloud VM. Streamlit would have the same hosting issue, so it is not better for the public static site yet.
:::

## Refresh Options

Run these from RStudio.

### Local Rebuild

Does not hit APIs or spend quota.

```{r local-refresh, eval=FALSE}
source("R/20_refresh_all.R")
refresh_world_cup_data(profile = "local-rebuild")
```

### Free Source Refresh

Pulls free/no-key public sources and rebuilds the local site.

```{r free-refresh, eval=FALSE}
source("R/20_refresh_all.R")
refresh_world_cup_data(profile = "free-refresh")
```

### Keyed APIs Later

Use after `.env` is configured locally.

```{r keyed-refresh, eval=FALSE}
source("R/20_refresh_all.R")
refresh_world_cup_data(
  profile = "free-refresh",
  include_keyed_apis = TRUE
)
```

Odds pulls require explicit quota permission.

```{r odds-refresh, eval=FALSE}
source("R/20_refresh_all.R")
refresh_world_cup_data(
  profile = "free-refresh",
  include_keyed_apis = TRUE,
  include_odds_quota = TRUE
)
```

## Next Work

1. Add rolling recent-form features.
2. Add squad-strength and player-availability signals.
3. Add news-derived injury and lineup signals.
4. Add odds-implied probabilities.
5. Compare ordinal logistic, multinomial logistic, and paired-goal models.
6. Calibrate predicted probabilities.
