---
title: "2026 World Cup Betting Model"
subtitle: "Reproducible data, models, and reports for a World Cup prediction workflow"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
root <- if (file.exists("data/processed/modeling")) "." else ".."
read_csv_if_exists <- function(path) {
  if (file.exists(path)) {
    readr::read_csv(path, show_col_types = FALSE)
  } else {
    data.frame()
  }
}
```

This site is the public-facing project notebook. It keeps the rendered analysis in HTML, hides code by default, and lets readers expand code with **Show code**.

No raw datasets, processed databases, API keys, `.Renviron`, credentials, or private files are published here.

## Current Models

```{r model-summary}
goals_metrics <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "goals_linear_model_metrics.csv"))
ordinal_metrics <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "result_ordinal_model_metrics.csv"))

if (nrow(goals_metrics) > 0 && nrow(ordinal_metrics) > 0) {
  data.frame(
    model = c("Goals OLS", "Ordinal result logistic"),
    target = c("team goals", "loss < draw < win"),
    rows_total = c(goals_metrics$rows_total[[1]], ordinal_metrics$rows_total[[1]]),
    primary_metric = c("test RMSE", "test accuracy"),
    value = c(
      round(goals_metrics$test_rmse[[1]], 4),
      round(ordinal_metrics$test_accuracy[[1]], 4)
    ),
    baseline_context = c(
      paste("mean test goals:", round(goals_metrics$mean_y_test[[1]], 4)),
      paste("majority baseline:", round(ordinal_metrics$baseline_majority_accuracy[[1]], 4))
    )
  )
} else {
  data.frame(note = "Model metrics are generated locally before rendering.")
}
```

## Reports

::: {.columns}
::: {.column width="50%"}
::: {.project-card}
### Goals Model

OLS baseline for team goals scored from pre-match Elo, opponent Elo, expected result, home listing, and neutral-site status.

[Open goals report](reports/01_goals_linear_regression.html)
:::
:::

::: {.column width="50%"}
::: {.project-card}
### Result Model

Ordinal logistic baseline for `loss < draw < win`, producing win/draw/loss probability estimates.

[Open result report](reports/02_ordinal_result_model.html)
:::
:::
:::

## Data Status

```{r table-inventory}
inventory <- read_csv_if_exists(file.path(root, "data", "processed", "metadata", "table_inventory.csv"))

if (nrow(inventory) > 0) {
  inventory |>
    dplyr::arrange(table_name) |>
    head(20)
} else {
  data.frame(note = "Table inventory is generated locally by R/07_export_metadata.R.")
}
```

## Refresh Options

Run these from RStudio.

### Local Rebuild

Does not hit APIs or spend quota.

```{r local-refresh, eval=FALSE}
source("R/20_refresh_all.R")
refresh_world_cup_data(profile = "local-rebuild")
```

### Free Source Refresh

Pulls free/no-key public sources and rebuilds the local site.

```{r free-refresh, eval=FALSE}
source("R/20_refresh_all.R")
refresh_world_cup_data(profile = "free-refresh")
```

### Keyed APIs Later

Use after `.env` is configured locally.

```{r keyed-refresh, eval=FALSE}
source("R/20_refresh_all.R")
refresh_world_cup_data(
  profile = "free-refresh",
  include_keyed_apis = TRUE
)
```

Odds pulls require explicit quota permission.

```{r odds-refresh, eval=FALSE}
source("R/20_refresh_all.R")
refresh_world_cup_data(
  profile = "free-refresh",
  include_keyed_apis = TRUE,
  include_odds_quota = TRUE
)
```

## Next Work

1. Add rolling recent-form features.
2. Add squad-strength and player-availability signals.
3. Add news-derived injury and lineup signals.
4. Add odds-implied probabilities.
5. Compare ordinal logistic, multinomial logistic, and paired-goal models.
6. Calibrate predicted probabilities.
