---
title: "2026 World Cup Betting Model"
subtitle: "Goals, match-result probabilities, and data diagnostics for the 2026 FIFA World Cup"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
root <- if (file.exists("data/processed/modeling")) "." else ".."
source(file.path(root, "R", "report_helpers.R"))
read_csv_if_exists <- function(path) {
  if (file.exists(path)) {
    readr::read_csv(path, show_col_types = FALSE)
  } else {
    data.frame()
  }
}
```

<div class="draft-label">DRAFT</div>

<div class="hero-actions">
  <a class="btn" href="https://github.com/LindaData/world-cup-2026-betting-model">View on GitHub</a>
  <a class="btn" href="reports/01_goals_linear_regression.html">Goals Model</a>
  <a class="btn" href="reports/02_ordinal_result_model.html">Result Model</a>
  <a class="btn" href="reports/03_distribution_review.html">Distributions</a>
  <a class="btn" href="https://lindadata.github.io/">LindaData Home</a>
</div>

This dashboard organizes the 2026 World Cup betting model, model diagnostics, and data-source decision in one shareable site.

Code is hidden by default so the report reads cleanly. Use **Show code** to inspect the R steps.

No raw datasets, processed databases, API keys, `.Renviron`, credentials, or private files are published here.

```{r hero-metrics, echo=FALSE, results="asis"}
goals_metrics <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "goals_linear_model_metrics.csv"))
ordinal_metrics <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "result_ordinal_model_metrics.csv"))

if (nrow(goals_metrics) > 0 && nrow(ordinal_metrics) > 0) {
  cat(metric_cards(data.frame(
    label = c("Rows Modeled", "Goals RMSE", "Result Accuracy", "Reference Accuracy"),
    value = c(
      fmt_integer(goals_metrics$rows_total[[1]]),
      fmt_number(goals_metrics$test_rmse[[1]], 3),
      fmt_percent(ordinal_metrics$test_accuracy[[1]], 1),
      fmt_percent(ordinal_metrics$baseline_majority_accuracy[[1]], 1)
    ),
    note = c("team-match rows", "lower is better", "test set", "reference class")
  )))
}
```

## Data API Decision

::: {.api-choice}
**Recommended API: API-Football Pro**

Use one paid API as the primary data connection: API-Football Pro at **$19/month**. Current spend is **$10/month**, so the projected monthly total is **$29/month**, well under the $100 cap.
:::

```{r api-decision, echo=FALSE}
api_options <- data.frame(
  provider = c("API-Football Pro", "The Odds API 20K", "football-data.org ML / Deep Data", "Sportmonks Starter"),
  price = c("$19/mo", "$30/mo", "EUR29/mo before add-ons", "EUR29/mo before add-ons"),
  fit = c(
    "Best one-API fit: fixtures, teams, players/coaches, lineups, injuries, statistics, predictions, and pre-match odds.",
    "Excellent odds feed, but not enough team/player model data by itself.",
    "Good football data, but odds and statistics are add-ons and coverage is narrower for this use case.",
    "Strong football data, but useful betting/advanced features add monthly cost."
  ),
  decision = c("Recommended", "Use for odds validation", "Defer", "Defer")
)
display_table(api_options, caption = "Paid API options reviewed for the primary data connection.")
```

Sources: [API-Football pricing](https://www.api-football.com/pricing), [The Odds API pricing](https://the-odds-api.com/), [football-data.org pricing](https://www.football-data.org/pricing), and [Sportmonks pricing](https://www.sportmonks.com/football-api/).

After subscribing, add the API key only to local `.env` as `API_FOOTBALL_KEY`. Do not commit the key.

## Project Map

::: {.panel-tabset}

## About

This project builds a reproducible data and modeling workflow for the 2026 FIFA World Cup.

The workflow has four layers:

- **Data collection:** match history, Elo ratings, 2026 venues, weather, player metadata, odds, and news signals.
- **Data storage:** local files and DuckDB tables for modeling.
- **Modeling:** interpretable models in R, with Python handling API pulls and heavier data jobs.
- **Publishing:** Quarto renders a shareable GitHub Pages site from the local analysis.

## Data Exploration

The starting modeling unit is one row per team-match. A single game creates two observations, one for each team.

```{r goals-sample}
goals_sample <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "goals_linear_model_sample_1000.csv"))

if (nrow(goals_sample) > 0) {
  goals_sample |>
    dplyr::select(
      date,
      team,
      opponent,
      tournament,
      y_goals_for,
      pre_elo,
      opponent_pre_elo,
      pre_match_expected_result
    ) |>
    dplyr::mutate(
      pre_match_expected_result = fmt_percent(pre_match_expected_result, 1),
      pre_elo = round(pre_elo, 0),
      opponent_pre_elo = round(opponent_pre_elo, 0)
    ) |>
    display_table(caption = "Sample of the team-match modeling grain.", max_rows = 8)
} else {
  display_table(empty_note("Sample rows are generated locally by R/10_fit_linear_goals_model.R."))
}
```

```{r goals-plot, fig.cap="Distribution of goals in the public sample used for the goals model."}
if (nrow(goals_sample) > 0 && requireNamespace("ggplot2", quietly = TRUE)) {
  ggplot2::ggplot(goals_sample, ggplot2::aes(x = y_goals_for)) +
    ggplot2::geom_histogram(binwidth = 1, boundary = -0.5, fill = "#157878", color = "white") +
    ggplot2::labs(
      title = "Goals scored by team-match row",
      x = "Goals scored",
      y = "Team-match rows"
    ) +
    ggplot2::theme_minimal(base_size = 12)
} else {
  plot.new()
  text(0.5, 0.5, "Goals sample is generated locally before rendering.")
}
```

## Modeling

The model suite has two core outputs:

- **Goals OLS:** predicts goals scored.
- **Ordinal logistic:** predicts `loss < draw < win`.

The reports show the model target, data grain, math, coefficients, predictions, and diagnostics.

```{r model-summary}
if (nrow(goals_metrics) > 0 && nrow(ordinal_metrics) > 0) {
  data.frame(
    model = c("Goals OLS", "Ordinal result logistic"),
    target = c("team goals", "loss < draw < win"),
    rows = fmt_integer(c(goals_metrics$rows_total[[1]], ordinal_metrics$rows_total[[1]])),
    primary_metric = c("Test RMSE", "Test accuracy"),
    value = c(
      fmt_number(goals_metrics$test_rmse[[1]], 3),
      fmt_percent(ordinal_metrics$test_accuracy[[1]], 1)
    ),
    context = c(
      paste("Mean test goals:", fmt_number(goals_metrics$mean_y_test[[1]], 3)),
      paste("Reference accuracy:", fmt_percent(ordinal_metrics$baseline_majority_accuracy[[1]], 1))
    )
  ) |>
    display_table(caption = "Current model performance.")
} else {
  display_table(empty_note("Model metrics are generated locally before rendering."))
}
```

## Data

The local database contains more tables than the public site publishes. The rendered site shows metadata only, which keeps credentials and source data off GitHub.

```{r table-inventory}
inventory <- read_csv_if_exists(file.path(root, "data", "processed", "metadata", "table_inventory.csv"))

if (nrow(inventory) > 0) {
  inventory |>
    dplyr::arrange(table_name) |>
    dplyr::mutate(row_count = fmt_integer(row_count)) |>
    display_table(caption = "Local tables available for modeling.", max_rows = 16)
} else {
  display_table(empty_note("Table inventory is generated locally by R/07_export_metadata.R."))
}
```

## Display Standards

<div class="principles">
  <div class="principle"><strong>One grain per table</strong>Rows should mean one thing: team-match, match, player, venue, or API pull.</div>
  <div class="principle"><strong>Readable labels</strong>Keep machine-safe names in source files; convert them to readable labels only in reports.</div>
  <div class="principle"><strong>Round for display only</strong>Do not round model inputs or stored outputs. Round the rendered table so readers can scan it.</div>
  <div class="principle"><strong>Separate public and private</strong>Publish metadata and rendered results, not raw datasets, databases, keys, or credentials.</div>
</div>

:::

## Reports

::: {.columns}
::: {.column width="50%"}
::: {.project-card}
### Goals Model

OLS model for team goals scored from pre-match Elo, opponent Elo, expected result, home listing, and neutral-site status.

<a class="project-button" href="reports/01_goals_linear_regression.html">Open goals report</a>
:::
:::

::: {.column width="50%"}
::: {.project-card}
### Result Model

Ordinal logistic model for `loss < draw < win`, producing win/draw/loss probability estimates.

<a class="project-button" href="reports/02_ordinal_result_model.html">Open result report</a>
:::
:::
:::

::: {.project-card}
### Distribution Review

Distribution notes for deciding whether soccer features are counts, continuous variables, ordered outcomes, probabilities, or mixture processes.

<a class="project-button" href="reports/03_distribution_review.html">Open distribution review</a>
:::

## Refresh Commands

Run these from RStudio.

### Local Rebuild

Does not hit APIs or spend quota.

```{r local-refresh, eval=FALSE}
source("R/20_refresh_all.R")
refresh_world_cup_data(profile = "local-rebuild")
```

### Free Source Refresh

Pulls free/no-key public sources and rebuilds the local site.

```{r free-refresh, eval=FALSE}
source("R/20_refresh_all.R")
refresh_world_cup_data(profile = "free-refresh")
```

### API-Football Refresh

Use after `.env` has `API_FOOTBALL_KEY`.

```{r keyed-refresh, eval=FALSE}
source("R/20_refresh_all.R")
refresh_world_cup_data(
  profile = "free-refresh",
  include_keyed_apis = TRUE
)
```
