---
title: "Data & Coverage"
subtitle: "Available data, sample modeling rows, metadata, and simple diagnostics"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
root <- if (file.exists("data/processed/modeling")) "." else ".."
source(file.path(root, "R", "report_helpers.R"))
read_csv_if_exists <- function(path) {
  if (file.exists(path)) {
    readr::read_csv(path, show_col_types = FALSE)
  } else {
    data.frame()
  }
}
```

## What Data Is Available

The model starts with free and public data, then uses connected football API coverage as a structured enrichment layer. The regression baseline is built from open historical match, player, venue, weather, and news metadata.

```{r available-data}
available_data <- data.frame(
  data_area = c(
    "International match history",
    "Team strength history",
    "Player goals",
    "2026 fixtures and venues",
    "Squad/player references",
    "Weather signals",
    "News metadata",
    "Live football API layer"
  ),
  current_status = c(
    "Available",
    "Available",
    "Available",
    "Available",
    "Partially available",
    "Available for scheduled venues",
    "Available as article metadata",
    "Connected as a coverage and enrichment layer"
  ),
  modeling_use = c(
    "Results, goals for/against, tournament context, neutral-site flag",
    "Pre-match Elo-style strength, opponent strength, expected result",
    "Historical scorer context and player goal totals",
    "Future match shell, host cities, venues, travel context",
    "Roster context and player identity enrichment where public references match",
    "Temperature, wind, humidity, and match-window conditions",
    "Team and tournament news volume/context features",
    "Fixtures, teams, standings, lineups, injuries, odds, events, and player stats"
  ),
  stringsAsFactors = FALSE
)

display_table(available_data, caption = "Available data areas and modeling use.")
```

## Sample Modeling Dataset

The core modeling grain is **one team in one match**. A single match appears twice: once from each team's perspective.

```{r sample-modeling-data}
sample_data <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "goals_linear_model_sample_1000.csv"))

if (nrow(sample_data) > 0) {
  sample_data |>
    dplyr::select(
      date,
      team,
      opponent,
      tournament,
      y_goals_for,
      pre_elo,
      opponent_pre_elo,
      pre_match_expected_result,
      listed_home,
      neutral
    ) |>
    dplyr::mutate(
      pre_elo = round(pre_elo, 0),
      opponent_pre_elo = round(opponent_pre_elo, 0),
      pre_match_expected_result = fmt_percent(pre_match_expected_result, 1)
    ) |>
    display_table(caption = "Sample team-match rows used by the free-data model baseline.", max_rows = 12)
} else {
  display_table(empty_note("Sample modeling rows are available after the goals model refresh."))
}
```

## Metadata

The table inventory shows what is available in the local modeling database. The rendered site embeds metadata summaries, not raw data files.

```{r table-inventory}
inventory <- read_csv_if_exists(file.path(root, "data", "processed", "metadata", "table_inventory.csv"))

if (nrow(inventory) > 0) {
  inventory |>
    dplyr::arrange(dplyr::desc(row_count)) |>
    dplyr::mutate(row_count = fmt_integer(row_count)) |>
    display_table(caption = "Modeling table inventory.", max_rows = 20)
} else {
  display_table(empty_note("Table inventory is available after the metadata refresh."))
}
```

```{r column-inventory}
columns <- read_csv_if_exists(file.path(root, "data", "processed", "metadata", "column_inventory.csv"))

if (nrow(columns) > 0) {
  type_column <- if ("data_type" %in% names(columns)) "data_type" else "column_type"
  columns |>
    dplyr::filter(table_name %in% c(
      "vw_goals_linear_model_frame",
      "vw_result_ordinal_model_frame",
      "vw_2026_fixture_model_frame",
      "vw_2026_team_model_features"
    )) |>
    dplyr::select(table_name, column_name, data_type = dplyr::all_of(type_column)) |>
    display_table(caption = "Selected metadata for model-ready tables.", max_rows = 40)
} else {
  display_table(empty_note("Column inventory is available after the metadata refresh."))
}
```

## Simple Visuals

```{r row-counts-plot, fig.cap="Largest modeling tables by row count."}
if (nrow(inventory) > 0 && requireNamespace("ggplot2", quietly = TRUE)) {
  inventory |>
    dplyr::arrange(dplyr::desc(row_count)) |>
    dplyr::slice_head(n = 12) |>
    ggplot2::ggplot(ggplot2::aes(x = stats::reorder(table_name, row_count), y = row_count)) +
    ggplot2::geom_col(fill = "#0f6b68") +
    ggplot2::coord_flip() +
    ggplot2::scale_y_continuous(labels = function(x) format(x, big.mark = ",", scientific = FALSE, trim = TRUE)) +
    ggplot2::labs(
      title = "Largest tables in the modeling database",
      x = NULL,
      y = "Rows"
    ) +
    ggplot2::theme_minimal(base_size = 12)
} else {
  plot.new()
  text(0.5, 0.5, "Row-count chart is available after the metadata refresh.")
}
```

```{r goals-plot, fig.cap="Goals scored per team-match row in the model sample."}
if (nrow(sample_data) > 0 && requireNamespace("ggplot2", quietly = TRUE)) {
  ggplot2::ggplot(sample_data, ggplot2::aes(x = y_goals_for)) +
    ggplot2::geom_histogram(binwidth = 1, boundary = -0.5, fill = "#d9a441", color = "white") +
    ggplot2::labs(
      title = "Goals scored by team-match row",
      x = "Goals scored",
      y = "Rows"
    ) +
    ggplot2::theme_minimal(base_size = 12)
} else {
  plot.new()
  text(0.5, 0.5, "Goals chart is available after the model refresh.")
}
```

```{r result-distribution, fig.cap="Observed win/draw/loss distribution from historical team-match rows."}
result_frame <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "result_ordinal_model_frame.csv"))

if (nrow(result_frame) > 0 && requireNamespace("ggplot2", quietly = TRUE)) {
  result_frame |>
    dplyr::count(y_result_ordered, name = "rows") |>
    dplyr::mutate(y_result_ordered = factor(y_result_ordered, levels = c("loss", "draw", "win"))) |>
    ggplot2::ggplot(ggplot2::aes(x = y_result_ordered, y = rows, fill = y_result_ordered)) +
    ggplot2::geom_col(show.legend = FALSE) +
    ggplot2::scale_y_continuous(labels = function(x) format(x, big.mark = ",", scientific = FALSE, trim = TRUE)) +
    ggplot2::scale_fill_manual(values = c("#b85c5c", "#d9a441", "#0f6b68")) +
    ggplot2::labs(
      title = "Historical result distribution",
      x = "Result from team perspective",
      y = "Rows"
    ) +
    ggplot2::theme_minimal(base_size = 12)
} else {
  plot.new()
  text(0.5, 0.5, "Result distribution is available after the result model refresh.")
}
```

## Coverage Notes

Late-breaking match information is treated as structured enrichment: confirmed starting lineups, injury status, suspensions, player minutes, event-level player statistics, and betting market movement can be added as model inputs as those fields become available.
