---
title: "2026 World Cup Forecasting Model: Project Overview"
author: "World Cup Forecasting Model Project"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)
```

# Purpose

This project builds a reproducible modeling workflow for the 2026 FIFA World Cup.

The current version is intentionally transparent:

- Python pulls raw data and API snapshots.
- SQL/DuckDB stores model-ready tables and views.
- R fits statistical models and knits presentable reports.
- The `docs/` folder holds the local website that will later publish to GitHub Pages.

# Current Modeling Stack

There are two baseline models right now.

| Model | Target | Report |
| --- | --- | --- |
| Goals OLS | Team goals scored in a match | `reports/01_goals_linear_regression.html` |
| Ordinal logistic | `loss < draw < win` | `reports/02_ordinal_result_model.html` |

These are baselines, not final forecasting models. They give us a working pipeline, interpretable coefficients, sample data, and repeatable metrics.

# Latest Metrics

```{r latest-metrics}
goals_metrics_path <- file.path("..", "data", "processed", "modeling", "goals_linear_model_metrics.csv")
ordinal_metrics_path <- file.path("..", "data", "processed", "modeling", "result_ordinal_model_metrics.csv")

if (!file.exists(goals_metrics_path)) {
  goals_metrics_path <- file.path("data", "processed", "modeling", "goals_linear_model_metrics.csv")
  ordinal_metrics_path <- file.path("data", "processed", "modeling", "result_ordinal_model_metrics.csv")
}

goals_metrics <- readr::read_csv(goals_metrics_path, show_col_types = FALSE)
ordinal_metrics <- readr::read_csv(ordinal_metrics_path, show_col_types = FALSE)

model_summary <- data.frame(
  model = c("Goals OLS", "Ordinal result logistic"),
  rows_total = c(goals_metrics$rows_total[[1]], ordinal_metrics$rows_total[[1]]),
  primary_metric = c("test_rmse", "test_accuracy"),
  metric_value = c(goals_metrics$test_rmse[[1]], ordinal_metrics$test_accuracy[[1]]),
  comparison = c(
    paste("mean_y_test =", round(goals_metrics$mean_y_test[[1]], 4)),
    paste("majority_baseline =", round(ordinal_metrics$baseline_majority_accuracy[[1]], 4))
  )
)

model_summary
```

# Data Status

The public/no-key data currently includes:

- Historical international match results.
- Historical player goal events.
- Computed team Elo match history.
- 2026 World Cup fixtures.
- 2026 squad/player tables from public sources.
- Wikidata player enrichment where available.
- Open-Meteo weather rows for mapped venues.
- GDELT news metadata test pull.

Missing or incomplete inputs for advanced betting work:

- Confirmed lineups.
- Player minutes and substitutions.
- Injuries and suspensions.
- Player-match statistics.
- Odds snapshots and closing-line movement.
- xG, shots, cards, and referee features.

# Refresh Options

Use these from RStudio.

## Rebuild From Local Data

This is the safest refresh. It does not hit APIs or spend quota.

```{r local-refresh, eval=FALSE}
source("R/20_refresh_all.R")
refresh_world_cup_data(profile = "local-rebuild")
```

## Pull Free No-Key Sources

This refreshes public files, Wikimedia/Wikidata, Open-Meteo, and GDELT.

```{r free-refresh, eval=FALSE}
source("R/20_refresh_all.R")
refresh_world_cup_data(profile = "free-refresh")
```

## Include Keyed APIs Later

Use this only after API keys are added to `.env`.

```{r keyed-refresh, eval=FALSE}
source("R/20_refresh_all.R")
refresh_world_cup_data(
  profile = "free-refresh",
  include_keyed_apis = TRUE
)
```

Odds pulls require a separate explicit option:

```{r odds-refresh, eval=FALSE}
source("R/20_refresh_all.R")
refresh_world_cup_data(
  profile = "free-refresh",
  include_keyed_apis = TRUE,
  include_odds_quota = TRUE
)
```

# RStudio Workflow

Open:

```text
world-cup-betting-model.Rproj
```

Then knit all reports:

```{r render-reports, eval=FALSE}
source("R/12_render_reports.R")
```

The local website starts at:

```text
docs/index.html
```

# Next Modeling Work

The next useful model improvements are:

1. Rolling team form before each match.
2. Squad-strength summaries from player caps, goals, age, position, and club.
3. News-derived injury and lineup signals.
4. Odds-implied probabilities.
5. Multinomial logistic regression for win/draw/loss.
6. Poisson or negative-binomial paired-goal models.
7. Probability calibration.
