---
title: "Goals Model"
subtitle: "Team goals scored from strength, opponent strength, and match context"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
root <- if (file.exists("data/processed/modeling")) "." else ".."
source(file.path(root, "R", "report_helpers.R"))
read_csv_if_exists <- function(path) {
  if (file.exists(path)) {
    readr::read_csv(path, show_col_types = FALSE)
  } else {
    data.frame()
  }
}
```

## Purpose

This report documents the goals model for the World Cup prediction project.

The target is team goals scored in a historical international match:

$$
Y_{i,m} = \text{goals scored by team } i \text{ in match } m
$$

Goals are non-negative counts. The distribution report compares count-model shapes used for soccer scoring.

## Unit Of Observation

The modeling frame has one row per **team-match**.

If Argentina plays France, the data has two rows:

- Argentina's goals as the target, with France as the opponent.
- France's goals as the target, with Argentina as the opponent.

## Sample Rows Embedded In Report

No sample CSV is published with the website. These rows are embedded into the rendered HTML from the local model frame.

```{r sample-rows}
sample_data <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "goals_linear_model_sample_1000.csv"))

if (nrow(sample_data) > 0) {
  sample_data |>
    dplyr::select(
      date,
      team,
      opponent,
      tournament,
      y_goals_for,
      pre_elo,
      opponent_pre_elo,
      pre_match_expected_result
    ) |>
    dplyr::mutate(
      pre_elo = round(pre_elo, 0),
      opponent_pre_elo = round(opponent_pre_elo, 0),
      pre_match_expected_result = fmt_percent(pre_match_expected_result, 1)
    ) |>
    display_table(caption = "Team-match sample used by the goals model.", max_rows = 10)
} else {
  display_table(empty_note("Sample rows are generated locally by R/10_fit_linear_goals_model.R."))
}
```

## Elo Feature

For one team against one opponent, the expected result before kickoff is:

$$
p =
\frac{1}{1 + 10^{-\left(\text{Team Elo} + \text{Home Adv.} - \text{Opponent Elo}\right)/400}}
$$

Where:

- `Team Elo` is the team's pre-match rating.
- `Opponent Elo` is the opponent's pre-match rating.
- `Home Adv.` is the home advantage adjustment.
- `p` is the expected result on a 0 to 1 scale.

## Model

The regression is:

$$
\text{Goals}
= \beta_0
+ \beta_1\text{Team Elo}
+ \beta_2\text{Opponent Elo}
+ \beta_3\text{Expected Result}
+ \beta_4\text{Listed Home}
+ \beta_5\text{Neutral Site}
+ \epsilon
$$

with:

$$
\epsilon_{i,m} \sim N(0, \sigma^2)
$$

## Results

```{r metrics}
metrics <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "goals_linear_model_metrics.csv"))
coefficients <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "goals_linear_model_coefficients.csv"))

if (nrow(metrics) > 0) {
  data.frame(
    rows_total = fmt_integer(metrics$rows_total),
    rows_train = fmt_integer(metrics$rows_train),
    rows_test = fmt_integer(metrics$rows_test),
    train_r_squared = fmt_percent(metrics$train_r_squared, 1),
    test_rmse = fmt_number(metrics$test_rmse, 3),
    test_mae = fmt_number(metrics$test_mae, 3),
    mean_y_test = fmt_number(metrics$mean_y_test, 3)
  ) |>
    display_table(caption = "Goals model fit summary.")
} else {
  display_table(empty_note("Metrics are generated locally."))
}
```

```{r coefficients}
if (nrow(coefficients) > 0) {
  coefficients |>
    dplyr::mutate(
      estimate = fmt_number(estimate, 4),
      std_error = fmt_number(std_error, 4),
      t_value = fmt_number(t_value, 2),
      p_value = fmt_p_value(p_value)
    ) |>
    display_table(caption = "Regression coefficients. Positive estimates increase expected goals.")
} else {
  display_table(empty_note("Coefficients are generated locally."))
}
```

```{r coefficient-plot, fig.cap="Direction and size of estimated goals-model effects."}
if (nrow(coefficients) > 0 && requireNamespace("ggplot2", quietly = TRUE)) {
  coefficients |>
    dplyr::filter(term != "(Intercept)") |>
    ggplot2::ggplot(ggplot2::aes(x = stats::reorder(term, estimate), y = estimate, fill = estimate > 0)) +
    ggplot2::geom_col(width = 0.65, show.legend = FALSE) +
    ggplot2::coord_flip() +
    ggplot2::scale_fill_manual(values = c("#b85c5c", "#157878")) +
    ggplot2::labs(
      title = "Goals model coefficients",
      x = NULL,
      y = "Estimated effect"
    ) +
    ggplot2::theme_minimal(base_size = 12)
} else {
  plot.new()
  text(0.5, 0.5, "Coefficient plot is generated locally.")
}
```

## Interpretation

The signs match soccer intuition:

- `pre_elo` is positive: stronger teams score more.
- `opponent_pre_elo` is negative: stronger opponents suppress scoring.
- `pre_match_expected_result` is positive: teams expected to perform better score more.
- `listed_homeTRUE` is positive: listed home teams score more.

## Reproduce Locally

```{r reproduce, eval=FALSE}
source("R/01_build_duckdb.R")
source("R/10_fit_linear_goals_model.R")
```
