---
title: "Goals Linear Regression"
subtitle: "Baseline model for team goals scored"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
root <- if (file.exists("data/processed/modeling")) "." else ".."
read_csv_if_exists <- function(path) {
  if (file.exists(path)) {
    readr::read_csv(path, show_col_types = FALSE)
  } else {
    data.frame()
  }
}
```

## Purpose

This report documents the first transparent baseline model for the World Cup prediction project.

The target is team goals scored in a historical international match:

$$
Y_{i,m} = \text{goals scored by team } i \text{ in match } m
$$

This is a baseline, not the final soccer model. Goals are non-negative counts, so future versions should compare Poisson, negative-binomial, bivariate Poisson, or Skellam-style models.

## Unit Of Observation

The modeling frame has one row per **team-match**.

If Argentina plays France, the data has two rows:

- Argentina's goals as the target, with France as the opponent.
- France's goals as the target, with Argentina as the opponent.

## Sample Rows Embedded In Report

No sample CSV is published with the website. These rows are embedded into the rendered HTML from the local model frame.

```{r sample-rows}
sample_data <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "goals_linear_model_sample_1000.csv"))

if (nrow(sample_data) > 0) {
  head(sample_data, 10)
} else {
  data.frame(note = "Sample rows are generated locally by R/10_fit_linear_goals_model.R.")
}
```

## Elo Feature

For team $i$ against opponent $j$, expected result before kickoff is:

$$
p_{i,m} =
\frac{1}
{1 + 10^{-\left(E_{i,m}^{pre} + H_m - E_{j,m}^{pre}\right)/400}}
$$

Where:

- $E_{i,m}^{pre}$ is team $i$'s pre-match Elo.
- $E_{j,m}^{pre}$ is opponent pre-match Elo.
- $H_m$ is the home advantage adjustment.
- $p_{i,m}$ is expected result on a 0 to 1 scale.

## Model

The first regression is:

$$
Y_{i,m}
= \beta_0
+ \beta_1 E_{i,m}^{pre}
+ \beta_2 E_{j,m}^{pre}
+ \beta_3 p_{i,m}
+ \beta_4 I(\text{listed home}_{i,m})
+ \beta_5 I(\text{neutral site}_m)
+ \epsilon_{i,m}
$$

with:

$$
\epsilon_{i,m} \sim N(0, \sigma^2)
$$

## Results

```{r metrics}
metrics <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "goals_linear_model_metrics.csv"))
coefficients <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "goals_linear_model_coefficients.csv"))

if (nrow(metrics) > 0) metrics else data.frame(note = "Metrics are generated locally.")
```

```{r coefficients}
if (nrow(coefficients) > 0) coefficients else data.frame(note = "Coefficients are generated locally.")
```

## Interpretation

The signs match soccer intuition:

- `pre_elo` is positive: stronger teams score more.
- `opponent_pre_elo` is negative: stronger opponents suppress scoring.
- `pre_match_expected_result` is positive: teams expected to perform better score more.
- `listed_homeTRUE` is positive: listed home teams score more.

## Reproduce Locally

```{r reproduce, eval=FALSE}
source("R/01_build_duckdb.R")
source("R/10_fit_linear_goals_model.R")
```
