---
title: "2026 World Cup Modeling Baseline: Goals Linear Regression"
author: "World Cup Betting Model Project"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)
```

# Purpose

This report documents the first transparent baseline model for the World Cup prediction project.

The goal is not to claim that ordinary least squares is the correct final model for soccer goals. It is a clean starting point:

- The dataset uses real historical international football data.
- The outcome is easy to inspect.
- The model uses only features known before kickoff.
- The results are reproducible from RStudio.

Later reports can switch the target to binary win/loss or ordinal loss/draw/win.

# Project Setup

Open the RStudio project:

```text
world-cup-betting-model.Rproj
```

Then run:

```{r setup-project, eval=FALSE}
source("R/00_setup.R")
source("R/01_build_duckdb.R")
source("R/06_data_inventory.R")
source("R/10_fit_linear_goals_model.R")
```

# Data Sources

The current free/public data layer includes:

- Historical international results.
- Historical player goal events.
- 2026 World Cup fixtures.
- 2026 World Cup squad players.
- Player caps, goals, clubs, positions, and ages.
- Venue and weather context.
- Wikidata player enrichment where available.
- GDELT news metadata, currently a partial/test pull.
- Computed team Elo ratings.

Large raw and processed files are kept local. A small shareable sample lives at:

```text
data/samples/goals_linear_model_sample_1000.csv
```

# Unit Of Observation

The modeling dataset has one row per **team-match**.

If Argentina plays France, the data has two rows:

- Argentina's goals as the target, with France as the opponent.
- France's goals as the target, with Argentina as the opponent.

This format is convenient because the target is always:

$$
Y_{i,m} = \text{goals scored by team } i \text{ in match } m
$$

# Sample Data

```{r load-sample}
sample_path <- file.path("..", "data", "samples", "goals_linear_model_sample_1000.csv")

if (!file.exists(sample_path)) {
  sample_path <- file.path("data", "samples", "goals_linear_model_sample_1000.csv")
}

sample_data <- readr::read_csv(sample_path, show_col_types = FALSE)
head(sample_data, 10)
```

The sample columns are:

```{r sample-columns}
data.frame(
  column = names(sample_data),
  example_value = as.character(sample_data[1, ])
)
```

# Modeling Target

The first target is:

$$
Y_{i,m} = \texttt{y\_goals\_for}
$$

This is the number of goals scored by team $i$ in match $m$.

# Pre-Match Elo Feature

The project computes a simple Elo rating from historical match results.

For team $i$ against opponent $j$, the expected result before kickoff is:

$$
p_{i,m} =
\frac{1}
{1 + 10^{-\left(E_{i,m}^{pre} + H_m - E_{j,m}^{pre}\right)/400}}
$$

Where:

- $E_{i,m}^{pre}$ is team $i$'s pre-match Elo.
- $E_{j,m}^{pre}$ is the opponent's pre-match Elo.
- $H_m$ is a home advantage adjustment.
- $p_{i,m}$ is the expected match result on a 0 to 1 scale.

This is not the final team-strength model. It is a transparent baseline signal.

# First Linear Model

The first regression is:

$$
Y_{i,m}
= \beta_0
+ \beta_1 E_{i,m}^{pre}
+ \beta_2 E_{j,m}^{pre}
+ \beta_3 p_{i,m}
+ \beta_4 I(\text{listed home}_{i,m})
+ \beta_5 I(\text{neutral site}_m)
+ \epsilon_{i,m}
$$

with:

$$
\epsilon_{i,m} \sim N(0, \sigma^2)
$$

This is deliberately simple. Goals are count data, so a later model should use a Poisson, negative-binomial, bivariate Poisson, or Skellam-style setup.

# Fit The Model

The model has already been fit by:

```{r fit-command, eval=FALSE}
source("R/10_fit_linear_goals_model.R")
```

Load the saved metrics and coefficients:

```{r load-model-output}
metrics_path <- file.path("..", "data", "processed", "modeling", "goals_linear_model_metrics.csv")
coef_path <- file.path("..", "data", "processed", "modeling", "goals_linear_model_coefficients.csv")

if (!file.exists(metrics_path)) {
  metrics_path <- file.path("data", "processed", "modeling", "goals_linear_model_metrics.csv")
  coef_path <- file.path("data", "processed", "modeling", "goals_linear_model_coefficients.csv")
}

metrics <- readr::read_csv(metrics_path, show_col_types = FALSE)
coefficients <- readr::read_csv(coef_path, show_col_types = FALSE)

metrics
```

```{r coefficients}
coefficients
```

# Interpretation

The important coefficient signs make sense:

- `pre_elo` is positive: stronger teams score more.
- `opponent_pre_elo` is negative: stronger opponents suppress scoring.
- `pre_match_expected_result` is positive: teams expected to perform better score more.
- `listed_homeTRUE` is positive: listed home teams score more.
- `neutralTRUE` is positive in this first model, but this should not be over-interpreted until the venue/home coding is refined.

# Why This Is Only A Baseline

Linear regression is easy to read, but goals are non-negative counts. OLS can predict impossible negative values and does not model the natural distribution of soccer scores.

Better next-step goal models:

$$
Y_{i,m} \sim \text{Poisson}(\lambda_{i,m})
$$

$$
\log(\lambda_{i,m}) = X_{i,m}\beta
$$

For win/draw/loss, an ordinal model could be:

$$
\text{logit}\left(P(Y_{i,m}^{ord} \leq k)\right)
= \theta_k - X_{i,m}\beta
$$

where:

$$
Y_{i,m}^{ord} \in \{\text{loss}, \text{draw}, \text{win}\}
$$

and $k$ indexes the ordered cutpoints.

# Next Modeling Steps

1. Add recent form windows: last 5, 10, and 20 matches before each match.
2. Add squad strength features: total caps, total goals, age profile, position mix.
3. Add weather and venue context.
4. Add news features from GDELT: injury terms, lineup terms, and article-volume spikes.
5. Add odds once API keys are available.
6. Move from OLS to Poisson or negative-binomial goals.
7. Build binary and ordinal outcome models.
8. Calibrate predicted probabilities.

# Reproducibility Checklist

Run these from RStudio:

```{r reproducibility, eval=FALSE}
source("R/00_setup.R")
source("R/01_build_duckdb.R")
source("R/06_data_inventory.R")
source("R/10_fit_linear_goals_model.R")
rmarkdown::render("reports/01_goals_linear_regression.Rmd")
```

