---
title: "2026 World Cup Modeling Baseline: Ordinal Result Model"
author: "World Cup Betting Model Project"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)
```

# Purpose

This report adds the first result-probability model to the project.

The outcome is ordered:

$$
Y_{i,m}^{ord} \in \{\text{loss}, \text{draw}, \text{win}\}
$$

with:

$$
\text{loss} < \text{draw} < \text{win}
$$

This is closer to the betting question than the first goals model because it estimates probabilities for match outcomes.

# Unit Of Observation

The data has one row per **team-match**.

If Argentina plays France, there are two rows:

- Argentina result versus France.
- France result versus Argentina.

That means the modeled outcome is the result from the listed team's perspective.

# Sample Data

```{r load-sample}
sample_path <- file.path("..", "data", "processed", "modeling", "result_ordinal_model_sample_1000.csv")

if (!file.exists(sample_path)) {
  sample_path <- file.path("data", "processed", "modeling", "result_ordinal_model_sample_1000.csv")
}

sample_data <- readr::read_csv(sample_path, show_col_types = FALSE)
head(sample_data, 10)
```

# Model

The proportional-odds ordinal logistic model is:

$$
\text{logit}\left(P(Y_{i,m}^{ord} \leq k)\right)
= \theta_k - X_{i,m}\beta
$$

where:

$$
k \in \{\text{loss}, \text{draw}\}
$$

and:

$$
X_{i,m}\beta
= \beta_1 E_{i,m}^{pre}
+ \beta_2 E_{j,m}^{pre}
+ \beta_3 p_{i,m}
+ \beta_4 I(\text{listed home}_{i,m})
+ \beta_5 I(\text{neutral site}_m)
$$

The predictors match the first goals model:

- Team pre-match Elo.
- Opponent pre-match Elo.
- Elo-derived expected result.
- Listed-home indicator.
- Neutral-site indicator.

# Fit The Model

The model is fit by:

```{r fit-command, eval=FALSE}
source("R/11_fit_ordinal_result_model.R")
```

Load the saved outputs:

```{r load-model-output}
metrics_path <- file.path("..", "data", "processed", "modeling", "result_ordinal_model_metrics.csv")
coef_path <- file.path("..", "data", "processed", "modeling", "result_ordinal_model_coefficients.csv")
class_path <- file.path("..", "data", "processed", "modeling", "result_ordinal_model_class_distribution.csv")
confusion_path <- file.path("..", "data", "processed", "modeling", "result_ordinal_model_confusion_matrix.csv")
pred_path <- file.path("..", "data", "processed", "modeling", "result_ordinal_model_test_predictions.csv")

if (!file.exists(metrics_path)) {
  metrics_path <- file.path("data", "processed", "modeling", "result_ordinal_model_metrics.csv")
  coef_path <- file.path("data", "processed", "modeling", "result_ordinal_model_coefficients.csv")
  class_path <- file.path("data", "processed", "modeling", "result_ordinal_model_class_distribution.csv")
  confusion_path <- file.path("data", "processed", "modeling", "result_ordinal_model_confusion_matrix.csv")
  pred_path <- file.path("data", "processed", "modeling", "result_ordinal_model_test_predictions.csv")
}

metrics <- readr::read_csv(metrics_path, show_col_types = FALSE)
coefficients <- readr::read_csv(coef_path, show_col_types = FALSE)
class_distribution <- readr::read_csv(class_path, show_col_types = FALSE)
confusion <- readr::read_csv(confusion_path, show_col_types = FALSE)
predictions <- readr::read_csv(pred_path, show_col_types = FALSE)

metrics
```

# Class Balance

```{r class-distribution}
class_distribution
```

# Coefficients

```{r coefficients}
coefficients
```

Positive slope coefficients move probability mass toward better outcomes.

For example, if `pre_elo` is positive, higher team Elo increases the probability of `win` and decreases the probability of `loss`, holding other features constant.

# Test Predictions

```{r predictions}
head(
  predictions[
    ,
    c(
      "date",
      "team",
      "opponent",
      "y_result_ordered",
      "predicted_result",
      "pred_prob_loss",
      "pred_prob_draw",
      "pred_prob_win"
    )
  ],
  12
)
```

# Confusion Matrix

```{r confusion}
confusion
```

# Interpretation

The model is useful because it produces three probabilities:

$$
P(\text{loss}), \quad P(\text{draw}), \quad P(\text{win})
$$

That is the shape we eventually need for betting-market comparison.

But this should still be treated as a baseline:

- The proportional-odds assumption may be too restrictive.
- Soccer outcomes are not naturally one-team independent because each match creates paired rows.
- Draws are hard to model and usually need richer features.
- Market odds will become a critical benchmark once we add them.

# Next Modeling Steps

1. Add rolling recent-form features before each match.
2. Add squad strength and player availability.
3. Add team-level news signals.
4. Add odds-implied probabilities.
5. Compare ordinal logistic against multinomial logistic regression.
6. Calibrate probabilities with held-out data.

# Reproducibility Checklist

Run these from RStudio:

```{r reproducibility, eval=FALSE}
source("R/00_setup.R")
source("R/01_build_duckdb.R")
source("R/11_fit_ordinal_result_model.R")
source("R/12_render_reports.R")
```
