---
title: "Ordinal Result Model"
subtitle: "Baseline model for loss, draw, and win probabilities"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
root <- if (file.exists("data/processed/modeling")) "." else ".."
read_csv_if_exists <- function(path) {
  if (file.exists(path)) {
    readr::read_csv(path, show_col_types = FALSE)
  } else {
    data.frame()
  }
}
```

## Purpose

This report adds the first result-probability model.

The ordered outcome is:

$$
Y_{i,m}^{ord} \in \{\text{loss}, \text{draw}, \text{win}\}
$$

with:

$$
\text{loss} < \text{draw} < \text{win}
$$

This is closer to betting-market comparison because it estimates match outcome probabilities.

## Unit Of Observation

The data has one row per **team-match**.

If Argentina plays France, there are two rows:

- Argentina result versus France.
- France result versus Argentina.

## Embedded Sample Rows

No sample CSV is published with the website. These rows are embedded into the rendered HTML from the local model frame.

```{r sample-rows}
sample_data <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "result_ordinal_model_sample_1000.csv"))

if (nrow(sample_data) > 0) {
  head(sample_data, 10)
} else {
  data.frame(note = "Sample rows are generated locally by R/11_fit_ordinal_result_model.R.")
}
```

## Model

The proportional-odds ordinal logistic model is:

$$
\text{logit}\left(P(Y_{i,m}^{ord} \leq k)\right)
= \theta_k - X_{i,m}\beta
$$

where:

$$
k \in \{\text{loss}, \text{draw}\}
$$

and:

$$
X_{i,m}\beta
= \beta_1 E_{i,m}^{pre}
+ \beta_2 E_{j,m}^{pre}
+ \beta_3 p_{i,m}
+ \beta_4 I(\text{listed home}_{i,m})
+ \beta_5 I(\text{neutral site}_m)
$$

## Results

```{r metrics}
metrics <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "result_ordinal_model_metrics.csv"))
class_distribution <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "result_ordinal_model_class_distribution.csv"))
coefficients <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "result_ordinal_model_coefficients.csv"))
confusion <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "result_ordinal_model_confusion_matrix.csv"))
predictions <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "result_ordinal_model_test_predictions.csv"))

if (nrow(metrics) > 0) metrics else data.frame(note = "Metrics are generated locally.")
```

## Class Balance

```{r class-balance}
if (nrow(class_distribution) > 0) class_distribution else data.frame(note = "Class balance is generated locally.")
```

## Coefficients

```{r coefficients}
if (nrow(coefficients) > 0) coefficients else data.frame(note = "Coefficients are generated locally.")
```

Positive slope coefficients move probability mass toward better outcomes.

## Test Prediction Examples

```{r predictions}
if (nrow(predictions) > 0) {
  head(
    predictions[
      ,
      c(
        "date",
        "team",
        "opponent",
        "y_result_ordered",
        "predicted_result",
        "pred_prob_loss",
        "pred_prob_draw",
        "pred_prob_win"
      )
    ],
    12
  )
} else {
  data.frame(note = "Predictions are generated locally.")
}
```

## Confusion Matrix

```{r confusion}
if (nrow(confusion) > 0) confusion else data.frame(note = "Confusion matrix is generated locally.")
```

## Interpretation

The model produces:

$$
P(\text{loss}), \quad P(\text{draw}), \quad P(\text{win})
$$

This is the shape needed for future odds comparison.

Still, this is a baseline:

- The proportional-odds assumption may be too restrictive.
- Paired team-match rows are not independent.
- Draws usually need richer features.
- Market odds will be the key benchmark once added.

## Reproduce Locally

```{r reproduce, eval=FALSE}
source("R/01_build_duckdb.R")
source("R/11_fit_ordinal_result_model.R")
```
