---
title: "Result Probability Model"
subtitle: "Loss, draw, and win probabilities"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
root <- if (file.exists("data/processed/modeling")) "." else ".."
source(file.path(root, "R", "report_helpers.R"))
read_csv_if_exists <- function(path) {
  if (file.exists(path)) {
    readr::read_csv(path, show_col_types = FALSE)
  } else {
    data.frame()
  }
}
```

## Purpose

This report documents the result-probability model.

The ordered outcome is:

$$
Y_{i,m}^{ord} \in \{\text{loss}, \text{draw}, \text{win}\}
$$

with:

$$
\text{loss} < \text{draw} < \text{win}
$$

This output supports betting-market comparison because it estimates match outcome probabilities.

## Unit Of Observation

The data has one row per **team-match**.

If Argentina plays France, there are two rows:

- Argentina result versus France.
- France result versus Argentina.

## Embedded Sample Rows

No sample CSV is published with the website. These rows are embedded into the rendered HTML from the local model frame.

```{r sample-rows}
sample_data <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "result_ordinal_model_sample_1000.csv"))

if (nrow(sample_data) > 0) {
  sample_data |>
    dplyr::select(
      date,
      team,
      opponent,
      y_result_ordered,
      pre_elo,
      opponent_pre_elo,
      pre_match_expected_result
    ) |>
    dplyr::mutate(
      pre_elo = round(pre_elo, 0),
      opponent_pre_elo = round(opponent_pre_elo, 0),
      pre_match_expected_result = fmt_percent(pre_match_expected_result, 1)
    ) |>
    display_table(caption = "Team-match sample used by the result model.", max_rows = 10)
} else {
  display_table(empty_note("Sample rows are available after the result model refresh."))
}
```

## Model

The model estimates the chance that a team result is at or below each ordered level:

$$
\text{logit}\left(P(\text{Result} \leq k)\right)
= \theta_k - \text{Team Score}
$$

where:

$$
k \in \{\text{loss}, \text{draw}\}
$$

and:

$$
\text{Team Score}
= \beta_1\text{Team Elo}
+ \beta_2\text{Opponent Elo}
+ \beta_3\text{Expected Result}
+ \beta_4\text{Listed Home}
+ \beta_5\text{Neutral Site}
$$

## Results

```{r metrics}
metrics <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "result_ordinal_model_metrics.csv"))
class_distribution <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "result_ordinal_model_class_distribution.csv"))
coefficients <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "result_ordinal_model_coefficients.csv"))
confusion <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "result_ordinal_model_confusion_matrix.csv"))
predictions <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "result_ordinal_model_test_predictions.csv"))

if (nrow(metrics) > 0) {
  data.frame(
    rows_total = fmt_integer(metrics$rows_total),
    rows_train = fmt_integer(metrics$rows_train),
    rows_test = fmt_integer(metrics$rows_test),
    test_accuracy = fmt_percent(metrics$test_accuracy, 1),
    reference_accuracy = fmt_percent(metrics$baseline_majority_accuracy, 1),
    test_log_loss = fmt_number(metrics$test_log_loss, 3),
    test_multiclass_brier = fmt_number(metrics$test_multiclass_brier, 3)
  ) |>
    display_table(caption = "Ordinal model fit summary.")
} else {
  display_table(empty_note("Metrics are available after the result model refresh."))
}
```

## Class Balance

```{r class-balance}
if (nrow(class_distribution) > 0) {
  class_distribution |>
    dplyr::mutate(
      rows = fmt_integer(rows),
      share = fmt_percent(share, 1)
    ) |>
    display_table(caption = "Outcome balance in the modeling frame.")
} else {
  display_table(empty_note("Class balance is available after the result model refresh."))
}
```

## Coefficients

```{r coefficients}
if (nrow(coefficients) > 0) {
  coefficients |>
    dplyr::mutate(
      estimate = fmt_number(estimate, 4),
      std_error = fmt_number(std_error, 4),
      t_value = fmt_number(t_value, 2),
      p_value_approx = fmt_p_value(p_value_approx)
    ) |>
    display_table(caption = "Ordinal logistic coefficients and thresholds.")
} else {
  display_table(empty_note("Coefficients are available after the result model refresh."))
}
```

Positive slope coefficients move probability mass toward better outcomes.

## Test Prediction Examples

```{r predictions}
if (nrow(predictions) > 0) {
  predictions |>
    dplyr::select(
      date,
      team,
      opponent,
      y_result_ordered,
      predicted_result,
      pred_prob_loss,
      pred_prob_draw,
      pred_prob_win
    ) |>
    dplyr::mutate(
      pred_prob_loss = fmt_percent(pred_prob_loss, 1),
      pred_prob_draw = fmt_percent(pred_prob_draw, 1),
      pred_prob_win = fmt_percent(pred_prob_win, 1)
    ) |>
    display_table(caption = "Example test-set predictions.", max_rows = 12)
} else {
  display_table(empty_note("Predictions are available after the result model refresh."))
}
```

## Confusion Matrix

```{r confusion}
if (nrow(confusion) > 0) {
  confusion |>
    dplyr::mutate(rows = fmt_integer(rows)) |>
    display_table(caption = "Confusion matrix counts by actual and predicted result.")
} else {
  display_table(empty_note("Confusion matrix is available after the result model refresh."))
}
```

```{r confusion-plot, fig.cap="Confusion matrix as a heatmap; darker cells have more team-match rows."}
if (nrow(confusion) > 0 && requireNamespace("ggplot2", quietly = TRUE)) {
  confusion |>
    ggplot2::ggplot(ggplot2::aes(x = predicted_result, y = actual_result, fill = rows)) +
    ggplot2::geom_tile(color = "white", linewidth = 0.8) +
    ggplot2::geom_text(ggplot2::aes(label = format(rows, big.mark = ",")), color = "white", fontface = "bold") +
    ggplot2::scale_fill_gradient(low = "#9cc7c7", high = "#157878") +
    ggplot2::labs(
      title = "Ordinal result model confusion matrix",
      x = "Predicted result",
      y = "Actual result",
      fill = "Rows"
    ) +
    ggplot2::theme_minimal(base_size = 12)
} else {
  plot.new()
  text(0.5, 0.5, "Confusion matrix plot is available after the result model refresh.")
}
```

## Interpretation

The model produces:

$$
P(\text{loss}), \quad P(\text{draw}), \quad P(\text{win})
$$

These probabilities are the shape used for odds comparison.

Model checks tracked here:

- Proportional-odds fit.
- Paired team-match dependence.
- Draw calibration.
- Comparison against market-implied probabilities.

## Reproduce Locally

```{r reproduce, eval=FALSE}
source("R/01_build_duckdb.R")
source("R/11_fit_ordinal_result_model.R")
```
