---
title: "Distributions"
subtitle: "Probability models and diagnostics for the World Cup modeling workflow"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
root <- if (file.exists("data/processed/modeling")) "." else ".."
source(file.path(root, "R", "report_helpers.R"))
read_csv_if_exists <- function(path) {
  if (file.exists(path)) {
    readr::read_csv(path, show_col_types = FALSE)
  } else {
    data.frame()
  }
}
goals_frame <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "goals_linear_model_frame.csv"))
if (nrow(goals_frame) == 0) {
  goals_frame <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "goals_linear_model_sample_1000.csv"))
}
result_distribution <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "result_ordinal_model_class_distribution.csv"))
distribution_map <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "distribution_candidate_map.csv"))
distribution_diagnostics <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "distribution_diagnostics_summary.csv"))
distribution_autocorrelation <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "distribution_autocorrelation_tests.csv"))
expanded_population_summary <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "expanded_population_summary.csv"))
```

::: {.reference-note}
These original project notes are influenced by public NC State statistics teaching resources: short conceptual framing, visible math, reproducible R/Quarto code, and simulation/diagnostic thinking.
:::

## Purpose

Each variable has an explicit statistical shape. Soccer data mixes low-count outcomes, ordered categorical results, probabilities, text/news counts, player availability flags, and continuous context variables.

The workflow lists plausible distributions, builds diagnostics, and filters models that do not fit the data or the betting question.

## Expanded Population

```{r expanded-population, echo=FALSE, results="asis"}
if (nrow(expanded_population_summary) > 0) {
  key_metrics <- expanded_population_summary |>
    dplyr::filter(metric %in% c("rows", "matches", "teams", "first_match", "last_match", "feature_columns")) |>
    dplyr::mutate(
      label = dplyr::case_when(
        metric == "rows" ~ "Team-Match Rows",
        metric == "matches" ~ "Matches",
        metric == "teams" ~ "Teams",
        metric == "first_match" ~ "First Match",
        metric == "last_match" ~ "Latest Match",
        metric == "feature_columns" ~ "Feature Columns",
        TRUE ~ metric
      ),
      note = "expanded local population"
    ) |>
    dplyr::select(label, value, note)
  cat(metric_cards(key_metrics))
} else {
  display_table(empty_note("Expanded population metrics are available after running R/22_build_expanded_feature_population.R."))
}
```

## Current Outcome Shapes

```{r outcome-metrics, echo=FALSE, results="asis"}
if (nrow(goals_frame) > 0 && "y_goals_for" %in% names(goals_frame)) {
  goal_counts <- table(goals_frame$y_goals_for)
  goal_mean <- mean(goals_frame$y_goals_for, na.rm = TRUE)
  goal_var <- stats::var(goals_frame$y_goals_for, na.rm = TRUE)
  zero_share <- mean(goals_frame$y_goals_for == 0, na.rm = TRUE)

  cat(metric_cards(data.frame(
    label = c("Mean Goals", "Goal Variance", "Zero-Goal Share", "Max Goals"),
    value = c(
      fmt_number(goal_mean, 3),
      fmt_number(goal_var, 3),
      fmt_percent(zero_share, 1),
      fmt_integer(max(goals_frame$y_goals_for, na.rm = TRUE))
    ),
    note = c("team-match rows", "compare with mean", "possible count-model clue", "tail check")
  )))
}
```

```{r goal-distribution-table}
if (nrow(goals_frame) > 0 && "y_goals_for" %in% names(goals_frame)) {
  goals_frame |>
    dplyr::count(y_goals_for, name = "rows") |>
    dplyr::mutate(
      share = fmt_percent(rows / sum(rows), 1),
      rows = fmt_integer(rows)
    ) |>
    display_table(caption = "Observed team-goal distribution.", max_rows = 12)
} else {
  display_table(empty_note("Goal distribution is available after the modeling refresh."))
}
```

```{r goal-distribution-plot, fig.cap="Observed team-goal counts with a Poisson reference curve using the empirical mean."}
if (nrow(goals_frame) > 0 && "y_goals_for" %in% names(goals_frame) && requireNamespace("ggplot2", quietly = TRUE)) {
  lambda_hat <- mean(goals_frame$y_goals_for, na.rm = TRUE)
  observed <- goals_frame |>
    dplyr::count(y_goals_for, name = "observed_rows") |>
    dplyr::mutate(observed_share = observed_rows / sum(observed_rows))
  poisson_ref <- data.frame(
    y_goals_for = seq(0, max(goals_frame$y_goals_for, na.rm = TRUE)),
    poisson_share = stats::dpois(seq(0, max(goals_frame$y_goals_for, na.rm = TRUE)), lambda = lambda_hat)
  )

  ggplot2::ggplot(observed, ggplot2::aes(x = y_goals_for, y = observed_share)) +
    ggplot2::geom_col(fill = "#157878", alpha = 0.82, width = 0.75) +
    ggplot2::geom_line(data = poisson_ref, ggplot2::aes(y = poisson_share), color = "#f2a65a", linewidth = 1.1) +
    ggplot2::geom_point(data = poisson_ref, ggplot2::aes(y = poisson_share), color = "#f2a65a", size = 2) +
    ggplot2::scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    ggplot2::labs(
      title = "Observed goals versus simple Poisson reference",
      x = "Goals scored by team",
      y = "Share of team-match rows"
    ) +
    ggplot2::theme_minimal(base_size = 12)
} else {
  plot.new()
  text(0.5, 0.5, "Goal distribution plot is available after the modeling refresh.")
}
```

## Candidate Distributions

<div class="distribution-grid">
  <div class="distribution-card"><strong>Goals scored</strong>Poisson, negative binomial, zero-inflated count models, bivariate Poisson, Skellam goal difference.<small>Goals are low nonnegative counts.</small></div>
  <div class="distribution-card"><strong>Match result</strong>Ordinal logistic, multinomial logistic, adjacent-category models, paired-score simulation.<small>Useful for win/draw/loss odds comparison.</small></div>
  <div class="distribution-card"><strong>Player and news counts</strong>Poisson or negative binomial with exposure offsets where needed.<small>Useful for goals, mentions, injuries, cards, and appearances.</small></div>
  <div class="distribution-card"><strong>Probabilities</strong>Beta, Dirichlet, calibration curves, Brier/log-loss summaries.<small>Useful once predictions are compared with market-implied probabilities.</small></div>
  <div class="distribution-card"><strong>Continuous context</strong>Normal, t, lognormal, gamma, empirical splines, or transformed variables.<small>Useful for weather, travel, rest, Elo differences, and player minutes.</small></div>
  <div class="distribution-card"><strong>Latent team strength</strong>Gaussian random effects, dynamic state-space ratings, hierarchical priors.<small>Useful when teams have uneven schedule strength and sparse recent games.</small></div>
</div>

```{r distribution-map}
if (nrow(distribution_map) > 0) {
  distribution_map |>
    display_table(caption = "How each distribution is used in the modeling workflow.")
} else {
  display_table(empty_note("Distribution mapping is available after running R/24_distribution_diagnostics.R."))
}
```

## Distribution Test Results

The table below does not ask whether a distribution is perfect. With a large
historical soccer dataset, almost every simple distribution will fail a strict
test. The useful question is practical: does the diagnostic support using the
distribution as a model, a feature, a simulation tool, or a residual check?

```{r distribution-diagnostics}
if (nrow(distribution_diagnostics) > 0) {
  distribution_diagnostics |>
    dplyr::mutate(
      rows = fmt_integer(rows),
      estimate = fmt_number(estimate, 3),
      statistic = fmt_number(statistic, 3),
      p_value = fmt_p_value(p_value)
    ) |>
    dplyr::select(
      distribution,
      target,
      rows,
      estimate,
      statistic,
      p_value,
      decision,
      interpretation
    ) |>
    display_table(caption = "Distribution diagnostics and modeling decision.")
} else {
  display_table(empty_note("Distribution diagnostics are available after running R/24_distribution_diagnostics.R."))
}
```

```{r distribution-test-plot, fig.cap="Distribution diagnostic p-values. Very small p-values are common with large soccer datasets and should be read with the modeling decision."}
if (nrow(distribution_diagnostics) > 0 && requireNamespace("ggplot2", quietly = TRUE)) {
  plot_data <- distribution_diagnostics |>
    dplyr::filter(!is.na(p_value), p_value > 0) |>
    dplyr::mutate(
      log10_p = -log10(p_value),
      distribution = factor(distribution, levels = rev(unique(distribution)))
    )

  ggplot2::ggplot(plot_data, ggplot2::aes(x = log10_p, y = distribution)) +
    ggplot2::geom_col(fill = "#0f6b68", width = 0.64) +
    ggplot2::geom_vline(xintercept = -log10(0.05), color = "#d9a441", linewidth = 0.9, linetype = "dashed") +
    ggplot2::labs(
      title = "Distribution checks by evidence against a simple fit",
      x = "-log10(p-value)",
      y = NULL
    ) +
    ggplot2::theme_minimal(base_size = 12)
} else {
  plot.new()
  text(0.5, 0.5, "Distribution diagnostic plot is available after the diagnostics refresh.")
}
```

## Autocorrelation Checks

Autocorrelation means model errors are related across nearby rows in time. If
errors stay positive or negative for a stretch, the model may be missing a time
pattern such as team form, schedule congestion, or changing tournament style.

```{r autocorrelation-tests}
if (nrow(distribution_autocorrelation) > 0) {
  distribution_autocorrelation |>
    dplyr::mutate(
      autocorrelation = fmt_number(autocorrelation, 3),
      durbin_watson = fmt_number(durbin_watson, 3)
    ) |>
    display_table(caption = "Residual autocorrelation checks.")
} else {
  display_table(empty_note("Autocorrelation checks are available after running R/24_distribution_diagnostics.R."))
}
```

## Modeling Implications

The current diagnostics point to these practical choices:

- Use **Poisson** and **negative binomial** families for goals, cards, injuries,
  and news-count features because these targets are event counts.
- Keep **Bernoulli**, **binomial**, and **beta** thinking around win flags,
  scoring flags, form-window summaries, and probability calibration.
- Use **gamma**, **exponential**, and **geometric** ideas for rest, event timing,
  and scoring drought features, but only after they improve validation metrics.
- Use **normal**, **t**, **chi-squared**, and **F** distributions mainly for
  diagnostics: residual shape, autocorrelation, heteroskedasticity, goodness of
  fit, and variance shifts.
- Use **uniform** draws for Monte Carlo simulation and randomized validation,
  not as a claim about match outcomes.

```{r candidate-table}
candidate_models <- data.frame(
  data_family = c(
    "Team goals",
    "Goal difference",
    "Win/draw/loss",
    "News article counts",
    "Player availability",
    "Market probabilities",
    "Weather and travel"
  ),
  statistical_shape = c(
    "Low nonnegative count",
    "Integer contrast",
    "Ordered or nominal category",
    "Sparse count",
    "Binary or proportion",
    "Compositional probabilities",
    "Continuous context"
  ),
  model_options = c(
    "Poisson, negative binomial, bivariate Poisson",
    "Skellam, ordered result model",
    "Ordinal logistic, multinomial logistic",
    "Poisson, negative binomial, topic indicators",
    "Logistic regression, hierarchical availability prior",
    "Calibration, beta/Dirichlet summaries, log loss",
    "Splines, t errors, transformed linear effects"
  ),
  diagnostic = c(
    "Mean-variance check, zero share, tail fit",
    "Symmetry and draw concentration",
    "Class balance, calibration by class",
    "Overdispersion and burstiness",
    "Missingness and source reliability",
    "Reliability curves and bookmaker margin",
    "Nonlinearity, skew, outliers"
  )
)

display_table(candidate_models, caption = "Distribution options for the model.")
```

## Review Notes Captured

The public materials reviewed emphasize a few habits that fit this project well:

- Use computing to build intuition, not just final answers.
- Put distribution assumptions beside visuals so model choices are auditable.
- Compare models on held-out or cross-validated performance, not training fit alone.
- Teach the reader the data grain before showing the model.
- Keep source code reproducible while making the public page readable.

Sources reviewed include Professor Justin Post's personal/course site, his ST 558 course repository, his Shiny teaching app page, and the Distributome project linked from those teaching resources.

Reference links:

- [Justin B. Post personal site](https://jbpost2.github.io/)
- [ST 558 Data Science for Statisticians](https://jbpost2.github.io/ST-558-Data-Science-for-Statisticians/)
- [R Shiny teaching apps](https://jbpost2.github.io/pages/ShinyApps.html)
- [Distributome](https://distributome.org/)

## Model Quality Checklist

1. Goals distribution diagnostics by tournament era and confederation.
2. Mean-versus-variance checks for count features.
3. Calibration plots for ordinal result probabilities.
4. Cross-validation by time split.
5. Predicted versus observed W/D/L rates by probability bucket.
6. Residual autocorrelation checks after every major model refresh.
7. Distribution-specific tests before promoting a challenger model.
