---
title: "KNN Similarity Model"
subtitle: "Nearest-neighbor comparison model for goals and match results"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
root <- if (file.exists("../data/processed/modeling")) ".." else "."
source(file.path(root, "R", "report_helpers.R"))
read_csv_if_exists <- function(path) {
  if (file.exists(path)) {
    readr::read_csv(path, show_col_types = FALSE)
  } else {
    data.frame()
  }
}
```

## Purpose

The KNN model is a similarity model. For a team in a match, it finds historical team-match rows that look most similar, then summarizes what happened in those nearby games.

This gives the project a useful contrast:

- OLS explains average goal changes with coefficients.
- Ordinal logistic explains ordered win/draw/loss probabilities.
- KNN asks: "What happened when past teams looked like this team in this matchup?"

## How It Works

KNN stands for **K-Nearest Neighbors**.

For each test row, the model:

1. Standardizes the model inputs so all features are on comparable scales.
2. Finds the `25` closest historical team-match rows.
3. Predicts goals by averaging the goals scored by those neighbors.
4. Predicts loss, draw, and win probabilities by counting neighbor outcomes.

```{r features}
features <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "knn_similarity_model_features.csv"))

if (nrow(features) > 0) {
  display_table(features, caption = "Features used to find similar team-match rows.")
} else {
  display_table(empty_note("KNN feature metadata is available after the KNN refresh."))
}
```

## Results

```{r metrics}
metrics <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "knn_similarity_model_metrics.csv"))

if (nrow(metrics) > 0) {
  metrics |>
    dplyr::transmute(
      rows_total = fmt_integer(rows_total),
      rows_train_evaluated = fmt_integer(rows_train_evaluated),
      rows_test_evaluated = fmt_integer(rows_test_evaluated),
      k_neighbors = k_neighbors,
      goals_rmse = fmt_number(goals_rmse, 3),
      goals_mae = fmt_number(goals_mae, 3),
      result_accuracy = fmt_percent(result_accuracy, 1),
      result_log_loss = fmt_number(result_log_loss, 3),
      result_multiclass_brier = fmt_number(result_multiclass_brier, 3),
      reference_accuracy = fmt_percent(reference_accuracy, 1)
    ) |>
    display_table(caption = "KNN model performance.")
} else {
  display_table(empty_note("KNN metrics are available after the KNN refresh."))
}
```

## Sample Predictions

```{r sample-predictions}
predictions <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "knn_similarity_model_sample_predictions.csv"))

if (nrow(predictions) > 0) {
  predictions |>
    dplyr::select(
      date,
      team,
      opponent,
      tournament,
      goals_for,
      pred_goals_for_knn,
      actual_result,
      predicted_result,
      pred_prob_loss,
      pred_prob_draw,
      pred_prob_win
    ) |>
    dplyr::mutate(
      pred_goals_for_knn = fmt_number(pred_goals_for_knn, 2),
      pred_prob_loss = fmt_percent(pred_prob_loss, 1),
      pred_prob_draw = fmt_percent(pred_prob_draw, 1),
      pred_prob_win = fmt_percent(pred_prob_win, 1)
    ) |>
    display_table(caption = "Sample KNN predictions.", max_rows = 12)
} else {
  display_table(empty_note("KNN predictions are available after the KNN refresh."))
}
```

## Interpretation

KNN is easy to explain because the prediction comes from comparable historical cases. That is also its limitation.

The model depends heavily on:

- The features used to define similarity.
- The number of neighbors.
- Whether older matches should count the same as recent matches.
- Whether rare outcomes, especially draws, are represented well enough in the neighbor set.

KNN is strongest as a comparison model. If it beats the simpler regression models, the feature set probably contains useful nonlinear patterns. If it loses, the added flexibility is not earning its keep.

## Reproduce Locally

```{r reproduce, eval=FALSE}
source("R/13_fit_knn_similarity_model.R")
```
