---
title: "Poisson Goals Model"
subtitle: "Count regression for expected team goals"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
root <- if (file.exists("data/processed/modeling")) "." else ".."
source(file.path(root, "R", "report_helpers.R"))
read_csv_if_exists <- function(path) {
  if (file.exists(path)) {
    readr::read_csv(path, show_col_types = FALSE)
  } else {
    data.frame()
  }
}
```

## Purpose

The Poisson goals model predicts how many goals a team is expected to score.

This is a better statistical shape for goals than simple linear regression because goals are counts: 0, 1, 2, 3, and so on. A Poisson model keeps the expected value positive and models scoring as a count process.

## Model

The model predicts expected goals, usually written as lambda:

$$
\lambda = E(\text{Goals})
$$

Instead of fitting goals directly, the model fits the log of expected goals:

$$
\log(\lambda)
= \beta_0
+ \beta_1\text{Team Elo}
+ \beta_2\text{Opponent Elo}
+ \beta_3\text{Expected Result}
+ \beta_4\text{Listed Home}
+ \beta_5\text{Neutral Site}
$$

Plain English: the model estimates a positive expected-goals number from team strength, opponent strength, expected result, and match context.

## Validation

Two checks are shown:

- **Random 80/20:** a general benchmark where rows are randomly split into training and test sets.
- **Train through 2018, test 2019+:** a harder forward-looking check that better resembles predicting future matches.

```{r metrics}
metrics <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "goals_poisson_model_metrics.csv"))

if (nrow(metrics) > 0) {
  metrics |>
    dplyr::mutate(
      validation_type = dplyr::case_when(
        validation_type == "random_80_20" ~ "Random 80/20",
        validation_type == "train_through_2018_test_2019_plus" ~ "Train through 2018; test 2019+",
        TRUE ~ validation_type
      ),
      rows_total = fmt_integer(rows_total),
      rows_train = fmt_integer(rows_train),
      rows_test = fmt_integer(rows_test),
      test_rmse = fmt_number(test_rmse, 3),
      test_mae = fmt_number(test_mae, 3),
      mean_y_test = fmt_number(mean_y_test, 3),
      mean_poisson_deviance = fmt_number(mean_poisson_deviance, 3)
    ) |>
    display_table(caption = "Poisson model validation summary.")
} else {
  display_table(empty_note("Poisson metrics are available after the goals model refresh."))
}
```

## Coefficients

The rate ratio column is the easiest way to read this model. A value above 1 increases expected goals. A value below 1 decreases expected goals.

```{r coefficients}
coefficients <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "goals_poisson_model_coefficients.csv"))

if (nrow(coefficients) > 0) {
  coefficients |>
    dplyr::mutate(
      estimate = fmt_number(estimate, 4),
      std_error = fmt_number(std_error, 4),
      z_value = fmt_number(z_value, 2),
      p_value = fmt_p_value(p_value),
      rate_ratio = fmt_number(rate_ratio, 4)
    ) |>
    display_table(caption = "Poisson model coefficients.")
} else {
  display_table(empty_note("Poisson coefficients are available after the goals model refresh."))
}
```

## Sample Predictions

```{r sample-predictions}
predictions <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "goals_poisson_model_test_predictions.csv"))

if (nrow(predictions) > 0) {
  predictions |>
    dplyr::filter(validation_type == "train_through_2018_test_2019_plus") |>
    dplyr::select(
      date,
      team,
      opponent,
      tournament,
      y_goals_for,
      pred_goals_for_poisson,
      residual
    ) |>
    dplyr::mutate(
      pred_goals_for_poisson = fmt_number(pred_goals_for_poisson, 2),
      residual = fmt_number(residual, 2)
    ) |>
    display_table(caption = "Sample forward-time Poisson predictions.", max_rows = 12)
} else {
  display_table(empty_note("Poisson predictions are available after the goals model refresh."))
}
```

## Interpretation

This model is intentionally compact. It supplies a count-based expected-goals benchmark, while roster, injury, lineup, odds, and live-form signals are handled as enrichment features in the broader scoring workflow.

## Reproduce Locally

```{r reproduce, eval=FALSE}
source("R/01_build_duckdb.R")
source("R/14_fit_poisson_goals_model.R")
```
