---
title: "Regression Diagnostics"
subtitle: "Collinearity, residual checks, and count-model fit"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
root <- if (file.exists("data/processed/modeling")) "." else ".."
source(file.path(root, "R", "report_helpers.R"))
read_csv_if_exists <- function(path) {
  if (file.exists(path)) {
    readr::read_csv(path, show_col_types = FALSE)
  } else {
    data.frame()
  }
}
```

## Purpose

This report checks whether the first regression models are behaving well enough to interpret.

The checks focus on three practical questions:

- Are the predictors repeating the same information?
- Do residuals show patterns the model missed?
- Is a count model a better shape for goals than simple linear regression?

## Summary

```{r summary-table}
diagnostics <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "regression_diagnostics_summary.csv"))

if (nrow(diagnostics) > 0) {
  diagnostics |>
    dplyr::mutate(
      value = dplyr::case_when(
        statistic %in% c("Breusch-Pagan p-value", "Jarque-Bera p-value") ~ fmt_p_value(value),
        TRUE ~ fmt_number(value, 3)
      )
    ) |>
    dplyr::select(check, statistic, value, status, public_interpretation) |>
    display_table(caption = "Regression diagnostic summary.")
} else {
  display_table(empty_note("Regression diagnostics are available after the local model refresh."))
}
```

## Data Used By The Models

```{r source-usage}
source_usage <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "api_capture_model_usage.csv"))

if (nrow(source_usage) > 0) {
  source_usage |>
    dplyr::mutate(rows_available = fmt_integer(rows_available)) |>
    display_table(caption = "Local data coverage used for modeling and scoring.")
} else {
  display_table(empty_note("Data coverage is available after the diagnostics refresh."))
}
```

## Collinearity

Collinearity means two or more predictors are carrying nearly the same signal. It does not always ruin predictions, but it can make individual coefficients unstable.

**VIF** means variance inflation factor. A VIF near 1 is clean. Values above 5 deserve attention. Values above 10 are usually too high for coefficient interpretation.

```{r vif-table}
vif <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "goals_model_vif.csv"))

if (nrow(vif) > 0) {
  vif |>
    dplyr::arrange(dplyr::desc(vif)) |>
    dplyr::mutate(vif = fmt_number(vif, 3)) |>
    display_table(caption = "Variance inflation factors for goals-model predictors.")
} else {
  display_table(empty_note("VIF results are available after the diagnostics refresh."))
}
```

```{r correlation-table}
correlations <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "goals_model_predictor_correlations.csv"))

if (nrow(correlations) > 0) {
  correlations |>
    dplyr::arrange(dplyr::desc(abs_correlation)) |>
    dplyr::mutate(
      correlation = fmt_number(correlation, 3),
      abs_correlation = fmt_number(abs_correlation, 3)
    ) |>
    display_table(caption = "Largest pairwise predictor correlations.", max_rows = 10)
} else {
  display_table(empty_note("Predictor correlations are available after the diagnostics refresh."))
}
```

## Residual Checks

A residual is:

$$
\text{Residual} = \text{Observed Goals} - \text{Predicted Goals}
$$

For OLS, residuals should not show strong structure after the model has done its work.

```{r residual-diagnostics}
residuals <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "goals_model_residual_diagnostics.csv"))

if (nrow(residuals) > 0) {
  residuals |>
    dplyr::mutate(value = dplyr::case_when(
      grepl("p_value", diagnostic) ~ fmt_p_value(value),
      TRUE ~ fmt_number(value, 4)
    )) |>
    display_table(caption = "OLS residual diagnostics.")
} else {
  display_table(empty_note("Residual diagnostics are available after the diagnostics refresh."))
}
```

## Poisson Checks

Poisson regression is a natural model for goals because goals are counts. The main issue is **overdispersion**.

Overdispersion means the observed goal counts vary more than a basic Poisson model expects. If this is large, the next goals model should consider a negative-binomial or paired-goal approach.

```{r poisson-diagnostics}
poisson <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "poisson_model_diagnostics.csv"))

if (nrow(poisson) > 0) {
  poisson |>
    dplyr::mutate(value = fmt_number(value, 4)) |>
    display_table(caption = "Poisson model diagnostics.")
} else {
  display_table(empty_note("Poisson diagnostics are available after the diagnostics refresh."))
}
```

## Interpretation

These diagnostics are the gate before model comparison:

- If VIF is low, coefficient directions are easier to trust.
- If Durbin-Watson is close to 2, residuals are not strongly sequential in the ordered rows.
- If heteroskedasticity appears, OLS standard errors should be treated carefully.
- If Poisson overdispersion appears, expected-goal predictions may still be useful, but uncertainty intervals need a wider model.

## Reproduce Locally

```{r reproduce, eval=FALSE}
source("R/15_regression_diagnostics.R")
```
