---
title: "Model Overview"
subtitle: "2026 match scores and win/draw/loss probabilities from local models"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
root <- if (file.exists("data/processed/modeling")) "." else ".."
source(file.path(root, "R", "report_helpers.R"))
read_csv_if_exists <- function(path) {
  if (file.exists(path)) {
    readr::read_csv(path, show_col_types = FALSE)
  } else {
    data.frame()
  }
}
```

## Purpose

This report scores the 2026 World Cup fixture list with the current local models.

Each fixture receives:

- OLS expected goals for the home and away team.
- Poisson expected goals for the home and away team.
- Ordinal win, draw, and loss probabilities from the home team's perspective.
- Weather context when venue-level weather summaries are available.

## Scoring Summary

```{r metrics}
metrics <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "world_cup_2026_fixture_prediction_metrics.csv"))

if (nrow(metrics) > 0) {
  metrics |>
    dplyr::mutate(
      value = dplyr::case_when(
        metric == "ordinal_completed_fixture_accuracy" ~ fmt_percent(value, 1),
        TRUE ~ fmt_number(value, 3)
      )
    ) |>
    dplyr::select(display_label, value, plain_english) |>
    display_table(caption = "Fixture scoring summary.")
} else {
  display_table(empty_note("Fixture prediction metrics are available after the scoring refresh."))
}
```

## Prediction Table

The table below shows a compact version of the scored fixture set. Probabilities are from the home-team perspective.

```{r predictions}
predictions <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "world_cup_2026_fixture_predictions.csv"))

if (nrow(predictions) > 0) {
  predictions |>
    dplyr::mutate(
      score = dplyr::case_when(
        !is.na(home_score) & !is.na(away_score) ~ paste0(home_score, "-", away_score),
        TRUE ~ ""
      ),
      poisson_expected_goals = paste0(
        fmt_number(pred_home_goals_poisson, 2),
        " - ",
        fmt_number(pred_away_goals_poisson, 2)
      ),
      ols_expected_goals = paste0(
        fmt_number(pred_home_goals_ols, 2),
        " - ",
        fmt_number(pred_away_goals_ols, 2)
      ),
      home_win = fmt_percent(pred_home_win_prob, 1),
      draw = fmt_percent(pred_draw_prob, 1),
      away_win = fmt_percent(pred_away_win_prob, 1),
      weather = dplyr::if_else(
        is.na(avg_temperature_2m),
        "",
        paste0(fmt_number(avg_temperature_2m, 1), " C, wind ", fmt_number(avg_wind_speed_10m, 1))
      )
    ) |>
    dplyr::select(
      date,
      home_team,
      away_team,
      score,
      poisson_expected_goals,
      ols_expected_goals,
      home_win,
      draw,
      away_win,
      predicted_result,
      city,
      weather
    ) |>
    display_table(caption = "Scored 2026 fixtures.", max_rows = 24)
} else {
  display_table(empty_note("Fixture predictions are available after the scoring refresh."))
}
```

## Completed Fixture Check

Completed matches provide an early live check of the model. The tournament sample is small, so this section is best read as monitoring rather than full model evaluation.

```{r completed}
if (nrow(predictions) > 0) {
  completed <- predictions |>
    dplyr::filter(!is.na(home_score), !is.na(away_score)) |>
    dplyr::mutate(
      actual_score = paste0(home_score, "-", away_score),
      poisson_expected_goals = paste0(
        fmt_number(pred_home_goals_poisson, 2),
        " - ",
        fmt_number(pred_away_goals_poisson, 2)
      ),
      probabilities = paste0(
        "H ",
        fmt_percent(pred_home_win_prob, 1),
        " / D ",
        fmt_percent(pred_draw_prob, 1),
        " / A ",
        fmt_percent(pred_away_win_prob, 1)
      )
    )

  if (nrow(completed) > 0) {
    completed |>
      dplyr::select(
        date,
        home_team,
        away_team,
        actual_score,
        actual_result,
        predicted_result,
        poisson_expected_goals,
        probabilities
      ) |>
      display_table(caption = "Completed fixtures scored by the current models.", max_rows = 16)
  } else {
    display_table(empty_note("Completed fixture checks will appear here as final scores enter the local fixture table."))
  }
} else {
  display_table(empty_note("Completed fixture checks are available after the scoring refresh."))
}
```

## Probability View

```{r probability-plot, fig.cap="Predicted home win, draw, and away win probabilities for upcoming fixtures."}
if (nrow(predictions) > 0 && requireNamespace("ggplot2", quietly = TRUE)) {
  wide_plot_data <- predictions |>
    dplyr::mutate(match = paste(home_team, "vs", away_team)) |>
    dplyr::arrange(date, source_match_id) |>
    dplyr::slice_head(n = 12) |>
    dplyr::select(match, pred_home_win_prob, pred_draw_prob, pred_away_win_prob)

  plot_data <- rbind(
    data.frame(match = wide_plot_data$match, outcome = "Home win", probability = wide_plot_data$pred_home_win_prob),
    data.frame(match = wide_plot_data$match, outcome = "Draw", probability = wide_plot_data$pred_draw_prob),
    data.frame(match = wide_plot_data$match, outcome = "Away win", probability = wide_plot_data$pred_away_win_prob)
  )
  plot_data$match <- factor(plot_data$match, levels = rev(unique(wide_plot_data$match)))

  ggplot2::ggplot(plot_data, ggplot2::aes(x = probability, y = match, fill = outcome)) +
    ggplot2::geom_col(position = "stack", width = 0.72) +
    ggplot2::scale_x_continuous(labels = function(x) paste0(round(100 * x), "%")) +
    ggplot2::scale_fill_manual(values = c("Home win" = "#0f6b68", "Draw" = "#d9a441", "Away win" = "#b85c5c")) +
    ggplot2::labs(
      title = "Fixture result probabilities",
      x = "Predicted probability",
      y = NULL,
      fill = NULL
    ) +
    ggplot2::theme_minimal(base_size = 12)
} else {
  plot.new()
  text(0.5, 0.5, "Probability plot is available after the scoring refresh.")
}
```

## Interpretation

The fixture table is the model's operational output:

- The Poisson goal estimates are the main expected-goals baseline.
- The ordinal model supplies win/draw/loss probabilities.
- OLS remains a transparent benchmark, but count models are better suited for goals.
- Weather is shown as context until enough historical weather coverage exists for coefficient estimation.

## Reproduce Locally

```{r reproduce, eval=FALSE}
source("R/16_score_2026_fixtures.R")
```
