---
title: "Matchday Prediction Board"
subtitle: "Winner, result probabilities, goals, cards coverage, and lineups"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
root <- if (file.exists("data/processed/modeling")) "." else ".."
source(file.path(root, "R", "report_helpers.R"))
read_csv_if_exists <- function(path) {
  if (file.exists(path)) {
    readr::read_csv(path, show_col_types = FALSE)
  } else {
    data.frame()
  }
}
```

## Prediction Board

This is the matchday view: one section for the current predictions and live-feed coverage.

```{r summary-cards, echo=FALSE, results="asis"}
summary <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "matchday_prediction_summary.csv"))
board <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "matchday_prediction_board.csv"))

summary_value <- function(metric_name) {
  if (nrow(summary) == 0) {
    return(NA_character_)
  }
  value <- summary$value[summary$metric == metric_name]
  if (length(value) == 0) NA_character_ else as.character(value[[1]])
}

if (nrow(summary) > 0) {
  cat(metric_cards(data.frame(
    label = c("Last Refresh", "Matches", "Today", "Upcoming", "Lineups", "Events"),
    value = c(
      summary_value("last_refreshed_local"),
      summary_value("matches_on_board"),
      summary_value("matches_today"),
      summary_value("upcoming_matches"),
      summary_value("lineup_rows_available"),
      summary_value("event_rows_available")
    ),
    note = c(
      "local laptop time",
      "prediction rows",
      "matchday count",
      "future matches",
      "API lineup rows",
      "API event rows"
    )
  )))
}
```

## Future Match Bracket

Double-click a matchup to open the model detail.

```{r bracket-interface, echo=FALSE, results="asis"}
escape_html <- function(x) {
  x <- ifelse(is.na(x), "", as.character(x))
  x <- gsub("&", "&amp;", x, fixed = TRUE)
  x <- gsub("<", "&lt;", x, fixed = TRUE)
  x <- gsub(">", "&gt;", x, fixed = TRUE)
  x <- gsub('"', "&quot;", x, fixed = TRUE)
  x
}

if (nrow(board) > 0) {
  board$kickoff_utc <- as.POSIXct(board$kickoff_utc_iso, format = "%Y-%m-%dT%H:%M:%SZ", tz = "UTC")
  future_board <- board |>
    dplyr::filter(
      match_timing != "Completed",
      (
        !is.na(kickoff_utc) & kickoff_utc > Sys.time()
      ) |
        (
          is.na(kickoff_utc) & as.Date(date) >= Sys.Date()
        )
    ) |>
    dplyr::arrange(kickoff_utc, date, source_match_id)

  if (nrow(future_board) > 0) {
    cat('<div class="bracket-board">')
    for (day in unique(as.character(future_board$date))) {
      day_rows <- future_board[as.character(future_board$date) == day, ]
      cat('<section class="bracket-day">')
      cat('<h3>', escape_html(format(as.Date(day), "%A, %B %d")), '</h3>', sep = "")
      cat('<div class="match-grid">')

      for (i in seq_len(nrow(day_rows))) {
        row <- day_rows[i, ]
        kickoff <- ifelse(
          is.na(row$kickoff_local_iso) || !nzchar(row$kickoff_local_iso),
          "Time pending",
          paste0(row$local_time, " ", row$utc_offset)
        )
        result_probs <- paste0(
          "H ",
          fmt_percent(row$ensemble_home_win_prob, 1),
          " / D ",
          fmt_percent(row$ensemble_draw_prob, 1),
          " / A ",
          fmt_percent(row$ensemble_away_win_prob, 1)
        )
        goals <- paste0(
          fmt_number(row$pred_home_goals_poisson, 2),
          " - ",
          fmt_number(row$pred_away_goals_poisson, 2)
        )
        cards <- if (
          is.finite(row$home_projected_yellow_cards) &&
            is.finite(row$away_projected_yellow_cards)
        ) {
          paste0(
            fmt_number(row$home_projected_yellow_cards, 1),
            " - ",
            fmt_number(row$away_projected_yellow_cards, 1)
          )
        } else {
          "Feed pending"
        }
        lineups <- paste0(row$home_lineup_status, " / ", row$away_lineup_status)
        refresh <- ifelse(
          is.na(row$refresh_utc_iso) || !nzchar(row$refresh_utc_iso),
          "Refresh time pending",
          paste("Refresh 30 min before:", row$refresh_utc_iso)
        )

        cat('<article class="match-card" tabindex="0">')
        cat('<div class="match-card-top">')
        cat('<span>', escape_html(kickoff), '</span>', sep = "")
        cat('<strong>', escape_html(row$confidence_band), '</strong>', sep = "")
        cat('</div>')
        cat('<div class="match-teams">')
        cat('<div><span>', escape_html(row$home_team), '</span><small>Home/listed</small></div>', sep = "")
        cat('<div class="versus">vs</div>')
        cat('<div><span>', escape_html(row$away_team), '</span><small>Away/listed</small></div>', sep = "")
        cat('</div>')
        cat('<div class="pick-line">')
        cat('<span>Winner</span><strong>', escape_html(row$predicted_winner), '</strong>', sep = "")
        cat('</div>')
        cat('<div class="model-picks">')
        cat('<span>OLS: ', escape_html(row$ols_predicted_winner), '</span>', sep = "")
        cat('<span>Poisson: ', escape_html(row$poisson_predicted_winner), '</span>', sep = "")
        cat('<span>Ordinal: ', escape_html(row$ordinal_predicted_winner), '</span>', sep = "")
        cat('</div>')
        cat('<div class="match-detail">')
        cat('<dl>')
        cat('<dt>Outcome probabilities</dt><dd>', escape_html(result_probs), '</dd>', sep = "")
        cat('<dt>Expected goals</dt><dd>', escape_html(goals), '</dd>', sep = "")
        cat('<dt>Most likely score</dt><dd>', escape_html(row$most_likely_score), '</dd>', sep = "")
        cat('<dt>Totals</dt><dd>O2.5 ', fmt_percent(row$over_2_5_prob, 1), ' / BTTS ', fmt_percent(row$both_teams_to_score_prob, 1), '</dd>', sep = "")
        cat('<dt>Yellow cards</dt><dd>', escape_html(cards), '</dd>', sep = "")
        cat('<dt>Lineups</dt><dd>', escape_html(lineups), '</dd>', sep = "")
        cat('<dt>Venue</dt><dd>', escape_html(paste(row$city, row$country, sep = ", ")), '</dd>', sep = "")
        cat('<dt>Refresh</dt><dd>', escape_html(refresh), '</dd>', sep = "")
        cat('</dl>')
        cat('</div>')
        cat('</article>')
      }

      cat('</div>')
      cat('</section>')
    }
    cat('</div>')
  } else {
    cat('<p class="empty-board">No future matches are available in the current local fixture table.</p>')
  }
} else {
  cat('<p class="empty-board">Matchday predictions are available after the matchday board refresh.</p>')
}
```

```{=html}
<script>
document.addEventListener("DOMContentLoaded", function () {
  document.querySelectorAll(".match-card").forEach(function (card) {
    card.addEventListener("dblclick", function () {
      card.classList.toggle("open");
    });
    card.addEventListener("keydown", function (event) {
      if (event.key === "Enter" || event.key === " ") {
        event.preventDefault();
        card.classList.toggle("open");
      }
    });
  });
});
</script>
```

```{r board-table}
if (nrow(board) > 0) {
  board$kickoff_utc <- as.POSIXct(board$kickoff_utc_iso, format = "%Y-%m-%dT%H:%M:%SZ", tz = "UTC")
  board_for_display <- board |>
    dplyr::filter(
      (
        !is.na(kickoff_utc) & kickoff_utc > Sys.time()
      ) |
        (
          is.na(kickoff_utc) & as.Date(date) >= Sys.Date() & match_timing != "Completed"
        )
    )

  display_board <- board_for_display |>
    dplyr::mutate(
      result_probability = paste0(
        "H ",
        fmt_percent(ensemble_home_win_prob, 1),
        " / D ",
        fmt_percent(ensemble_draw_prob, 1),
        " / A ",
        fmt_percent(ensemble_away_win_prob, 1)
      ),
      expected_goals = paste0(
        fmt_number(pred_home_goals_poisson, 2),
        " - ",
        fmt_number(pred_away_goals_poisson, 2)
      ),
      totals = paste0(
        "O2.5 ",
        fmt_percent(over_2_5_prob, 1),
        " / BTTS ",
        fmt_percent(both_teams_to_score_prob, 1)
      ),
      yellow_cards = dplyr::case_when(
        is.finite(home_projected_yellow_cards) & is.finite(away_projected_yellow_cards) ~
          paste0(
            fmt_number(home_projected_yellow_cards, 1),
            " - ",
            fmt_number(away_projected_yellow_cards, 1)
          ),
        TRUE ~ "Feed pending"
      ),
      lineup = paste0(home_lineup_status, " / ", away_lineup_status),
      confidence = paste0(confidence_band, " (", fmt_percent(prediction_confidence, 1), ")"),
      venue = paste(city, country, sep = ", ")
    ) |>
    dplyr::select(
      date,
      match_timing,
      match_label,
      predicted_winner,
      confidence,
      result_probability,
      expected_goals,
      most_likely_score,
      totals,
      yellow_cards,
      lineup,
      venue
    )

  if (nrow(display_board) > 0) {
    display_board |>
      display_table(caption = "Future matchday predictions.", max_rows = 24)
  } else {
    display_table(empty_note("No future matches are available in the current rendered board."))
  }
} else {
  display_table(empty_note("Matchday predictions are available after the matchday board refresh."))
}
```

## Lineup Detail

Confirmed lineups appear here when the match feed posts starting XIs.

```{r lineup-detail}
if (nrow(board) > 0) {
  lineup_detail <- board |>
    dplyr::filter(home_lineup_status == "Confirmed" | away_lineup_status == "Confirmed") |>
    dplyr::select(
      date,
      match_label,
      home_formation,
      away_formation,
      home_starting_lineup,
      away_starting_lineup
    )

  if (nrow(lineup_detail) > 0) {
    lineup_detail |>
      display_table(caption = "Confirmed starting lineups.", max_rows = 8)
  } else {
    display_table(empty_note("Starting lineups update when the fixture feed posts them."))
  }
} else {
  display_table(empty_note("Lineup detail is available after the matchday board refresh."))
}
```

## Post-Match Accuracy

This section updates after final scores enter the fixture table.

```{r accuracy}
accuracy <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "matchday_model_accuracy.csv"))

if (nrow(accuracy) > 0) {
  accuracy |>
    dplyr::mutate(
      completed_matches = fmt_integer(completed_matches),
      outcome_accuracy = fmt_percent(outcome_accuracy, 1)
    ) |>
    display_table(caption = "Post-match model accuracy on completed fixtures.")
} else {
  display_table(empty_note("Post-match accuracy updates after completed scores are available."))
}
```

## Models And Data Used

```{r data-sources}
sources <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "matchday_prediction_data_sources.csv"))

if (nrow(sources) > 0) {
  sources |>
    display_table(caption = "Prediction outputs, model source, and current role.")
} else {
  display_table(empty_note("Model-source notes are available after the matchday board refresh."))
}
```

The board uses three active layers:

- **Winner and outcome:** an ensemble of the ordinal logistic result model and probabilities derived from the Poisson score grid.
- **Goals and likely score:** the Poisson goals model produces expected goals; the score grid turns those expected goals into scoreline probabilities.
- **Totals:** over 2.5 goals and both-teams-to-score probabilities come from the same Poisson score grid.

Cards and lineups are live-enrichment layers. They update when the football API returns player card history, event rows, or confirmed lineups for a fixture.

## How To Read It

- **Predicted winner:** highest ensemble probability. A draw remains possible for group-stage style matches.
- **Confidence:** the model's top probability bucket, not a guarantee.
- **Expected goals:** model average score, not the most likely exact score.
- **Most likely score:** the single scoreline with the highest probability in the Poisson grid.
- **O2.5:** probability that total goals are 3 or more.
- **BTTS:** probability that both teams score at least one goal.

## Reproduce Locally

```{r reproduce, eval=FALSE}
source("R/16_score_2026_fixtures.R")
source("R/17_matchday_prediction_board.R")
```
