---
title: "Predictions"
subtitle: "Today, upcoming matches, bracket projections, and post-match model review."
description: "World Cup 2026 static match predictions with advance probabilities, regulation-level draw probabilities, expected scores, model performance, and a mobile-friendly bracket."
toc: false
number-sections: false
code-tools: false
code-fold: false
execute:
  echo: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
root <- if (file.exists("data/processed/modeling")) "." else ".."
source(file.path(root, "R", "report_helpers.R"))
source(file.path(root, "R", "matchday_components.R"))
source(file.path(root, "R", "static_readability_overrides.R"))
bundle <- load_matchday_bundle(root)
board <- bundle$board
summary <- bundle$summary
```

<section class="page-intro prediction-intro">
  <span class="section-kicker">Forecast board</span>
  <h1>Predictions</h1>
  <p>Today's slate, the next match, upcoming fixtures, bracket projections, and model accuracy in one place.</p>
</section>

<nav class="prediction-page-nav" aria-label="Prediction page sections">
  <a href="#today">Today</a>
  <a href="#next-match">Next match</a>
  <a href="#upcoming">Upcoming</a>
  <a href="#bracket">Bracket</a>
  <a href="#champion-outlook">Champion outlook</a>
  <a href="#model-performance">Accuracy</a>
  <a href="#methodology-details">Methodology</a>
</nav>

```{r status, results="asis"}
cat('<section class="prediction-status-bar" aria-label="Prediction status">')
cat('<span class="status-chip">Beta</span>')
cat('<span class="status-chip">', escape_html(tournament_phase(summary)), '</span>', sep = "")
cat('<span class="status-chip muted">Last updated ', escape_html(summary_value(summary, "last_refreshed_local")), '</span>', sep = "")
cat('<span class="status-chip muted">Static GitHub Pages output</span>')
cat('</section>')
```

```{r static-note, results="asis"}
cat(render_static_api_note(summary, compact = TRUE))
```

```{r today, results="asis"}
cat(render_today_section(board, summary))
```

```{r next-match, results="asis"}
cat(render_next_match_section(board))
```

```{r upcoming, results="asis"}
cat(render_upcoming_section(board, limit = 12))
```

## Tournament Bracket {#bracket}

```{r bracket, results="asis"}
cat(render_bracket(root))
```

```{r champion-outlook, results="asis"}
cat(render_champion_section(bundle, compact = FALSE))
```

```{r accuracy, results="asis"}
cat(render_performance_section(bundle, compact = FALSE))
cat(render_challenger_section(bundle, compact = FALSE))
```

## Methodology Details {#methodology-details}

```{r methodology, results="asis"}
cat(render_data_source_section(bundle))
cat(render_methodology_short())
cat(render_technical_prediction_table(board))
```

## Reading The Forecast

- **Win probability** is the model-estimated chance that a specific team wins the match in regulation or group-stage scoring.
- **Advance probability** appears for knockout matches. It includes the path through regulation, extra time, or penalties.
- **Level after regulation** is the chance that a knockout match is tied after 90 minutes. It is not treated as a final draw.
- **Draw probability** applies to matches where a draw can be the final result.
- **Prediction strength** is the top outcome probability grouped into plain language. It is not a guarantee.
- **Model agreement** counts how many baseline models point to the same winner, draw, or advancing team.
- **Expected goals** is the model average score, not the single most likely score.
- **Most likely score** is the exact scoreline with the largest probability in the score grid.
- **Over 2.5 goals** means the match has at least 3 total goals.
- **Both teams to score** means each team scores at least one goal.

<p class="responsible-note">Forecasts are probabilistic and should be used for research, discussion, and model review. They are not financial advice or guaranteed betting edges.</p>
