---
title: "Matchday Prediction Board"
subtitle: "Winner, result probabilities, goals, cards coverage, and lineups"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
root <- if (file.exists("data/processed/modeling")) "." else ".."
source(file.path(root, "R", "report_helpers.R"))
read_csv_if_exists <- function(path) {
  if (file.exists(path)) {
    readr::read_csv(path, show_col_types = FALSE)
  } else {
    data.frame()
  }
}

summary <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "matchday_prediction_summary.csv"))
board <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "matchday_prediction_board.csv"))

escape_html <- function(x) {
  x <- ifelse(is.na(x), "", as.character(x))
  x <- gsub("&", "&amp;", x, fixed = TRUE)
  x <- gsub("<", "&lt;", x, fixed = TRUE)
  x <- gsub(">", "&gt;", x, fixed = TRUE)
  x <- gsub('"', "&quot;", x, fixed = TRUE)
  x
}

safe_text <- function(x, fallback = "Pending") {
  if (length(x) == 0 || is.null(x)) {
    return(fallback)
  }
  text <- as.character(x[[1]])
  if (is.na(text) || !nzchar(text)) {
    fallback
  } else {
    text
  }
}

safe_number <- function(x) {
  if (length(x) == 0 || is.null(x)) {
    return(NA_real_)
  }
  suppressWarnings(as.numeric(x[[1]]))
}

display_percent <- function(x, digits = 1, fallback = "Pending") {
  value <- safe_number(x)
  if (is.na(value) || !is.finite(value)) {
    fallback
  } else {
    fmt_percent(value, digits)
  }
}

display_number <- function(x, digits = 2, fallback = "Pending") {
  value <- safe_number(x)
  if (is.na(value) || !is.finite(value)) {
    fallback
  } else {
    fmt_number(value, digits)
  }
}

probability_width <- function(x) {
  value <- safe_number(x)
  if (is.na(value) || !is.finite(value)) {
    return(0)
  }
  max(0, min(100, 100 * value))
}

parse_utc_time <- function(x) {
  if (length(x) == 0 || is.null(x)) {
    return(as.POSIXct(NA_real_, origin = "1970-01-01", tz = "UTC"))
  }
  if (inherits(x, "POSIXt")) {
    return(as.POSIXct(x, tz = "UTC"))
  }
  text <- as.character(x)
  text <- sub("Z$", "", text)
  text <- gsub("T", " ", text, fixed = TRUE)
  as.POSIXct(text, format = "%Y-%m-%d %H:%M:%S", tz = "UTC")
}

format_utc_time <- function(x, fallback = "Time pending") {
  time <- parse_utc_time(x)
  time <- time[1]
  if (is.na(time)) {
    fallback
  } else {
    format(time, "%a, %b %d, %I:%M %p %Z", tz = Sys.timezone())
  }
}

prepare_future_board <- function(board) {
  if (nrow(board) == 0) {
    return(board)
  }
  board |>
    dplyr::mutate(
      kickoff_utc = parse_utc_time(kickoff_utc_iso)
    ) |>
    dplyr::filter(
      match_timing != "Completed",
      (
        !is.na(kickoff_utc) & kickoff_utc > Sys.time()
      ) |
        (
          is.na(kickoff_utc) & as.Date(date) >= Sys.Date()
        )
    ) |>
    dplyr::arrange(kickoff_utc, date, source_match_id)
}

local_match_date <- function(kickoff_iso, fallback_date) {
  kickoff <- parse_utc_time(kickoff_iso)
  fallback <- suppressWarnings(as.Date(fallback_date))
  fallback <- rep(fallback, length.out = length(kickoff))
  dates <- rep(as.Date(NA), length(kickoff))
  has_kickoff <- !is.na(kickoff)
  dates[has_kickoff] <- as.Date(format(kickoff[has_kickoff], "%Y-%m-%d", tz = Sys.timezone()))
  needs_fallback <- is.na(dates)
  dates[needs_fallback] <- fallback[needs_fallback]
  dates
}

prepare_daily_board <- function(board, target_date = Sys.Date()) {
  if (nrow(board) == 0) {
    return(board)
  }
  board |>
    dplyr::mutate(
      kickoff_utc = parse_utc_time(kickoff_utc_iso),
      match_day_local = local_match_date(kickoff_utc_iso, date)
    ) |>
    dplyr::filter(match_day_local == target_date) |>
    dplyr::arrange(kickoff_utc, date, source_match_id)
}

summary_value <- function(metric, fallback = "Pending") {
  if (nrow(summary) == 0 || !"metric" %in% names(summary) || !"value" %in% names(summary)) {
    return(fallback)
  }
  value <- summary$value[summary$metric == metric]
  if (length(value) == 0 || is.na(value[[1]]) || !nzchar(as.character(value[[1]]))) {
    fallback
  } else {
    as.character(value[[1]])
  }
}

probability_bar <- function(label, value, class_name) {
  paste0(
    '<div class="prob-row">',
    '<div class="prob-row-label"><span>', escape_html(label), '</span><strong>',
    display_percent(value, 1),
    '</strong></div>',
    '<div class="prob-track"><span class="', class_name, '" style="width:',
    fmt_number(probability_width(value), 1),
    '%"></span></div>',
    '</div>'
  )
}

today_probability_chip <- function(label, value, class_name) {
  paste0(
    '<div class="today-prob-chip">',
    '<span>', escape_html(label), '</span>',
    '<strong>', display_percent(value, 1), '</strong>',
    '<div class="today-prob-track"><i class="', class_name, '" style="width:',
    fmt_number(probability_width(value), 1),
    '%"></i></div>',
    '</div>'
  )
}

team_key <- function(x) {
  x <- iconv(as.character(x), to = "ASCII//TRANSLIT")
  tolower(gsub("[^a-z0-9]", "", x))
}

as_plain_record <- function(row) {
  if (nrow(row) == 0) {
    return(NULL)
  }
  list(
    name = as.character(row$team[[1]]),
    group = as.character(row$group[[1]]),
    seed = paste0(row$group[[1]], row$position[[1]]),
    position = as.integer(row$position[[1]]),
    points = round(as.numeric(row$points[[1]]), 2),
    gd = round(as.numeric(row$gd[[1]]), 2),
    gf = round(as.numeric(row$gf[[1]]), 2),
    elo = round(as.numeric(row$elo[[1]]), 0)
  )
}

build_projected_bracket <- function() {
  if (!requireNamespace("jsonlite", quietly = TRUE)) {
    return(NULL)
  }

  predictions <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "world_cup_2026_fixture_predictions.csv"))
  squads <- read_csv_if_exists(file.path(root, "data", "processed", "public_csv", "dim_2026_world_cup_squad_players.csv"))

  if (nrow(predictions) == 0 || nrow(squads) == 0) {
    return(NULL)
  }

  team_groups <- squads |>
    dplyr::select(team, group_name) |>
    dplyr::distinct() |>
    dplyr::mutate(
      group = gsub("^Group\\s+", "", group_name),
      team_join_key = team_key(team)
    ) |>
    dplyr::select(team_join_key, group)

  predictions <- predictions |>
    dplyr::mutate(
      home_join_key = team_key(home_team),
      away_join_key = team_key(away_team)
    ) |>
    dplyr::left_join(team_groups, by = c("home_join_key" = "team_join_key")) |>
    dplyr::rename(home_group = group) |>
    dplyr::left_join(team_groups, by = c("away_join_key" = "team_join_key")) |>
    dplyr::rename(away_group = group) |>
    dplyr::mutate(
      group = dplyr::coalesce(home_group, away_group),
      completed = !is.na(home_score) & !is.na(away_score),
      home_points = dplyr::case_when(
        completed & home_score > away_score ~ 3,
        completed & home_score == away_score ~ 1,
        completed & home_score < away_score ~ 0,
        TRUE ~ 3 * pred_home_win_prob + pred_draw_prob
      ),
      away_points = dplyr::case_when(
        completed & away_score > home_score ~ 3,
        completed & away_score == home_score ~ 1,
        completed & away_score < home_score ~ 0,
        TRUE ~ 3 * pred_away_win_prob + pred_draw_prob
      ),
      home_goals_for = dplyr::if_else(completed, as.numeric(home_score), pred_home_goals_poisson),
      away_goals_for = dplyr::if_else(completed, as.numeric(away_score), pred_away_goals_poisson),
      home_goals_against = away_goals_for,
      away_goals_against = home_goals_for
    ) |>
    dplyr::filter(!is.na(group))

  home_rows <- data.frame(
    group = predictions$group,
    team = predictions$home_team,
    points = predictions$home_points,
    gf = predictions$home_goals_for,
    ga = predictions$home_goals_against,
    completed = predictions$completed,
    elo = predictions$home_latest_elo
  )
  away_rows <- data.frame(
    group = predictions$group,
    team = predictions$away_team,
    points = predictions$away_points,
    gf = predictions$away_goals_for,
    ga = predictions$away_goals_against,
    completed = predictions$completed,
    elo = predictions$away_latest_elo
  )

  standings <- rbind(home_rows, away_rows) |>
    dplyr::group_by(group, team) |>
    dplyr::summarise(
      matches = dplyr::n(),
      completed = sum(completed, na.rm = TRUE),
      points = sum(points, na.rm = TRUE),
      gf = sum(gf, na.rm = TRUE),
      ga = sum(ga, na.rm = TRUE),
      gd = gf - ga,
      elo = max(elo, na.rm = TRUE),
      .groups = "drop"
    ) |>
    dplyr::arrange(group, dplyr::desc(points), dplyr::desc(gd), dplyr::desc(gf), dplyr::desc(elo), team) |>
    dplyr::group_by(group) |>
    dplyr::mutate(position = dplyr::row_number()) |>
    dplyr::ungroup()

  standings$elo[is.infinite(standings$elo)] <- NA_real_

  third_pool <- standings |>
    dplyr::filter(position == 3) |>
    dplyr::arrange(dplyr::desc(points), dplyr::desc(gd), dplyr::desc(gf), dplyr::desc(elo), team) |>
    dplyr::mutate(third_rank = dplyr::row_number(), qualifies = third_rank <= 8)

  slot_pairs <- data.frame(
    match_id = seq_len(16),
    slot_a = c("A2", "F1", "E1", "I1", "K2", "H1", "D1", "G1", "C1", "E2", "A1", "L1", "J1", "D2", "B1", "K1"),
    slot_b = c("B2", "C2", "ABCDF3", "CDFGH3", "L2", "J2", "BEFIJ3", "AEHIJ3", "F2", "I2", "CEFHI3", "EHIJK3", "H2", "G2", "EFGIJ3", "DEIJL3")
  )

  used_third_teams <- character()
  resolve_slot <- function(slot) {
    if (grepl("^[A-L][12]$", slot)) {
      group <- substr(slot, 1, 1)
      position <- as.integer(substr(slot, 2, 2))
      row <- standings[standings$group == group & standings$position == position, ]
      return(as_plain_record(row))
    }

    if (grepl("3$", slot)) {
      candidate_groups <- strsplit(sub("3$", "", slot), "")[[1]]
      row <- third_pool |>
        dplyr::filter(group %in% candidate_groups, !(team %in% used_third_teams), qualifies) |>
        dplyr::arrange(third_rank) |>
        dplyr::slice_head(n = 1)
      if (nrow(row) > 0) {
        used_third_teams <<- c(used_third_teams, row$team[[1]])
      }
      return(as_plain_record(row))
    }

    NULL
  }

  round32 <- lapply(seq_len(nrow(slot_pairs)), function(i) {
    row <- slot_pairs[i, ]
    list(
      match = as.integer(row$match_id),
      slotA = as.character(row$slot_a),
      slotB = as.character(row$slot_b),
      teamA = resolve_slot(as.character(row$slot_a)),
      teamB = resolve_slot(as.character(row$slot_b))
    )
  })

  group_summary <- standings |>
    dplyr::filter(position <= 3) |>
    dplyr::mutate(
      seed = paste0(group, position),
      points = round(points, 2),
      gd = round(gd, 2),
      gf = round(gf, 2),
      elo = round(elo, 0)
    ) |>
    dplyr::select(group, seed, team, position, points, gd, gf, elo) |>
    as.data.frame()

  third_summary <- third_pool |>
    dplyr::mutate(
      seed = paste0(group, "3"),
      points = round(points, 2),
      gd = round(gd, 2),
      gf = round(gf, 2),
      elo = round(elo, 0)
    ) |>
    dplyr::select(third_rank, qualifies, group, seed, team, points, gd, gf, elo) |>
    as.data.frame()

  list(
    round32 = round32,
    groupSeeds = group_summary,
    thirdWatch = third_summary,
    notes = list(
      seeding = "Projected group tables combine completed results with model expected points for remaining matches.",
      knockout = "Knockout advancement uses the current Elo-style strength signal until knockout match-specific data exists."
    )
  )
}
```

## Prediction Board

Daily match forecasts, upcoming fixtures, model picks, and post-match accuracy for the 2026 World Cup.

<div class="matchday-toolbar">
  <a href="#daily-match-view">Today</a>
  <a href="#next-match-forecast">Next</a>
  <a href="#next-upcoming-matches">Upcoming</a>
  <a href="#interactive-bracket-simulator">Simulator</a>
  <a href="#post-match-accuracy">Accuracy</a>
</div>

## Daily Match View

```{r daily-match-view, echo=FALSE, results="asis"}
today_date <- Sys.Date()
today_board <- prepare_daily_board(board, today_date)
future_board_for_day <- prepare_future_board(board)
last_refresh <- summary_value("last_refreshed_local")
today_label <- format(today_date, "%A, %B %d, %Y")
match_word <- if (nrow(today_board) == 1) "match" else "matches"

cat('<section class="daily-view-shell">')
cat('<div class="daily-view-header">')
cat('<div>')
cat('<div class="section-kicker">Today</div>')
cat('<h2>', escape_html(today_label), '</h2>', sep = "")
cat('<p>Open this section first for the day&apos;s slate, model pick, likely score, goals, cards coverage, and lineup status.</p>')
cat('</div>')
cat('<div class="daily-refresh-card">')
cat('<span>Last refresh</span>')
cat('<strong>', escape_html(last_refresh), '</strong>', sep = "")
cat('<small>Daily refresh is scheduled each morning; match forecasts refresh again before kickoff.</small>')
cat('</div>')
cat('<div class="daily-count-card">')
cat('<strong>', nrow(today_board), '</strong>', sep = "")
cat('<span>', escape_html(match_word), ' today</span>', sep = "")
cat('</div>')
cat('</div>')

if (nrow(today_board) > 0) {
  cat('<div class="today-match-list">')
  for (i in seq_len(nrow(today_board))) {
    row <- today_board[i, ]
    home_label <- safe_text(row$home_team, "Home")
    away_label <- safe_text(row$away_team, "Away")
    winner <- safe_text(row$predicted_winner, "No pick")
    status <- safe_text(row$match_timing, "Scheduled")
    kickoff <- format_utc_time(row$kickoff_utc_iso)
    refresh <- format_utc_time(row$refresh_utc_iso, "Refresh pending")
    actual_result <- safe_text(row$actual_result, "")
    result_line <- if (nzchar(actual_result)) paste("Result:", actual_result) else paste("Status:", status)
    venue <- paste(
      safe_text(row$city, "Venue"),
      safe_text(row$country, "location"),
      sep = ", "
    )
    expected_goals <- paste0(
      display_number(row$pred_home_goals_poisson, 2),
      " - ",
      display_number(row$pred_away_goals_poisson, 2)
    )
    card_projection <- if (
      is.finite(safe_number(row$home_projected_yellow_cards)) &&
        is.finite(safe_number(row$away_projected_yellow_cards))
    ) {
      paste0(
        display_number(row$home_projected_yellow_cards, 1),
        " - ",
        display_number(row$away_projected_yellow_cards, 1)
      )
    } else {
      safe_text(row$yellow_card_model_status, "Pending")
    }
    lineup_state <- paste(
      safe_text(row$home_lineup_status, "Pending"),
      safe_text(row$away_lineup_status, "Pending"),
      sep = " / "
    )

    cat('<article class="today-match-card">')
    cat('<div class="today-card-top">')
    cat('<span>', escape_html(kickoff), '</span>', sep = "")
    cat('<strong>', escape_html(status), '</strong>', sep = "")
    cat('</div>')
    cat('<h3>', escape_html(home_label), ' <span>vs</span> ', escape_html(away_label), '</h3>', sep = "")
    cat('<div class="today-pick-row">')
    cat('<span>Model pick</span>')
    cat('<strong>', escape_html(winner), '</strong>', sep = "")
    cat('<em>', escape_html(safe_text(row$confidence_band, "Confidence")), ': ', display_percent(row$prediction_confidence, 1), '</em>', sep = "")
    cat('</div>')
    cat('<div class="today-prob-grid">')
    cat(today_probability_chip(paste(home_label, "win"), row$ensemble_home_win_prob, "home-prob"))
    cat(today_probability_chip("Draw", row$ensemble_draw_prob, "draw-prob"))
    cat(today_probability_chip(paste(away_label, "win"), row$ensemble_away_win_prob, "away-prob"))
    cat('</div>')
    cat('<div class="today-detail-grid">')
    cat('<div><span>Likely score</span><strong>', escape_html(safe_text(row$most_likely_score, "Pending")), '</strong></div>', sep = "")
    cat('<div><span>Expected goals</span><strong>', escape_html(expected_goals), '</strong></div>', sep = "")
    cat('<div><span>Over 2.5 goals</span><strong>', display_percent(row$over_2_5_prob, 1), '</strong></div>', sep = "")
    cat('<div><span>Both score</span><strong>', display_percent(row$both_teams_to_score_prob, 1), '</strong></div>', sep = "")
    cat('<div><span>Yellow cards</span><strong>', escape_html(card_projection), '</strong></div>', sep = "")
    cat('<div><span>Lineups</span><strong>', escape_html(lineup_state), '</strong></div>', sep = "")
    cat('</div>')
    cat('<div class="today-card-footer">')
    cat('<span>', escape_html(result_line), '</span>', sep = "")
    cat('<span>', escape_html(venue), '</span>', sep = "")
    cat('<span>Forecast refresh: ', escape_html(refresh), '</span>', sep = "")
    cat('</div>')
    cat('</article>')
  }
  cat('</div>')
} else {
  future_with_days <- future_board_for_day |>
    dplyr::mutate(match_day_local = local_match_date(kickoff_utc_iso, date))
  future_with_days <- future_with_days |> dplyr::filter(!is.na(match_day_local))
  cat('<div class="daily-empty-card">')
  cat('<strong>No matches are scheduled for today in the current board.</strong>')
  if (nrow(future_with_days) > 0) {
    next_match <- future_with_days[1, ]
    next_day <- min(future_with_days$match_day_local, na.rm = TRUE)
    next_day_count <- sum(future_with_days$match_day_local == next_day, na.rm = TRUE)
    next_word <- if (next_day_count == 1) "match" else "matches"
    cat('<span>Next match: ', escape_html(safe_text(next_match$home_team, "Home")), ' vs ', escape_html(safe_text(next_match$away_team, "Away")), ' on ', escape_html(format_utc_time(next_match$kickoff_utc_iso)), '.</span>', sep = "")
    cat('<span>Next matchday: ', escape_html(format(next_day, "%A, %B %d")), ' with ', next_day_count, ' ', escape_html(next_word), '.</span>', sep = "")
  }
  cat('</div>')
}

cat('</section>')
```

## Next Match Forecast

```{r next-match-forecast, echo=FALSE, results="asis"}
future_board <- prepare_future_board(board)

if (nrow(future_board) > 0) {
  next_match <- future_board[1, ]
  home_label <- safe_text(next_match$home_team, "Home team")
  away_label <- safe_text(next_match$away_team, "Away team")
  winner <- safe_text(next_match$predicted_winner, "No pick")
  kickoff <- format_utc_time(next_match$kickoff_utc_iso)
  refresh <- format_utc_time(next_match$refresh_utc_iso, "Refresh time pending")
  venue <- paste(
    safe_text(next_match$city, "Venue"),
    safe_text(next_match$country, "location"),
    sep = ", "
  )
  expected_goals <- paste0(
    display_number(next_match$pred_home_goals_poisson, 2),
    " - ",
    display_number(next_match$pred_away_goals_poisson, 2)
  )
  card_projection <- if (
    is.finite(safe_number(next_match$home_projected_yellow_cards)) &&
      is.finite(safe_number(next_match$away_projected_yellow_cards))
  ) {
    paste0(
      display_number(next_match$home_projected_yellow_cards, 1),
      " - ",
      display_number(next_match$away_projected_yellow_cards, 1)
    )
  } else {
    "Feed pending"
  }
  lineup_state <- paste(
    safe_text(next_match$home_lineup_status, "Pending"),
    safe_text(next_match$away_lineup_status, "Pending"),
    sep = " / "
  )

  cat('<section class="next-match-hero">')
  cat('<div class="next-match-main">')
  cat('<div class="section-kicker">Next Match</div>')
  cat('<h2>', escape_html(home_label), ' <span>vs</span> ', escape_html(away_label), '</h2>', sep = "")
  cat('<div class="next-match-meta">')
  cat('<span>', escape_html(kickoff), '</span>', sep = "")
  cat('<span>', escape_html(venue), '</span>', sep = "")
  cat('<span>Forecast refresh: ', escape_html(refresh), '</span>', sep = "")
  cat('</div>')
  cat('<div class="winner-call">')
  cat('<span>Model pick</span>')
  cat('<strong>', escape_html(winner), '</strong>', sep = "")
  cat('<em>', escape_html(safe_text(next_match$confidence_band, "Confidence")), ' confidence: ', display_percent(next_match$prediction_confidence, 1), '</em>', sep = "")
  cat('</div>')
  cat('<div class="probability-panel">')
  cat(probability_bar(paste(home_label, "win"), next_match$ensemble_home_win_prob, "home-prob"))
  cat(probability_bar("Draw", next_match$ensemble_draw_prob, "draw-prob"))
  cat(probability_bar(paste(away_label, "win"), next_match$ensemble_away_win_prob, "away-prob"))
  cat('</div>')
  cat('</div>')

  cat('<div class="next-match-side">')
  cat('<div class="quick-stat"><span>Expected goals</span><strong>', escape_html(expected_goals), '</strong></div>', sep = "")
  cat('<div class="quick-stat"><span>Most likely score</span><strong>', escape_html(safe_text(next_match$most_likely_score, "Pending")), '</strong></div>', sep = "")
  cat('<div class="quick-stat"><span>Over 2.5 goals</span><strong>', display_percent(next_match$over_2_5_prob, 1), '</strong></div>', sep = "")
  cat('<div class="quick-stat"><span>Both teams score</span><strong>', display_percent(next_match$both_teams_to_score_prob, 1), '</strong></div>', sep = "")
  cat('<div class="quick-stat"><span>Yellow cards</span><strong>', escape_html(card_projection), '</strong></div>', sep = "")
  cat('<div class="quick-stat"><span>Lineups</span><strong>', escape_html(lineup_state), '</strong></div>', sep = "")
  cat('</div>')
  cat('</section>')

  cat('<section class="model-vote-panel">')
  cat('<div class="model-vote"><span>OLS</span><strong>', escape_html(safe_text(next_match$ols_predicted_winner, "Pending")), '</strong></div>', sep = "")
  cat('<div class="model-vote"><span>Poisson</span><strong>', escape_html(safe_text(next_match$poisson_predicted_winner, "Pending")), '</strong></div>', sep = "")
  cat('<div class="model-vote"><span>Ordinal</span><strong>', escape_html(safe_text(next_match$ordinal_predicted_winner, "Pending")), '</strong></div>', sep = "")
  cat('<div class="model-vote"><span>Ensemble</span><strong>', escape_html(winner), '</strong></div>', sep = "")
  cat('</section>')
} else {
  cat('<p class="empty-board">No future matches are available in the current local fixture table.</p>')
}
```

## Next Upcoming Matches

```{r upcoming-matches, echo=FALSE, results="asis"}
if (nrow(future_board) > 0) {
  upcoming <- future_board |> dplyr::slice_head(n = 8)
  cat('<div class="upcoming-match-strip">')
  for (i in seq_len(nrow(upcoming))) {
    row <- upcoming[i, ]
    result_probs <- paste0(
      "H ",
      display_percent(row$ensemble_home_win_prob, 0),
      " / D ",
      display_percent(row$ensemble_draw_prob, 0),
      " / A ",
      display_percent(row$ensemble_away_win_prob, 0)
    )
    cat('<article class="upcoming-match-card">')
    cat('<div class="upcoming-time">', escape_html(format_utc_time(row$kickoff_utc_iso)), '</div>', sep = "")
    cat('<h3>', escape_html(safe_text(row$home_team, "Home")), ' vs ', escape_html(safe_text(row$away_team, "Away")), '</h3>', sep = "")
    cat('<div class="upcoming-pick"><span>Pick</span><strong>', escape_html(safe_text(row$predicted_winner, "Pending")), '</strong></div>', sep = "")
    cat('<div class="upcoming-probs">', escape_html(result_probs), '</div>', sep = "")
    cat('<div class="upcoming-score">Likely score: ', escape_html(safe_text(row$most_likely_score, "Pending")), '</div>', sep = "")
    cat('</article>')
  }
  cat('</div>')
} else {
  cat('<p class="empty-board">No upcoming matches are available in the current local fixture table.</p>')
}
```

## Interactive Bracket Simulator

```{r interactive-bracket, echo=FALSE, results="asis"}
bracket_data <- build_projected_bracket()

if (is.null(bracket_data)) {
  cat('<p class="empty-board">The interactive bracket is available after fixture predictions and squad group metadata are refreshed.</p>')
} else {
  bracket_json <- jsonlite::toJSON(bracket_data, auto_unbox = TRUE, na = "null", dataframe = "rows")
  bracket_json <- gsub("<", "\\\\u003c", bracket_json, fixed = TRUE)
  cat('<section class="model-bracket-shell">')
  cat('<div class="model-bracket-header">')
  cat('<div>')
  cat('<div class="section-kicker">Bracket Simulator</div>')
  cat('<h2>Projected Knockout Path</h2>')
  cat('<p>Seeded from current group projections. Click a team to override a match winner; reset returns to the model path.</p>')
  cat('</div>')
  cat('<div class="model-bracket-controls">')
  cat('<button type="button" class="bracket-control" data-bracket-action="model">Model path</button>')
  cat('<button type="button" class="bracket-control" data-bracket-action="reset">Reset picks</button>')
  cat('<button type="button" class="bracket-control" data-bracket-action="layout">Compact view</button>')
  cat('</div>')
  cat('</div>')
  cat('<div class="bracket-overview-grid">')
  cat('<div class="seed-watch-panel"><h3>Projected Group Seeds</h3><div id="group-seed-watch"></div></div>')
  cat('<div class="seed-watch-panel"><h3>Best Third Watch</h3><div id="third-place-watch"></div></div>')
  cat('</div>')
  cat('<div id="model-bracket-root" class="interactive-bracket-grid" aria-live="polite"></div>')
  cat('<div id="model-bracket-champion" class="bracket-champion-card"></div>')
  cat('<script type="application/json" id="wc-bracket-data">', bracket_json, '</script>', sep = "")
  cat('</section>')
}
```

```{=html}
<script>
document.addEventListener("DOMContentLoaded", function () {
  const dataNode = document.getElementById("wc-bracket-data");
  const root = document.getElementById("model-bracket-root");
  const championNode = document.getElementById("model-bracket-champion");
  const seedWatch = document.getElementById("group-seed-watch");
  const thirdWatch = document.getElementById("third-place-watch");

  if (!dataNode || !root || !championNode) {
    return;
  }

  const data = JSON.parse(dataNode.textContent);
  const roundNames = ["Round of 32", "Round of 16", "Quarterfinals", "Semifinals", "Final"];
  const picks = {};

  function cloneTeam(team) {
    return team ? Object.assign({}, team) : null;
  }

  function cloneMatch(match) {
    return {
      slotA: match.slotA,
      slotB: match.slotB,
      teamA: cloneTeam(match.teamA),
      teamB: cloneTeam(match.teamB)
    };
  }

  function teamStrength(team) {
    const elo = team && Number(team.elo);
    return Number.isFinite(elo) ? elo : 1500;
  }

  function advanceProbability(teamA, teamB) {
    const diff = teamStrength(teamA) - teamStrength(teamB);
    return 1 / (1 + Math.pow(10, -diff / 400));
  }

  function formatPct(value) {
    if (!Number.isFinite(value)) {
      return "";
    }
    return Math.round(value * 100) + "%";
  }

  function modelWinner(match) {
    if (!match.teamA && !match.teamB) {
      return null;
    }
    if (match.teamA && !match.teamB) {
      return match.teamA;
    }
    if (!match.teamA && match.teamB) {
      return match.teamB;
    }
    return advanceProbability(match.teamA, match.teamB) >= 0.5 ? match.teamA : match.teamB;
  }

  function clearLaterPicks(roundIndex) {
    Object.keys(picks).forEach(function (key) {
      const keyRound = Number(key.split("-")[0]);
      if (keyRound > roundIndex) {
        delete picks[key];
      }
    });
  }

  function buildRounds() {
    const rounds = [data.round32.map(cloneMatch)];

    for (let roundIndex = 0; roundIndex < 4; roundIndex += 1) {
      const currentRound = rounds[roundIndex];
      const nextRound = Array.from({ length: Math.ceil(currentRound.length / 2) }, function (_, index) {
        return { slotA: "W" + (roundIndex + 1) + "." + (2 * index + 1), slotB: "W" + (roundIndex + 1) + "." + (2 * index + 2), teamA: null, teamB: null };
      });

      currentRound.forEach(function (match, matchIndex) {
        const key = roundIndex + "-" + matchIndex;
        const validNames = [match.teamA && match.teamA.name, match.teamB && match.teamB.name].filter(Boolean);
        if (picks[key] && !validNames.includes(picks[key])) {
          delete picks[key];
        }
        const pickedName = picks[key];
        const winner = pickedName
          ? validNames.includes(pickedName)
            ? (match.teamA && match.teamA.name === pickedName ? match.teamA : match.teamB)
            : modelWinner(match)
          : modelWinner(match);
        const nextMatch = nextRound[Math.floor(matchIndex / 2)];
        if (matchIndex % 2 === 0) {
          nextMatch.teamA = cloneTeam(winner);
        } else {
          nextMatch.teamB = cloneTeam(winner);
        }
      });

      rounds.push(nextRound);
    }

    return rounds;
  }

  function teamButton(team, slot, probability, selected, roundIndex, matchIndex) {
    const button = document.createElement("button");
    button.type = "button";
    button.className = "bracket-team";
    if (selected) {
      button.classList.add("selected");
    }
    if (!team) {
      button.classList.add("empty");
      button.disabled = true;
      button.innerHTML = "<span class='team-name'>" + slot + "</span><small>Pending seed</small>";
      return button;
    }

    button.innerHTML =
      "<span class='team-name'>" + team.name + "</span>" +
      "<span class='team-meta'>" + team.seed + " · " + formatPct(probability) + "</span>";
    button.setAttribute("aria-label", "Pick " + team.name + " to advance");
    button.addEventListener("click", function () {
      picks[roundIndex + "-" + matchIndex] = team.name;
      clearLaterPicks(roundIndex);
      render();
    });
    return button;
  }

  function matchCard(match, roundIndex, matchIndex) {
    const card = document.createElement("article");
    card.className = "bracket-match-card";
    const pA = match.teamA && match.teamB ? advanceProbability(match.teamA, match.teamB) : null;
    const pB = Number.isFinite(pA) ? 1 - pA : null;
    const winner = modelWinner(match);
    const pickedName = picks[roundIndex + "-" + matchIndex];
    const selectedName = pickedName || (winner && winner.name);

    const label = document.createElement("div");
    label.className = "bracket-match-label";
    label.textContent = "Match " + (matchIndex + 1);
    card.appendChild(label);
    card.appendChild(teamButton(match.teamA, match.slotA, pA, match.teamA && match.teamA.name === selectedName, roundIndex, matchIndex));
    card.appendChild(teamButton(match.teamB, match.slotB, pB, match.teamB && match.teamB.name === selectedName, roundIndex, matchIndex));

    if (pickedName) {
      const badge = document.createElement("span");
      badge.className = "user-pick-badge";
      badge.textContent = "Manual pick";
      card.appendChild(badge);
    }

    return card;
  }

  function renderSeeds() {
    if (seedWatch) {
      const seedsByGroup = {};
      data.groupSeeds.forEach(function (row) {
        seedsByGroup[row.group] = seedsByGroup[row.group] || [];
        seedsByGroup[row.group].push(row);
      });
      seedWatch.innerHTML = Object.keys(seedsByGroup).sort().map(function (group) {
        const rows = seedsByGroup[group].slice(0, 2).map(function (row) {
          return "<span><strong>" + row.seed + "</strong>" + row.team + "</span>";
        }).join("");
        return "<div class='seed-group'><b>Group " + group + "</b>" + rows + "</div>";
      }).join("");
    }

    if (thirdWatch) {
      thirdWatch.innerHTML = data.thirdWatch.map(function (row) {
        return "<div class='third-row" + (row.qualifies ? " qualifies" : "") + "'>" +
          "<span>" + row.third_rank + ". " + row.team + "</span>" +
          "<strong>" + row.points + " pts</strong>" +
          "</div>" +
          (row.third_rank === 8 ? "<div class='third-cut'>Cut line</div>" : "");
      }).join("");
    }
  }

  function render() {
    const rounds = buildRounds();
    root.innerHTML = "";
    rounds.forEach(function (round, roundIndex) {
      const column = document.createElement("section");
      column.className = "bracket-round";
      const heading = document.createElement("h3");
      heading.textContent = roundNames[roundIndex];
      column.appendChild(heading);
      const list = document.createElement("div");
      list.className = "bracket-round-list";
      round.forEach(function (match, matchIndex) {
        list.appendChild(matchCard(match, roundIndex, matchIndex));
      });
      column.appendChild(list);
      root.appendChild(column);
    });

    const finalMatch = rounds[rounds.length - 1][0];
    const champion = picks["4-0"]
      ? (finalMatch.teamA && finalMatch.teamA.name === picks["4-0"] ? finalMatch.teamA : finalMatch.teamB)
      : modelWinner(finalMatch);
    championNode.innerHTML = champion
      ? "<span>Projected Champion</span><strong>" + champion.name + "</strong><small>" + champion.seed + " · strength " + champion.elo + "</small>"
      : "";
  }

  document.querySelectorAll("[data-bracket-action]").forEach(function (button) {
    button.addEventListener("click", function () {
      const action = button.getAttribute("data-bracket-action");
      if (action === "layout") {
        root.classList.toggle("is-compact");
        button.textContent = root.classList.contains("is-compact") ? "Horizontal view" : "Compact view";
        return;
      }
      Object.keys(picks).forEach(function (key) {
        delete picks[key];
      });
      render();
    });
  });

  renderSeeds();
  render();
});
</script>
```

## Board Status

```{r summary-cards, echo=FALSE, results="asis"}
summary_value <- function(metric_name) {
  if (nrow(summary) == 0) {
    return(NA_character_)
  }
  value <- summary$value[summary$metric == metric_name]
  if (length(value) == 0) NA_character_ else as.character(value[[1]])
}

if (nrow(summary) > 0) {
  cat(metric_cards(data.frame(
    label = c("Last Refresh", "Matches", "Today", "Upcoming", "Lineups", "Events"),
    value = c(
      summary_value("last_refreshed_local"),
      summary_value("matches_on_board"),
      summary_value("matches_today"),
      summary_value("upcoming_matches"),
      summary_value("lineup_rows_available"),
      summary_value("event_rows_available")
    ),
    note = c(
      "local laptop time",
      "prediction rows",
      "matchday count",
      "future matches",
      "API lineup rows",
      "API event rows"
    )
  )))
}
```

## Future Fixture Cards

Double-click a matchup to open the model detail.

```{r bracket-interface, echo=FALSE, results="asis"}
if (nrow(board) > 0) {
  future_board <- prepare_future_board(board)

  if (nrow(future_board) > 0) {
    cat('<div class="bracket-board">')
    for (day in unique(as.character(future_board$date))) {
      day_rows <- future_board[as.character(future_board$date) == day, ]
      cat('<section class="bracket-day">')
      cat('<h3>', escape_html(format(as.Date(day), "%A, %B %d")), '</h3>', sep = "")
      cat('<div class="match-grid">')

      for (i in seq_len(nrow(day_rows))) {
        row <- day_rows[i, ]
        kickoff <- format_utc_time(row$kickoff_utc_iso)
        result_probs <- paste0(
          "H ",
          display_percent(row$ensemble_home_win_prob, 1),
          " / D ",
          display_percent(row$ensemble_draw_prob, 1),
          " / A ",
          display_percent(row$ensemble_away_win_prob, 1)
        )
        goals <- paste0(
          display_number(row$pred_home_goals_poisson, 2),
          " - ",
          display_number(row$pred_away_goals_poisson, 2)
        )
        cards <- if (
          is.finite(safe_number(row$home_projected_yellow_cards)) &&
            is.finite(safe_number(row$away_projected_yellow_cards))
        ) {
          paste0(
            display_number(row$home_projected_yellow_cards, 1),
            " - ",
            display_number(row$away_projected_yellow_cards, 1)
          )
        } else {
          "Feed pending"
        }
        lineups <- paste0(row$home_lineup_status, " / ", row$away_lineup_status)
        refresh <- paste("Refresh 30 min before:", format_utc_time(row$refresh_utc_iso, "time pending"))

        cat('<article class="match-card" tabindex="0">')
        cat('<div class="match-card-top">')
        cat('<span>', escape_html(kickoff), '</span>', sep = "")
        cat('<strong>', escape_html(row$confidence_band), '</strong>', sep = "")
        cat('</div>')
        cat('<div class="match-teams">')
        cat('<div><span>', escape_html(row$home_team), '</span><small>Home/listed</small></div>', sep = "")
        cat('<div class="versus">vs</div>')
        cat('<div><span>', escape_html(row$away_team), '</span><small>Away/listed</small></div>', sep = "")
        cat('</div>')
        cat('<div class="pick-line">')
        cat('<span>Winner</span><strong>', escape_html(row$predicted_winner), '</strong>', sep = "")
        cat('</div>')
        cat('<div class="model-picks">')
        cat('<span>OLS: ', escape_html(row$ols_predicted_winner), '</span>', sep = "")
        cat('<span>Poisson: ', escape_html(row$poisson_predicted_winner), '</span>', sep = "")
        cat('<span>Ordinal: ', escape_html(row$ordinal_predicted_winner), '</span>', sep = "")
        cat('</div>')
        cat('<div class="match-detail">')
        cat('<dl>')
        cat('<dt>Outcome probabilities</dt><dd>', escape_html(result_probs), '</dd>', sep = "")
        cat('<dt>Expected goals</dt><dd>', escape_html(goals), '</dd>', sep = "")
        cat('<dt>Most likely score</dt><dd>', escape_html(row$most_likely_score), '</dd>', sep = "")
        cat('<dt>Totals</dt><dd>O2.5 ', display_percent(row$over_2_5_prob, 1), ' / BTTS ', display_percent(row$both_teams_to_score_prob, 1), '</dd>', sep = "")
        cat('<dt>Yellow cards</dt><dd>', escape_html(cards), '</dd>', sep = "")
        cat('<dt>Lineups</dt><dd>', escape_html(lineups), '</dd>', sep = "")
        cat('<dt>Venue</dt><dd>', escape_html(paste(row$city, row$country, sep = ", ")), '</dd>', sep = "")
        cat('<dt>Refresh</dt><dd>', escape_html(refresh), '</dd>', sep = "")
        cat('</dl>')
        cat('</div>')
        cat('</article>')
      }

      cat('</div>')
      cat('</section>')
    }
    cat('</div>')
  } else {
    cat('<p class="empty-board">No future matches are available in the current local fixture table.</p>')
  }
} else {
  cat('<p class="empty-board">Matchday predictions are available after the matchday board refresh.</p>')
}
```

```{=html}
<script>
document.addEventListener("DOMContentLoaded", function () {
  document.querySelectorAll(".match-card").forEach(function (card) {
    card.addEventListener("dblclick", function () {
      card.classList.toggle("open");
    });
    card.addEventListener("keydown", function (event) {
      if (event.key === "Enter" || event.key === " ") {
        event.preventDefault();
        card.classList.toggle("open");
      }
    });
  });
});
</script>
```

```{r board-table}
if (nrow(board) > 0) {
  board_for_display <- prepare_future_board(board)

  display_board <- board_for_display |>
    dplyr::mutate(
      result_probability = paste0(
        "H ",
        fmt_percent(ensemble_home_win_prob, 1),
        " / D ",
        fmt_percent(ensemble_draw_prob, 1),
        " / A ",
        fmt_percent(ensemble_away_win_prob, 1)
      ),
      expected_goals = paste0(
        fmt_number(pred_home_goals_poisson, 2),
        " - ",
        fmt_number(pred_away_goals_poisson, 2)
      ),
      totals = paste0(
        "O2.5 ",
        fmt_percent(over_2_5_prob, 1),
        " / BTTS ",
        fmt_percent(both_teams_to_score_prob, 1)
      ),
      yellow_cards = dplyr::case_when(
        is.finite(home_projected_yellow_cards) & is.finite(away_projected_yellow_cards) ~
          paste0(
            fmt_number(home_projected_yellow_cards, 1),
            " - ",
            fmt_number(away_projected_yellow_cards, 1)
          ),
        TRUE ~ "Feed pending"
      ),
      lineup = paste0(home_lineup_status, " / ", away_lineup_status),
      confidence = paste0(confidence_band, " (", fmt_percent(prediction_confidence, 1), ")"),
      venue = paste(city, country, sep = ", ")
    ) |>
    dplyr::select(
      date,
      match_timing,
      match_label,
      predicted_winner,
      confidence,
      result_probability,
      expected_goals,
      most_likely_score,
      totals,
      yellow_cards,
      lineup,
      venue
    )

  if (nrow(display_board) > 0) {
    display_board |>
      display_table(caption = "Future matchday predictions.", max_rows = 24)
  } else {
    display_table(empty_note("No future matches are available in the current rendered board."))
  }
} else {
  display_table(empty_note("Matchday predictions are available after the matchday board refresh."))
}
```

## Lineup Detail

Confirmed lineups appear here when the match feed posts starting XIs.

```{r lineup-detail}
if (nrow(board) > 0) {
  lineup_detail <- board |>
    dplyr::filter(home_lineup_status == "Confirmed" | away_lineup_status == "Confirmed") |>
    dplyr::select(
      date,
      match_label,
      home_formation,
      away_formation,
      home_starting_lineup,
      away_starting_lineup
    )

  if (nrow(lineup_detail) > 0) {
    lineup_detail |>
      display_table(caption = "Confirmed starting lineups.", max_rows = 8)
  } else {
    display_table(empty_note("Starting lineups update when the fixture feed posts them."))
  }
} else {
  display_table(empty_note("Lineup detail is available after the matchday board refresh."))
}
```

## Post-Match Accuracy

Each completed match is graded against the final score. "Right" means the model's top result matched the actual result. Goal miss is the absolute difference between predicted total goals and actual total goals.

```{r accuracy, results='asis'}
accuracy <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "matchday_model_accuracy.csv"))
accuracy_detail <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "matchday_model_accuracy_detail.csv"))

accuracy_badge <- function(correct) {
  value <- as.logical(correct)[[1]]
  if (is.na(value)) {
    return('<span class="accuracy-badge accuracy-pending">Pending</span>')
  }
  if (value) {
    '<span class="accuracy-badge accuracy-correct">Right</span>'
  } else {
    '<span class="accuracy-badge accuracy-wrong">Wrong</span>'
  }
}

goal_error_chip <- function(value) {
  numeric_value <- safe_number(value)
  if (is.na(numeric_value) || !is.finite(numeric_value)) {
    return('<span class="error-chip error-neutral">No goal model</span>')
  }
  class_name <- if (numeric_value <= 0.5) {
    "error-good"
  } else if (numeric_value <= 1.25) {
    "error-warn"
  } else {
    "error-bad"
  }
  paste0(
    '<span class="error-chip ', class_name, '">',
    display_number(numeric_value, 2),
    ' goal miss</span>'
  )
}

probability_line <- function(value) {
  numeric_value <- safe_number(value)
  if (is.na(numeric_value) || !is.finite(numeric_value)) {
    return("")
  }
  paste0('<small>Actual result prob: ', display_percent(numeric_value, 1), '</small>')
}

model_cell <- function(row, pick_col, correct_col, prob_col = NULL, error_col = NULL) {
  pick <- safe_text(row[[pick_col]], "No pick")
  pieces <- c(
    paste0('<strong>', escape_html(pick), '</strong>'),
    accuracy_badge(row[[correct_col]])
  )
  if (!is.null(prob_col)) {
    pieces <- c(pieces, probability_line(row[[prob_col]]))
  }
  if (!is.null(error_col)) {
    pieces <- c(pieces, goal_error_chip(row[[error_col]]))
  }
  paste0('<div class="accuracy-model-cell">', paste(pieces[nzchar(pieces)], collapse = ""), '</div>')
}

if (nrow(accuracy_detail) > 0) {
  accuracy_detail <- accuracy_detail |>
    dplyr::arrange(dplyr::desc(as.Date(date)), dplyr::desc(source_match_id))

  cards <- data.frame(
    label = c(
      "Completed Matches",
      "Ensemble Right",
      "Poisson Goal Miss",
      "Average Actual-Result Probability"
    ),
    value = c(
      fmt_integer(nrow(accuracy_detail)),
      fmt_percent(mean(accuracy_detail$ensemble_correct, na.rm = TRUE), 1),
      paste0(fmt_number(mean(accuracy_detail$poisson_total_goal_error, na.rm = TRUE), 2), " goals"),
      fmt_percent(mean(accuracy_detail$ensemble_probability_actual, na.rm = TRUE), 1)
    ),
    note = c(
      "Matches with final scores available.",
      "Share of completed matches the combined model called correctly.",
      "Average absolute miss on total goals.",
      "Average confidence assigned to what actually happened."
    ),
    stringsAsFactors = FALSE
  )
  cat(metric_cards(cards))

  display_rows <- head(accuracy_detail, 40)
  table_rows <- vapply(seq_len(nrow(display_rows)), function(i) {
    row <- display_rows[i, , drop = FALSE]
    paste0(
      '<tr>',
      '<td class="accuracy-match-cell"><strong>', escape_html(row$match_label), '</strong><small>',
      escape_html(row$date), '</small></td>',
      '<td class="accuracy-actual-cell"><strong>', escape_html(row$actual_score), '</strong><span>',
      escape_html(row$actual_winner), '</span><small>', escape_html(row$actual_result), '</small></td>',
      '<td>', model_cell(row, "ensemble_pick", "ensemble_correct", "ensemble_probability_actual"), '</td>',
      '<td>', model_cell(row, "ols_pick", "ols_correct", NULL, "ols_total_goal_error"), '</td>',
      '<td>', model_cell(row, "poisson_pick", "poisson_correct", "poisson_probability_actual", "poisson_total_goal_error"), '</td>',
      '<td>', model_cell(row, "ordinal_pick", "ordinal_correct", "ordinal_probability_actual"), '</td>',
      '</tr>'
    )
  }, character(1))

  cat(
    '<div class="accuracy-table-wrap">',
    '<table class="accuracy-table">',
    '<caption>Completed-match model review. The table shows up to the 40 most recent completed matches.</caption>',
    '<thead><tr>',
    '<th>Match</th>',
    '<th>Actual</th>',
    '<th>Ensemble</th>',
    '<th>OLS Goals</th>',
    '<th>Poisson</th>',
    '<th>Ordinal</th>',
    '</tr></thead>',
    '<tbody>',
    paste(table_rows, collapse = ""),
    '</tbody></table></div>',
    sep = ""
  )

  if (nrow(accuracy) > 0) {
    summary_rows <- accuracy |>
      dplyr::mutate(
        completed_matches = fmt_integer(completed_matches),
        outcome_accuracy = fmt_percent(outcome_accuracy, 1),
        avg_total_goal_error = dplyr::if_else(
          is.na(avg_total_goal_error),
          "Not a goal model",
          paste0(fmt_number(avg_total_goal_error, 2), " goals")
        ),
        avg_team_goal_mae = dplyr::if_else(
          is.na(avg_team_goal_mae),
          "Not a goal model",
          paste0(fmt_number(avg_team_goal_mae, 2), " goals")
        )
      )
    print(display_table(summary_rows, caption = "Model-level accuracy summary."))
  }
} else {
  display_table(empty_note("Post-match accuracy updates after completed scores are available."))
}
```

## Models And Data Used

```{r data-sources}
sources <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "matchday_prediction_data_sources.csv"))

if (nrow(sources) > 0) {
  sources |>
    display_table(caption = "Prediction outputs, model source, and current role.")
} else {
  display_table(empty_note("Model-source notes are available after the matchday board refresh."))
}
```

The board uses three active layers:

- **Winner and outcome:** an ensemble of the ordinal logistic result model and probabilities derived from the Poisson score grid.
- **Goals and likely score:** the Poisson goals model produces expected goals; the score grid turns those expected goals into scoreline probabilities.
- **Totals:** over 2.5 goals and both-teams-to-score probabilities come from the same Poisson score grid.

Cards and lineups are live-enrichment layers. They update when the football API returns player card history, event rows, or confirmed lineups for a fixture.

## How To Read It

- **Predicted winner:** highest ensemble probability. A draw remains possible for group-stage style matches.
- **Confidence:** the model's top probability bucket, not a guarantee.
- **Expected goals:** model average score, not the most likely exact score.
- **Most likely score:** the single scoreline with the highest probability in the Poisson grid.
- **O2.5:** probability that total goals are 3 or more.
- **BTTS:** probability that both teams score at least one goal.

## Reproduce Locally

```{r reproduce, eval=FALSE}
source("R/16_score_2026_fixtures.R")
source("R/17_matchday_prediction_board.R")
```
