---
title: "Multi-Sport Model Expansion"
subtitle: "API-Sports inventory, sport sections, and reusable modeling plan"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
root <- if (file.exists("data/processed")) "." else ".."
source(file.path(root, "R", "report_helpers.R"))

read_csv_if_exists <- function(path) {
  if (file.exists(path)) {
    readr::read_csv(path, show_col_types = FALSE)
  } else {
    data.frame()
  }
}

slug <- function(x) {
  x <- iconv(as.character(x), to = "ASCII//TRANSLIT")
  tolower(gsub("[^a-z0-9]+", "-", x))
}

escape_html <- function(x) {
  x <- ifelse(is.na(x), "", as.character(x))
  x <- gsub("&", "&amp;", x, fixed = TRUE)
  x <- gsub("<", "&lt;", x, fixed = TRUE)
  x <- gsub(">", "&gt;", x, fixed = TRUE)
  x <- gsub('"', "&quot;", x, fixed = TRUE)
  x
}

inventory_path <- file.path(root, "data", "processed", "multisport", "apisports_sport_inventory.csv")
endpoint_path <- file.path(root, "data", "processed", "multisport", "apisports_endpoint_inventory.csv")
dictionary_path <- file.path(root, "data", "processed", "multisport", "apisports_field_dictionary.csv")
training_summary_path <- file.path(root, "data", "processed", "multisport", "apisports_training_pull_summary.csv")
training_log_path <- file.path(root, "data", "processed", "multisport", "apisports_training_request_log.csv")
training_sample_path <- file.path(root, "data", "processed", "multisport", "apisports_training_sample_records.csv")
config_path <- file.path(root, "config", "sports.json")

sport_inventory <- read_csv_if_exists(inventory_path)
endpoint_inventory <- read_csv_if_exists(endpoint_path)
field_dictionary <- read_csv_if_exists(dictionary_path)
training_summary <- read_csv_if_exists(training_summary_path)
training_log <- read_csv_if_exists(training_log_path)
training_samples <- read_csv_if_exists(training_sample_path)

if (nrow(sport_inventory) == 0 && file.exists(config_path)) {
  config <- jsonlite::fromJSON(config_path)
  sport_inventory <- config$sports |>
    dplyr::transmute(
      sport_key,
      sport_label = label,
      host,
      api_version,
      historical_depth_note,
      reference_priority,
      model_family,
      endpoints_configured = vapply(inventory_endpoints, nrow, numeric(1)),
      endpoints_ok = NA_real_,
      plan = "Pending pull",
      subscription_active = NA,
      subscription_end = NA,
      requests_current = NA,
      requests_limit_day = config$request_budget_policy$default_plan_limit_per_day
    )
}

sport_inventory <- sport_inventory |>
  dplyr::arrange(as.numeric(reference_priority), sport_label)

pulled_at <- if ("pulled_at_utc" %in% names(sport_inventory) && nrow(sport_inventory) > 0) {
  sport_inventory$pulled_at_utc[[1]]
} else {
  "Inventory pending"
}
```

## Sport Inventory

This page organizes every sport covered by the current API-Sports setup. It uses a common metadata layer first, then adds sport-specific modeling only after the raw, cleaned, and model-ready datasets are reliable.

```{r inventory-summary, echo=FALSE, results="asis"}
if (nrow(sport_inventory) > 0) {
  endpoints_ok <- suppressWarnings(sum(as.numeric(sport_inventory$endpoints_ok), na.rm = TRUE))
  endpoints_configured <- suppressWarnings(sum(as.numeric(sport_inventory$endpoints_configured), na.rm = TRUE))
  limit_day <- suppressWarnings(max(as.numeric(sport_inventory$requests_limit_day), na.rm = TRUE))
  if (!is.finite(limit_day)) {
    limit_day <- NA_real_
  }
  cat(metric_cards(data.frame(
    label = c("Sports", "Endpoint Checks", "Plan Limit", "Last Inventory"),
    value = c(
      fmt_integer(nrow(sport_inventory)),
      paste0(fmt_integer(endpoints_ok), " / ", fmt_integer(endpoints_configured)),
      ifelse(is.na(limit_day), "Pending", paste0(fmt_integer(limit_day), " per day")),
      pulled_at
    ),
    note = c(
      "configured sports",
      "successful metadata calls",
      "reported daily request limit",
      "UTC timestamp"
    )
  )))
}
```

```{r sport-selector, echo=FALSE, results="asis"}
if (nrow(sport_inventory) > 0) {
  cat('<div class="sport-jump">')
  cat('<label for="sport-select">Jump to sport</label>')
  cat('<select id="sport-select">')
  cat('<option value="">Choose a sport</option>')
  for (i in seq_len(nrow(sport_inventory))) {
    row <- sport_inventory[i, ]
    target <- paste0("#sport-", slug(row$sport_key))
    cat('<option value="', escape_html(target), '">', escape_html(row$sport_label), '</option>', sep = "")
  }
  cat('</select>')
  cat('</div>')
  cat('<script>')
  cat('document.addEventListener("DOMContentLoaded",function(){var s=document.getElementById("sport-select");if(!s)return;s.addEventListener("change",function(){if(s.value){document.querySelector(s.value)?.scrollIntoView({behavior:"smooth",block:"start"});}});});')
  cat('</script>')
}
```

```{r sport-table, echo=FALSE}
if (nrow(sport_inventory) > 0) {
  sport_inventory |>
    dplyr::select(
      sport_label,
      host,
      historical_depth_note,
      model_family,
      endpoints_configured,
      endpoints_ok,
      plan,
      requests_limit_day
    ) |>
    display_table(caption = "Configured API-Sports coverage by sport.", max_rows = 20)
} else {
  display_table(empty_note("Run the API-Sports inventory pull to populate this table."))
}
```

## Training Data Pull

The first multi-sport pull stores full raw API responses locally, then publishes only compact summaries and small samples here. These files are the starting point for cleaned tables and model-ready datasets.

```{r training-summary, echo=FALSE, results="asis"}
if (nrow(training_summary) > 0) {
  total_requests <- sum(as.numeric(training_summary$requests_attempted), na.rm = TRUE)
  total_ok <- sum(as.numeric(training_summary$requests_ok), na.rm = TRUE)
  total_results <- sum(as.numeric(training_summary$total_results_reported), na.rm = TRUE)
  total_errors <- sum(as.numeric(training_summary$requests_with_errors), na.rm = TRUE)
  cat(metric_cards(data.frame(
    label = c("API Requests", "Successful Calls", "Reported Rows", "Calls With Notes"),
    value = c(
      fmt_integer(total_requests),
      fmt_integer(total_ok),
      fmt_integer(total_results),
      fmt_integer(total_errors)
    ),
    note = c(
      "historical pull",
      "HTTP 200 responses",
      "provider reported results",
      "plan, empty, or endpoint notes"
    )
  )))
}
```

```{r training-summary-table, echo=FALSE}
if (nrow(training_summary) > 0) {
  training_summary |>
    dplyr::select(
      sport_label,
      requests_attempted,
      requests_ok,
      requests_with_errors,
      total_results_reported,
      request_cap
    ) |>
    display_table(caption = "Historical API-Sports data pull by sport.", max_rows = 20)
} else {
  display_table(empty_note("Training data pulls have not run yet."))
}
```

```{r training-sample-table, echo=FALSE}
if (nrow(training_samples) > 0) {
  training_samples |>
    dplyr::select(
      sport_label,
      endpoint_name,
      path,
      results,
      sample_top_level_fields
    ) |>
    display_table(caption = "Representative sample fields from downloaded training data.", max_rows = 30)
} else {
  display_table(empty_note("Sample records appear after the training data pull runs."))
}
```

```{r training-errors, echo=FALSE}
if (nrow(training_log) > 0) {
  notable <- training_log |>
    dplyr::filter(status_code != "200" | (!is.na(errors) & nzchar(errors))) |>
    dplyr::select(sport_label, endpoint_name, status_code, results, errors) |>
    dplyr::slice_head(n = 12)
  if (nrow(notable) > 0) {
    display_table(notable, caption = "First endpoint notes to review before the next pull.", max_rows = 12)
  } else {
    display_table(empty_note("No endpoint notes were recorded in the latest training pull."))
  }
} else {
  display_table(empty_note("Endpoint notes appear after the training data pull runs."))
}
```

## Sport Sections

```{r sport-sections, echo=FALSE, results="asis"}
if (nrow(sport_inventory) == 0) {
  cat('<p class="empty-board">No sport inventory is available yet.</p>')
} else {
  for (i in seq_len(nrow(sport_inventory))) {
    sport <- sport_inventory[i, ]
    sport_key <- as.character(sport$sport_key)
    sport_id <- paste0("sport-", slug(sport_key))
    sport_endpoints <- endpoint_inventory |>
      dplyr::filter(.data$sport_key == sport_key) |>
      dplyr::select(endpoint_name, path, status_code, results, errors, sample_fields) |>
      dplyr::slice_head(n = 8)
    sport_fields <- field_dictionary |>
      dplyr::filter(.data$sport_key == sport_key) |>
      dplyr::select(field_name, definition, source, sample_value) |>
      dplyr::slice_head(n = 6)

    cat('<section class="sport-section-card" id="', escape_html(sport_id), '">', sep = "")
    cat('<div class="sport-section-header">')
    cat('<div>')
    cat('<div class="section-kicker">Sport</div>')
    cat('<h3>', escape_html(sport$sport_label), '</h3>', sep = "")
    cat('<p>', escape_html(sport$historical_depth_note), '</p>', sep = "")
    cat('</div>')
    cat('<div class="sport-section-meta">')
    cat('<span>Host</span><strong>', escape_html(sport$host), '</strong>', sep = "")
    cat('<span>Model family</span><strong>', escape_html(sport$model_family), '</strong>', sep = "")
    cat('</div>')
    cat('</div>')

    cat('<div class="sport-section-grid">')
    cat('<div class="sport-plan-card"><span>Endpoint checks</span><strong>', escape_html(sport$endpoints_ok), ' / ', escape_html(sport$endpoints_configured), '</strong></div>', sep = "")
    cat('<div class="sport-plan-card"><span>Plan</span><strong>', escape_html(sport$plan), '</strong></div>', sep = "")
    cat('<div class="sport-plan-card"><span>Daily limit</span><strong>', escape_html(sport$requests_limit_day), '</strong></div>', sep = "")
    cat('</div>')

    cat('<div class="sport-sample-block">')
    cat('<h4>Endpoint Sample</h4>')
    if (nrow(sport_endpoints) > 0) {
      print(display_table(sport_endpoints, max_rows = 8))
    } else {
      print(display_table(empty_note("Endpoint sample appears after the inventory pull.")))
    }
    cat('</div>')

    cat('<div class="sport-sample-block">')
    cat('<h4>Metadata Sample</h4>')
    if (nrow(sport_fields) > 0) {
      print(display_table(sport_fields, max_rows = 6))
    } else {
      print(display_table(empty_note("Field dictionary sample appears after the inventory pull.")))
    }
    cat('</div>')
    cat('</section>')
  }
}
```

## Reference Implementation

Basketball is the recommended first non-soccer reference implementation. It has a clean team-versus-team event structure, no draw outcome in most competitions, and enough overlap with the World Cup workflow to prove the reusable ingestion, validation, modeling, and reporting layers.

The first production pass for each sport will use the same sequence:

1. Ingest raw API snapshots with timestamped manifests.
2. Build cleaned game, team, player, standings, odds, and statistics tables where available.
3. Create model-ready rows with one row per team-game or event participant.
4. Run validation tests before modeling.
5. Fit baseline models first, then Bayesian models where the sport has enough historical depth.
6. Backtest by time, compare calibration, and publish small result samples.

## Current Lineage Rules

- Raw JSON stays local in `data/raw/apisports_inventory/`.
- Sanitized summaries are written to `data/processed/multisport/`.
- Published pages show samples, endpoint status, timestamps, and field definitions.
- The public site shows compact summaries only; full raw feeds remain in local project storage.
