---
title: "Model Challengers"
subtitle: "Stepwise regression, XGBoost readiness, and tree-based goal models."
description: "Challenger model comparison for the World Cup forecasting system, including stepwise regression, XGBoost readiness, random forest fallback, and feature importance."
toc: true
number-sections: true
code-fold: true
code-tools: true
execute:
  echo: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
root <- if (file.exists("data/processed/modeling")) "." else ".."
source(file.path(root, "R", "report_helpers.R"))
read_csv_if_exists <- function(path) {
  if (file.exists(path)) {
    readr::read_csv(path, show_col_types = FALSE)
  } else {
    data.frame()
  }
}
challengers <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "model_challenger_metrics.csv"))
importance <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "model_challenger_feature_importance.csv"))
status <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "model_challenger_status.csv"))
predictions <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "model_challenger_prediction_sample.csv"))
```

## Purpose

The production forecast should not change just because a model sounds more advanced.
This page compares challenger models against the current regression stack before any
promotion into the public prediction cards.

Current challengers:

- **Stepwise OLS:** an interpretable regression that selects terms by AIC.
- **XGBoost goals:** a gradient-boosted count model when the local `xgboost`
  package is installed.
- **Random forest goals:** a free local tree fallback while XGBoost is not
  installed.

## Challenger Leaderboard

```{r leaderboard, echo=FALSE, results="asis"}
if (nrow(challengers) > 0) {
  cards <- challengers |>
    dplyr::arrange(test_rmse) |>
    dplyr::mutate(
      label = model,
      value = paste0(fmt_number(test_rmse, 3), " RMSE"),
      note = paste0(model_family, " | MAE ", fmt_number(test_mae, 3))
    ) |>
    dplyr::select(label, value, note)
  cat(metric_cards(cards))
} else {
  display_table(empty_note("Challenger metrics are available after running R/18_fit_model_challengers.R."))
}
```

```{r metric-table}
if (nrow(challengers) > 0) {
  challengers |>
    dplyr::arrange(test_rmse) |>
    dplyr::mutate(
      test_rmse = fmt_number(test_rmse, 3),
      test_mae = fmt_number(test_mae, 3),
      mean_prediction = fmt_number(mean_prediction, 3),
      rows_train = fmt_integer(rows_train),
      rows_test = fmt_integer(rows_test)
    ) |>
    display_table(caption = "Held-out challenger model performance.")
} else {
  display_table(empty_note("No challenger metrics are available yet."))
}
```

## RMSE Visual

```{r rmse-plot, fig.cap="Held-out RMSE for the challenger goal models. Lower is better."}
if (nrow(challengers) > 0 && requireNamespace("ggplot2", quietly = TRUE)) {
  plot_data <- challengers |>
    dplyr::arrange(test_rmse) |>
    dplyr::mutate(model = factor(model, levels = rev(model)))

  ggplot2::ggplot(plot_data, ggplot2::aes(x = test_rmse, y = model, fill = model_family)) +
    ggplot2::geom_col(width = 0.64, show.legend = FALSE) +
    ggplot2::geom_text(
      ggplot2::aes(label = round(test_rmse, 3)),
      hjust = -0.12,
      color = "#24313d",
      size = 3.7
    ) +
    ggplot2::scale_fill_manual(values = c(
      "Interpretable regression" = "#0f6b68",
      "Gradient-boosted trees" = "#d9a441",
      "Tree ensemble fallback" = "#3d5873"
    )) +
    ggplot2::coord_cartesian(xlim = c(0, max(plot_data$test_rmse, na.rm = TRUE) * 1.14)) +
    ggplot2::labs(
      title = "Goal model challenger leaderboard",
      x = "RMSE",
      y = NULL
    ) +
    ggplot2::theme_minimal(base_size = 12)
} else {
  plot.new()
  text(0.5, 0.5, "RMSE plot is available after challenger metrics are refreshed.")
}
```

## Feature Importance

```{r feature-importance-plot, fig.cap="Top feature importance by challenger model. Importance is scaled within each model."}
if (nrow(importance) > 0 && requireNamespace("ggplot2", quietly = TRUE)) {
  top_importance <- importance |>
    dplyr::group_by(model) |>
    dplyr::slice_max(importance_scaled, n = 6, with_ties = FALSE) |>
    dplyr::ungroup() |>
    dplyr::mutate(feature = gsub("TRUE$", "", feature))

  ggplot2::ggplot(top_importance, ggplot2::aes(x = importance_scaled, y = reorder(feature, importance_scaled))) +
    ggplot2::geom_col(fill = "#0f6b68", width = 0.68) +
    ggplot2::facet_wrap(~ model, scales = "free_y") +
    ggplot2::scale_x_continuous(labels = function(x) paste0(round(100 * x), "%")) +
    ggplot2::labs(
      title = "Top features by challenger",
      x = "Scaled importance",
      y = NULL
    ) +
    ggplot2::theme_minimal(base_size = 12)
} else {
  plot.new()
  text(0.5, 0.5, "Feature importance is available after challenger models are fit.")
}
```

```{r feature-importance-table}
if (nrow(importance) > 0) {
  importance |>
    dplyr::group_by(model) |>
    dplyr::slice_max(importance_scaled, n = 8, with_ties = FALSE) |>
    dplyr::ungroup() |>
    dplyr::mutate(
      importance = fmt_number(importance, 3),
      importance_scaled = fmt_percent(importance_scaled, 1)
    ) |>
    display_table(caption = "Top model-specific feature importance.")
} else {
  display_table(empty_note("Feature importance is available after challenger models are fit."))
}
```

## XGBoost Status

```{r xgboost-status}
if (nrow(status) > 0) {
  status |>
    display_table(caption = "Local challenger package status.")
} else {
  display_table(empty_note("Package status is available after the challenger run."))
}
```

## Prediction Sample

```{r prediction-sample}
if (nrow(predictions) > 0) {
  predictions |>
    dplyr::arrange(model, dplyr::desc(abs(residual))) |>
    dplyr::mutate(
      predicted_goals = fmt_number(predicted_goals, 2),
      residual = fmt_number(residual, 2)
    ) |>
    dplyr::select(model, date, team, opponent, actual_goals, predicted_goals, residual) |>
    display_table(caption = "Largest held-out challenger residuals.", max_rows = 20)
} else {
  display_table(empty_note("Prediction samples are available after challenger models are fit."))
}
```

## Interpretation

The challenger models are useful for two separate reasons:

- Stepwise regression helps simplify the equation while staying interpretable.
- Tree models can capture nonlinear interactions without hand-writing every term.

The tree results should be treated as a candidate signal, not an automatic upgrade.
Before promotion, the next pass should compare calibration, tournament backtesting,
and stability across multiple time splits.

## Reproduce Locally

```{r reproduce, eval=FALSE}
source("R/18_fit_model_challengers.R")
```
