---
title: "Model Challengers"
subtitle: "Expanded historical model comparison for goals and match results."
description: "Challenger model comparison for the World Cup forecasting system, including count models, smooth models, ordinal and multinomial classifiers, stepwise regression, tree models, and feature importance."
toc: true
number-sections: true
code-fold: true
code-tools: true
execute:
  echo: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
root <- if (file.exists("data/processed/modeling")) "." else ".."
source(file.path(root, "R", "report_helpers.R"))
read_csv_if_exists <- function(path) {
  if (file.exists(path)) {
    readr::read_csv(path, show_col_types = FALSE)
  } else {
    data.frame()
  }
}
challengers <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "model_challenger_metrics.csv"))
importance <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "model_challenger_feature_importance.csv"))
status <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "model_challenger_status.csv"))
predictions <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "model_challenger_prediction_sample.csv"))
expanded_goals <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "expanded_goal_model_metrics.csv"))
expanded_results <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "expanded_result_model_metrics.csv"))
expanded_status <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "expanded_model_suite_status.csv"))
population_goals <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "population_expansion_goal_metrics.csv"))
population_results <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "population_expansion_result_metrics.csv"))
population_status <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "population_expansion_status.csv"))
population_summary <- read_csv_if_exists(file.path(root, "data", "processed", "modeling", "expanded_population_summary.csv"))

validation_label <- function(x) {
  dplyr::case_when(
    x == "random_80_20" ~ "Random 80/20 split",
    x == "train_through_2018_test_2019_plus" ~ "Train through 2018, test 2019+",
    x == "all_history_random_80_20" ~ "All history random 80/20 split",
    x == "all_history_time_2019_plus" ~ "All history train through 2018, test 2019+",
    x == "modern_1990_time_2019_plus" ~ "Modern era train 1990-2018, test 2019+",
    x == "recent_2000_time_2019_plus" ~ "Recent era train 2000-2018, test 2019+",
    TRUE ~ x
  )
}
```

## Purpose

The production forecast should not change just because a model sounds more advanced.
This page compares challenger models against the current regression stack before any
promotion into the public prediction cards.

Current challengers:

- **Stepwise OLS:** an interpretable regression that selects terms by AIC.
- **Count models:** Poisson, quasi-Poisson, negative binomial, and two-stage
  zero-aware goal models.
- **Smooth models:** a GAM Poisson model that can learn curved relationships.
- **Result classifiers:** ordinal logistic, multinomial logistic, regularized
  multinomial, classification tree, and random forest result models.
- **Expanded population sweep:** rolling form, opponent form, head-to-head
  history, rest, experience, tournament context, GBM, random forest, SVM, and
  shrinkage baselines.
- **XGBoost goals:** a gradient-boosted count model when the local `xgboost`
  package is installed.
- **Random forest goals:** a free local tree fallback while XGBoost is not
  installed.

How to read the tables:

- **RMSE** is the average size of the goal prediction error. Lower is better.
- **MAE** is the average absolute goal error. Lower is better.
- **Brier score** measures probability quality for win, draw, and loss forecasts.
  Lower is better.
- **Log loss** heavily penalizes confident wrong probabilities. Lower is better.
- **Accuracy** is the share of matches where the top predicted result matched
  the actual result. Higher is better.
- **Time split** is the stricter validation check because newer matches are held
  out from older training data.

## Challenger Leaderboard

```{r leaderboard, echo=FALSE, results="asis"}
if (nrow(challengers) > 0) {
  cards <- challengers |>
    dplyr::arrange(test_rmse) |>
    dplyr::mutate(
      label = model,
      value = paste0(fmt_number(test_rmse, 3), " RMSE"),
      note = paste0(model_family, " | MAE ", fmt_number(test_mae, 3))
    ) |>
    dplyr::select(label, value, note)
  cat(metric_cards(cards))
} else {
  display_table(empty_note("Challenger metrics are available after running R/18_fit_model_challengers.R."))
}
```

```{r metric-table}
if (nrow(challengers) > 0) {
  challengers |>
    dplyr::arrange(test_rmse) |>
    dplyr::mutate(
      test_rmse = fmt_number(test_rmse, 3),
      test_mae = fmt_number(test_mae, 3),
      mean_prediction = fmt_number(mean_prediction, 3),
      rows_train = fmt_integer(rows_train),
      rows_test = fmt_integer(rows_test)
    ) |>
    display_table(caption = "Held-out challenger model performance.")
} else {
  display_table(empty_note("No challenger metrics are available yet."))
}
```

## Expanded Goal Model Suite

This section compares goal-count models on the same historical modeling frame.
These are candidate engines for projected scores and expected goals.

```{r expanded-goal-cards, echo=FALSE, results="asis"}
if (nrow(expanded_goals) > 0) {
  goal_cards <- expanded_goals |>
    dplyr::filter(status == "fit") |>
    dplyr::group_by(validation_type) |>
    dplyr::slice_min(test_rmse, n = 1, with_ties = FALSE) |>
    dplyr::ungroup() |>
    dplyr::mutate(
      label = validation_label(validation_type),
      value = paste0(fmt_number(test_rmse, 3), " RMSE"),
      note = paste0(model, " | MAE ", fmt_number(test_mae, 3))
    ) |>
    dplyr::select(label, value, note)
  cat(metric_cards(goal_cards))
} else {
  display_table(empty_note("Expanded goal metrics are available after running R/21_fit_expanded_model_suite.R."))
}
```

```{r expanded-goal-table}
if (nrow(expanded_goals) > 0) {
  expanded_goals |>
    dplyr::arrange(validation_type, test_rmse) |>
    dplyr::mutate(
      validation_type = validation_label(validation_type),
      test_rmse = fmt_number(test_rmse, 3),
      test_mae = fmt_number(test_mae, 3),
      mean_prediction = fmt_number(mean_prediction, 3),
      mean_poisson_deviance = fmt_number(mean_poisson_deviance, 3),
      rows_train = fmt_integer(rows_train),
      rows_test = fmt_integer(rows_test)
    ) |>
    dplyr::select(
      validation_type,
      model,
      model_family,
      status,
      rows_train,
      rows_test,
      test_rmse,
      test_mae,
      mean_prediction,
      mean_poisson_deviance,
      note
    ) |>
    display_table(caption = "Expanded held-out goal model comparison.")
} else {
  display_table(empty_note("No expanded goal model metrics are available yet."))
}
```

```{r expanded-goal-plot, fig.cap="Goal-model RMSE by validation split. Lower is better."}
if (nrow(expanded_goals) > 0 && requireNamespace("ggplot2", quietly = TRUE)) {
  goal_plot <- expanded_goals |>
    dplyr::filter(status == "fit") |>
    dplyr::mutate(
      validation_type = validation_label(validation_type),
      model = reorder(model, test_rmse)
    )

  ggplot2::ggplot(goal_plot, ggplot2::aes(x = test_rmse, y = model)) +
    ggplot2::geom_col(fill = "#0f6b68", width = 0.66) +
    ggplot2::geom_text(
      ggplot2::aes(label = round(test_rmse, 3)),
      hjust = -0.08,
      color = "#24313d",
      size = 3.4
    ) +
    ggplot2::facet_wrap(~ validation_type, scales = "free_y") +
    ggplot2::coord_cartesian(xlim = c(0, max(goal_plot$test_rmse, na.rm = TRUE) * 1.14)) +
    ggplot2::labs(
      title = "Expanded goal model comparison",
      x = "RMSE",
      y = NULL
    ) +
    ggplot2::theme_minimal(base_size = 12)
} else {
  plot.new()
  text(0.5, 0.5, "Expanded goal RMSE plot is available after the model suite is refreshed.")
}
```

## Expanded Result Probability Suite

This section compares models that estimate the probability of loss, draw, or
win. For knockout matches, this is still regulation-time result modeling; the
advancement layer handles extra time and penalties separately.

```{r expanded-result-cards, echo=FALSE, results="asis"}
if (nrow(expanded_results) > 0) {
  result_cards <- expanded_results |>
    dplyr::filter(status == "fit") |>
    dplyr::group_by(validation_type) |>
    dplyr::slice_min(test_multiclass_brier, n = 1, with_ties = FALSE) |>
    dplyr::ungroup() |>
    dplyr::mutate(
      label = validation_label(validation_type),
      value = paste0(fmt_number(test_multiclass_brier, 3), " Brier"),
      note = paste0(model, " | accuracy ", fmt_percent(test_accuracy, 1))
    ) |>
    dplyr::select(label, value, note)
  cat(metric_cards(result_cards))
} else {
  display_table(empty_note("Expanded result metrics are available after running R/21_fit_expanded_model_suite.R."))
}
```

```{r expanded-result-table}
if (nrow(expanded_results) > 0) {
  expanded_results |>
    dplyr::arrange(validation_type, test_multiclass_brier) |>
    dplyr::mutate(
      validation_type = validation_label(validation_type),
      test_accuracy = fmt_percent(test_accuracy, 1),
      test_log_loss = fmt_number(test_log_loss, 3),
      test_multiclass_brier = fmt_number(test_multiclass_brier, 3),
      baseline_majority_accuracy = fmt_percent(baseline_majority_accuracy, 1),
      rows_train = fmt_integer(rows_train),
      rows_test = fmt_integer(rows_test)
    ) |>
    dplyr::select(
      validation_type,
      model,
      model_family,
      status,
      rows_train,
      rows_test,
      test_accuracy,
      test_log_loss,
      test_multiclass_brier,
      baseline_majority_accuracy,
      note
    ) |>
    display_table(caption = "Expanded held-out win/draw/loss probability comparison.")
} else {
  display_table(empty_note("No expanded result model metrics are available yet."))
}
```

```{r expanded-result-plot, fig.cap="Result-model Brier score by validation split. Lower is better."}
if (nrow(expanded_results) > 0 && requireNamespace("ggplot2", quietly = TRUE)) {
  result_plot <- expanded_results |>
    dplyr::filter(status == "fit") |>
    dplyr::mutate(
      validation_type = validation_label(validation_type),
      model = reorder(model, test_multiclass_brier)
    )

  ggplot2::ggplot(result_plot, ggplot2::aes(x = test_multiclass_brier, y = model)) +
    ggplot2::geom_col(fill = "#d9a441", width = 0.66) +
    ggplot2::geom_text(
      ggplot2::aes(label = round(test_multiclass_brier, 3)),
      hjust = -0.08,
      color = "#24313d",
      size = 3.4
    ) +
    ggplot2::facet_wrap(~ validation_type, scales = "free_y") +
    ggplot2::coord_cartesian(xlim = c(0, max(result_plot$test_multiclass_brier, na.rm = TRUE) * 1.14)) +
    ggplot2::labs(
      title = "Expanded result probability comparison",
      x = "Brier score",
      y = NULL
    ) +
    ggplot2::theme_minimal(base_size = 12)
} else {
  plot.new()
  text(0.5, 0.5, "Expanded result probability plot is available after the model suite is refreshed.")
}
```

## Expanded Population Feature Sweep

This sweep uses a wider local R feature population: historical rows, rolling
team form, opponent form, head-to-head history, rest spacing, team experience,
and tournament context. These are not automatically promoted. They first need
calibration checks, knockout handling review, and repeatable backtesting.

```{r population-summary, echo=FALSE, results="asis"}
if (nrow(population_summary) > 0) {
  pop_cards <- population_summary |>
    dplyr::filter(metric %in% c("rows", "matches", "teams", "feature_columns")) |>
    dplyr::mutate(
      label = dplyr::case_when(
        metric == "rows" ~ "Team-Match Rows",
        metric == "matches" ~ "Matches",
        metric == "teams" ~ "Teams",
        metric == "feature_columns" ~ "Feature Columns",
        TRUE ~ metric
      ),
      note = "expanded challenger population"
    ) |>
    dplyr::select(label, value, note)
  cat(metric_cards(pop_cards))
} else {
  display_table(empty_note("Expanded population summary is available after running R/22_build_expanded_feature_population.R."))
}
```

```{r population-goal-cards, echo=FALSE, results="asis"}
if (nrow(population_goals) > 0) {
  cards <- population_goals |>
    dplyr::filter(status == "fit") |>
    dplyr::group_by(validation_type) |>
    dplyr::slice_min(test_rmse, n = 1, with_ties = FALSE) |>
    dplyr::ungroup() |>
    dplyr::mutate(
      label = validation_label(validation_type),
      value = paste0(fmt_number(test_rmse, 3), " RMSE"),
      note = paste0(model, " | MAE ", fmt_number(test_mae, 3))
    ) |>
    dplyr::select(label, value, note)
  cat(metric_cards(cards))
} else {
  display_table(empty_note("Population goal metrics are available after running R/23_fit_population_expansion_models.R."))
}
```

```{r population-goal-table}
if (nrow(population_goals) > 0) {
  population_goals |>
    dplyr::filter(status == "fit") |>
    dplyr::arrange(validation_type, test_rmse) |>
    dplyr::mutate(
      validation_type = validation_label(validation_type),
      test_rmse = fmt_number(test_rmse, 3),
      test_mae = fmt_number(test_mae, 3),
      mean_prediction = fmt_number(mean_prediction, 3),
      mean_poisson_deviance = fmt_number(mean_poisson_deviance, 3),
      rows_train = fmt_integer(rows_train),
      rows_test = fmt_integer(rows_test)
    ) |>
    dplyr::select(validation_type, model, model_family, rows_train, rows_test, test_rmse, test_mae, mean_prediction, mean_poisson_deviance, note) |>
    display_table(caption = "Expanded population goal-model comparison.")
} else {
  display_table(empty_note("No expanded population goal metrics are available yet."))
}
```

```{r population-result-cards, echo=FALSE, results="asis"}
if (nrow(population_results) > 0) {
  cards <- population_results |>
    dplyr::filter(status == "fit") |>
    dplyr::group_by(validation_type) |>
    dplyr::slice_min(test_multiclass_brier, n = 1, with_ties = FALSE) |>
    dplyr::ungroup() |>
    dplyr::mutate(
      label = validation_label(validation_type),
      value = paste0(fmt_number(test_multiclass_brier, 3), " Brier"),
      note = paste0(model, " | accuracy ", fmt_percent(test_accuracy, 1))
    ) |>
    dplyr::select(label, value, note)
  cat(metric_cards(cards))
} else {
  display_table(empty_note("Population result metrics are available after running R/23_fit_population_expansion_models.R."))
}
```

```{r population-result-table}
if (nrow(population_results) > 0) {
  population_results |>
    dplyr::filter(status == "fit") |>
    dplyr::arrange(validation_type, test_multiclass_brier) |>
    dplyr::mutate(
      validation_type = validation_label(validation_type),
      test_accuracy = fmt_percent(test_accuracy, 1),
      test_log_loss = fmt_number(test_log_loss, 3),
      test_multiclass_brier = fmt_number(test_multiclass_brier, 3),
      baseline_majority_accuracy = fmt_percent(baseline_majority_accuracy, 1),
      rows_train = fmt_integer(rows_train),
      rows_test = fmt_integer(rows_test)
    ) |>
    dplyr::select(validation_type, model, model_family, rows_train, rows_test, test_accuracy, test_log_loss, test_multiclass_brier, baseline_majority_accuracy, note) |>
    display_table(caption = "Expanded population win/draw/loss probability comparison.")
} else {
  display_table(empty_note("No expanded population result metrics are available yet."))
}
```

## RMSE Visual

```{r rmse-plot, fig.cap="Held-out RMSE for the challenger goal models. Lower is better."}
if (nrow(challengers) > 0 && requireNamespace("ggplot2", quietly = TRUE)) {
  plot_data <- challengers |>
    dplyr::arrange(test_rmse) |>
    dplyr::mutate(model = factor(model, levels = rev(model)))

  ggplot2::ggplot(plot_data, ggplot2::aes(x = test_rmse, y = model, fill = model_family)) +
    ggplot2::geom_col(width = 0.64, show.legend = FALSE) +
    ggplot2::geom_text(
      ggplot2::aes(label = round(test_rmse, 3)),
      hjust = -0.12,
      color = "#24313d",
      size = 3.7
    ) +
    ggplot2::scale_fill_manual(values = c(
      "Interpretable regression" = "#0f6b68",
      "Gradient-boosted trees" = "#d9a441",
      "Tree ensemble fallback" = "#3d5873"
    )) +
    ggplot2::coord_cartesian(xlim = c(0, max(plot_data$test_rmse, na.rm = TRUE) * 1.14)) +
    ggplot2::labs(
      title = "Goal model challenger leaderboard",
      x = "RMSE",
      y = NULL
    ) +
    ggplot2::theme_minimal(base_size = 12)
} else {
  plot.new()
  text(0.5, 0.5, "RMSE plot is available after challenger metrics are refreshed.")
}
```

## Feature Importance

```{r feature-importance-plot, fig.cap="Top feature importance by challenger model. Importance is scaled within each model."}
if (nrow(importance) > 0 && requireNamespace("ggplot2", quietly = TRUE)) {
  top_importance <- importance |>
    dplyr::group_by(model) |>
    dplyr::slice_max(importance_scaled, n = 6, with_ties = FALSE) |>
    dplyr::ungroup() |>
    dplyr::mutate(feature = gsub("TRUE$", "", feature))

  ggplot2::ggplot(top_importance, ggplot2::aes(x = importance_scaled, y = reorder(feature, importance_scaled))) +
    ggplot2::geom_col(fill = "#0f6b68", width = 0.68) +
    ggplot2::facet_wrap(~ model, scales = "free_y") +
    ggplot2::scale_x_continuous(labels = function(x) paste0(round(100 * x), "%")) +
    ggplot2::labs(
      title = "Top features by challenger",
      x = "Scaled importance",
      y = NULL
    ) +
    ggplot2::theme_minimal(base_size = 12)
} else {
  plot.new()
  text(0.5, 0.5, "Feature importance is available after challenger models are fit.")
}
```

```{r feature-importance-table}
if (nrow(importance) > 0) {
  importance |>
    dplyr::group_by(model) |>
    dplyr::slice_max(importance_scaled, n = 8, with_ties = FALSE) |>
    dplyr::ungroup() |>
    dplyr::mutate(
      importance = fmt_number(importance, 3),
      importance_scaled = fmt_percent(importance_scaled, 1)
    ) |>
    display_table(caption = "Top model-specific feature importance.")
} else {
  display_table(empty_note("Feature importance is available after challenger models are fit."))
}
```

## Package Status

```{r xgboost-status}
if (nrow(status) > 0) {
  status |>
    display_table(caption = "Local challenger package status.")
} else {
  display_table(empty_note("Package status is available after the challenger run."))
}
```

```{r expanded-status}
if (nrow(expanded_status) > 0) {
  expanded_status |>
    display_table(caption = "Expanded model-suite package status.")
} else {
  display_table(empty_note("Expanded package status is available after the expanded model-suite run."))
}
```

```{r population-status}
if (nrow(population_status) > 0) {
  population_status |>
    display_table(caption = "Expanded population sweep package status.")
} else {
  display_table(empty_note("Expanded population package status is available after the population sweep."))
}
```

## Prediction Sample

```{r prediction-sample}
if (nrow(predictions) > 0) {
  predictions |>
    dplyr::arrange(model, dplyr::desc(abs(residual))) |>
    dplyr::mutate(
      predicted_goals = fmt_number(predicted_goals, 2),
      residual = fmt_number(residual, 2)
    ) |>
    dplyr::select(model, date, team, opponent, actual_goals, predicted_goals, residual) |>
    display_table(caption = "Largest held-out challenger residuals.", max_rows = 20)
} else {
  display_table(empty_note("Prediction samples are available after challenger models are fit."))
}
```

## Interpretation

The challenger models are useful for two separate reasons:

- Stepwise regression helps simplify the equation while staying interpretable.
- Tree models can capture nonlinear interactions without hand-writing every term.
- Count models better match the fact that goals are non-negative integers.
- Smooth models can improve accuracy when the relationship between team strength
  and goals is curved rather than straight.
- Multinomial and ordinal models give direct win, draw, and loss probabilities.

The tree results should be treated as a candidate signal, not an automatic upgrade.
Before promotion, the next pass should compare calibration, tournament backtesting,
knockout advancement behavior, and stability across multiple time splits.

## Reproduce Locally

```{r reproduce, eval=FALSE}
source("R/18_fit_model_challengers.R")
source("R/21_fit_expanded_model_suite.R")
source("R/22_build_expanded_feature_population.R")
source("R/23_fit_population_expansion_models.R")
```
