---
title: "Film Study State Engine"
output:
  html_document:
    toc: true
    toc_depth: 2
params:
  data_dir: "data/processed/film_study"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
data_dir <- params$data_dir

read_csv_if_exists <- function(name) {
  path <- file.path(data_dir, name)
  if (!file.exists(path)) {
    return(data.frame())
  }
  read.csv(path, stringsAsFactors = FALSE)
}

summary_path <- file.path(data_dir, "film_study_state_engine_summary.json")
summary_lines <- if (file.exists(summary_path)) readLines(summary_path, warn = FALSE) else "State engine not built yet."
state_summary <- read_csv_if_exists("film_study_state_engine_summary.csv")
transition_summary <- read_csv_if_exists("film_study_state_engine_transitions.csv")
snapshot <- read_csv_if_exists("film_study_state_engine_current_snapshot.csv")
```

# Overview

This private state engine treats each tagged event state like a move position. The pitch is split into a 12 by 8 grid, and each state is anchored to the current event family plus the current grid cell. A board-style alias is also created so the same cell can be read as a coordinate such as `F5` or `J2`. From each state it estimates:

- immediate next-move probabilities
- possession-level downstream outcome probabilities
- a simple advantage score based on those downstream outcomes
- likely next event type and next grid destination
- average grid progress, step length, and whether the next action tends to enter dangerous territory

The board coordinates are a reading aid, not a claim that football behaves like chess. The point is to turn continuous movement into a stable state graph that can be modeled as transitions, paths, and downstream value.

# Engine Summary

```{r engine-summary}
cat(paste(summary_lines, collapse = "\n"))
```

# State Table

```{r state-table}
state_summary
```

# Reading the State Table

- `board_cell` is the board-style alias for the current grid cell.
- `avg_forward_cell_progress` is the average column gain to the next tagged event.
- `final_third_entry_rate` is the share of visits from this state that move into the final third on the next tagged action.
- `most_likely_next_board_cell` is the destination cell the empirical engine sees most often from this state.

# Transition Table

```{r transition-table}
transition_summary
```

# Current-State Snapshot

```{r snapshot}
snapshot
```
